### 2015-12-07 paper/Paper_sweave.Rnw
test

\documentclass{article}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{\bf Comparing the Power of Plot Designs to Reveal Correlation}

\author{Dianne Cook\\
Department of Econometrics and Business Statistics, Monash University\\
and\\
Nathaniel Tomasetti\\
Department of Econometrics and Business Statistics, Monash University\\
  \maketitle
}

\begin{abstract}
Visual inference in EDA is prone to type 1 errors from the over-interpretation of randomness [4, 6]. Two competing plot designs, the scatter plot and overlaid line graph are both popular in the analysis of time series data. Lineups [2, 10, 11, 16, 27] allow the visual inference power of a graphic display to be evaluated, and were used to compare the plot designs. We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the decision for 2088 lineup evaluations. The results show that the scatter plot is both the faster and more powerful plot design, despite its inability to display the time dimension. 
\end{abstract}

\noindent
{\it Keywords:}  Lineups, visual inference, power comparison, scatter plot, line graph, data visualisation, visual analytics, time series, temporal data
\vfill

\begin{multicols}{2}
\twocolumn

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)

gen_true_data <- function(n, r, dep=TRUE, smoothed = FALSE) {
  d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2))) ##True plot, 0.5 for sigma off-diagonals requires ~300 iterations for all lineups.
  if (dep) {
    d$X1 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X1))
    d$X2 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X2))
  }
  if (smoothed) {
    d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
    d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
    d <- as.data.frame(d)
  }
  d$X1 <- scale(d$X1)
  d$X2 <- scale(d$X2)
  d$t <- 1:n
  return(d)
}

gen_null <- function(n, m=20, dep=TRUE, smoothed = FALSE){
  nd <- NULL
  for(i in 1:(m-1)) { 
    d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, 0, 0, 1), 2, 2)))
    if (dep) {
      d$X1 <- as.vector(arima.sim(list(ar=n/100), n))
      d$X2 <- as.vector(arima.sim(list(ar=n/100), n))
    }
    if (smoothed) {
      d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
      d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
      d <- as.data.frame(d)
    }
    d$X1 <- scale(d$X1)
    d$X2 <- scale(d$X2)
    nd <- rbind(nd, d)
  } 
  nd$.n <- rep(1:(m-1), each = n) 
  nd$t <- 1:n
  return(nd)
}
@

\section{Introduction}

In order to work with data, it first must be understood. Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first. This is especially so today, for vast databases that have been assembled in the big data era, that now need the data scientist to unravel the meaning of the numbers. Without preset hypotheses to test, the power of statistical inference is impotent, and without hypotheses the data analyst can stumble blindly trying to build up models of structure in the data.  To understand data requires good visualisation. This idea was formed as early as the 18th century, when William Playfair institutionalized the then revolutionary idea of graphing government and economic data. Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information [20]. Since then, advances in computing power have allowed statistical graphics to flourish, spearheaded by Tukey in 1965 into the new domain of exploratory data analysis (EDA) [7, 23]. EDA can be thought of as a well understood [3, 25] set of tools and techniques required to visualise information, to physically see what the data contains. In particular, it incorporates a free roaming approach, where the analyst is able to explore structure to find whatever relationships and structure exists within. Critically, the analyst does not have to have any pre-conceived ideas or hypotheses -- they are not specifically looking for any one particular thing. EDA emphasises letting the data inform us and can lead to the discovery of otherwise unexpected relationships, many of which may seem to become completely obvious after discovery. With the knowledge provided by EDA, ideas are generated about what relationships between variables may potentially exist. This then enables the analyst to use these new hypotheses upon which to apply classical inference and rigorously conduct tests with new data. EDA is also related to the field of model diagnostics (MD), where a model can be continuously refined through the visualisation of its fit, its residuals, and the interactions with its variables. Both EDA and MD follow the same framework: Visualise the data, look for patterns that suggest an underlying relationship, and if one is found implement it into the model then continue the exploration of the data. It has long been thought that EDA and statistical inference were worlds apart, but recent work [2, 16] bridges this chasm. Framing a data plot as a test statistic, which when compared to plots of null data, places EDA into the statistical inference recipe. 
The null hypothesis underlying a particular plot, is generically that there is no pattern, and particular types of plots implicitly regulate what “no  pattern” means. For example, a scatter plot of two variables is used to explore for some sort of association, so the implicit hypothesis is that there is no association between the two variables.  The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association. The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true. If an impartial observer asked to pick the plot that is different from the rest, picks the data plot this suggests there exists a pattern that is not the result of chance, and the null hypothesis is rejected. For a scatter plot, this departure from the null, might be a single outlier, or few outliers, a non-linear pattern, or clusters, which the human eye detects as more different in this data plot than in any of the nulls. This is the reason why visualisation remains important today, human eyes can detect patterns which would not be detected mathematically.  But eyes need calibration, which the lineup protocol provides.
On its own, with a single plot, because of random sampling in collecting data, it is easy for the analyst to imagine a pattern when no real structure exists in the population. Visual skills of an observer in EDA is prone to Type I error, where the null hypothesis is rejected when it is actually true, caused by the inherently random formation of patterns when visualised which are attributed to structure rather than chance. Daniel [4] warns against this, by providing 40 pages of plots from the null distribution, he encourages data analysts to understand the patterns that can be created from data without inherent structure. (This is what Buja et al [2] call the Rorschach protocol.) By being aware of what can appear in this type of data, the analyst should be more wary of claiming any visual feature they find is a true relationship. But this is not enough, even seasoned data analysts can be misled.  Diaconis [6] introduced the notion of 'magical thinking' which argues that people commonly suffer from the over-interpretation of randomness, particularly if it matches a pre-conception. If it suits their particular bias, an analyst may make false discoveries of structure and false rejections of the null hypothesis. Whilst using both can complement each other, the treatment of Type I error has led EDA and inferential statistics to be considered as very disparate pursuits. The mathematical rigor that governs classical inference relies on a framework that acknowledges and controls for the rate of Type I error.  Section 2 describes the visual inference protocols which put EDA more firmly into the rigorous framework of statistical inference.

\begin{figure}[hbtp]
\centering
\includegraphics{cny_line.png}
\caption{Figure 1: Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as an overlaid line graph. (Both sets of values were standardized to bring them to the same scale for plotting.) Do you think a relationship exists between the two?}
\end{figure}

In this research we utilise the methods in visual inference to study plot design to investigate the relationship between two temporal variables. There is a substantial literature suggesting that scatter plots are the appropriate display to read association [3, 8, 14], however, when the two variables are temporal it is common to display them as time series, drawn on the same plot. Temporal dependence is present in many macroeconomic variables [16, 17] and there is a strong argument for its presence in financial data [21, 24]. It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data.

\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics{nzd_scatter.png}
\end{subfigure}
\begin{subfigure}[t]{0.49\textwidth}
\includegraphics{nzd_line.png}
\end{subfigure}
\caption{Figure 2: Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatter plot (left) and overlaid line graph (right). Which one can you read the correlation from more accurately? Point-wise linear correlation between the two series, ignoring autodependence, is 0.64.}
\end{figure}

To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26]. This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure 1), even if no relationship exists. Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18]. However, keep in mind magical thinking [6], the desire to find any pattern may lead to a false discovery of positive correlation. The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier. If positive correlation does not exist, the observer may be fooled into thinking it does anyway. Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot. Figure 1 highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together. However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship. Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm. Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure 3). When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small. However, this is incorrect, as looking at the small horizontal distance is comparing the lines at different dates. It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993. Vanderplas [25] examines this effect in more detail. It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope. As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice. However, in practice many data analysts are split between the two major alternatives for temporal data. The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.

\begin{figure}[hbtp]
\centering
\includegraphics{murders.jpg}
\caption{Figure 3 [19]: Murders in New York State (Orange) and New York City (Blue). The difference between the two represents murders outside of the city. What size would you judge the difference to be in 1994?}
\end{figure}

There has been other research on the perception of  temporal displays, e.g. Javed et al. [12], but they do not examine perception of association. Javed et al.  examined optimal ways to visualise local features, such as, which variable had the highest value at a given point in time, and global features, such as, comparing the size of the overall slope of many different variables. The finding was that a different graphical layout is optimal for each type of task. However, they did not test for the perception of correlation, which can be treated as both a local and a global feature within the graph. The slope of each variable must be compared at a point in time, and they must have a persistent relationship across at least the majority of the series. However there is a quandary, the visual features that improve local tasks are unsuited to global tasks and vice versa; so we must now find a form that is well suited to assess both types of tasks simultaneously. This paper addresses this deficiency. Section 2 describes the lineup method utilised to rigorously compare plot designs. Section 3 explains the experimental design.  Section 4 contains the results of the experiment and Section 5 discusses the implications of the findings.

\section{The Lineup Protocol}

With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing? In order to use visualisation effectively, we will get the most benefit if the type of graphic display chosen is best suited to the task at hand, be it EDA, MD, or even the presentation of results. We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within. This can be found by using the lineup protocol developed in 2009 by Buja et al. [2], which is easily implemented with the nullabor R package. They created a statistically rigorous framework with properties explored primarily in Majumder et. al [16] and further in Hofmann et. al [11] that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
The lineup protocol is inspired in part by the police lineup [27]. We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people). If the real data plot is selected by an uninvolved observer as being most different from the other plots there is evidence that the structure of that data has led to a significant difference in the plot (That the criminal is sufficiently different from the innocents). This constitutes a rejection of the null hypothesis. If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error. p_i, the probability that plot i is chosen in the lineup depends not only on the plot design d, but also the signal strength, q_i, of that plot and of every other competing plot in the lineup. It can be defined as some unknown function f_(i,d) (q_1,…,q_20) [11]. If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution. The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test. For a lineup of m plots, the probability of selecting any plot when they are all generated by the null distribution is 1/m, setting the Type 1 error rate, α, of a lineup hypothesis test to equal 1/m. To control our error rate, it is initially obvious that we can simply change m, with an increase in null lineups having an inverse reduction in m. We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots. A much more powerful option is to have multiple different viewers of each lineup, with K observers; the probability that at a particular plot was picked at least x times under the null hypothesis is binomially distributed [2, 11] with:

$$p-value = P(X \geq x | H_{0}) = 1 - B_{n, 1/m}(x-1)$$

If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ . This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$. The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.

\begin{figure}[hbtp]
\centering
<<lineup.example, dependson='setup', fig.width=6, fig.height=6, outwidth='0.7//textwidth', echo=FALSE>>=
set.seed(1234)
a <- gen_true_data(24, 0.95)
td <- melt(a, id="t")
b <- gen_null(24)
b$.n <- NULL
nd_l <- melt(b, id="t")
nd_l$.n <- rep(1:19, rep(24, 19))
pos <- 7
td <- data.frame(.n=rep(pos, nrow(td)), td)
lg <- nd_l$.n < pos
nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
                    value=nd_l$value[lg])
nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
                    value=nd_l$value[!lg])
nd_l2$.n <- nd_l2$.n + 1
d <- rbind(nd_l1, td, nd_l2)

ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
    theme_bw() + theme(axis.title.x = element_blank(),
                       axis.title.y = element_blank(),
                       axis.text.x = element_blank(),
                       axis.text.y = element_blank(),
                       axis.ticks = element_blank(),
                       legend.position="none")
@
\caption{Figure 4: A lineup ($m = 20$) of line graphs. Which plot shows the series with the strongest correlation?}
\end{figure}

Essentially, the more time our subjects pick the correct plot out of the lineup, the more confident we are that the real plot was visually significantly different to the nulls; and that the type of graphic involved has the power to display association. The lineup effectively allows us to conduct a hypothesis test on visual features, with the competing hypotheses:
\begin{indent}
H_0 : There is no association between the two series (evidence for H0: the associated data is indistinguishable from the nulls).
H_1: There is an association between the two series (evidence against H0: the associated data can be distinguished from the nulls).
\end{indent}

As the power of a statistical test is defined as the probability of rejecting $H_{0}$ when it is false, the power of a lineup test is viewed as the probability of detecting the true plot. We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ correctly detected the true plot out of the lineup.
We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as [10]:

$$\hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2n-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /n_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /n_{2} }$$

Where $\hat\pi_{i} = (x_{i}+1)/(n_{i}+1)$, where $x$ is the number of times a true lineup was correctly identified and $n$ is the Welch-Satterwaite estimate for the degrees of freedom.

Some individuals may have a better natural ability at detecting the correct correlation pattern in the graphics, but by having each participant viewing multiple lineups and each lineup being viewed by multiple different participants, this can be controlled via a random effect variable in the model. We recruited participants through Amazon's Mechanical Turk, where the lineup protocol has been applied to a number of problems in prior papers [9, 16, 27]. Amazon’s Mechanical Turk [1] (MTurk) is a labour crowd-sourcing platform developed to give easy access to workers with basic tasks paid in line with the United States minimum wage. MTurk can be used to recruit subjects to read a variety of graphics and report on particular visual tasks. The time taken to decide, confidence in the decision and the reason for that decision can also be recorded. Heer and Bostock [9] use MTurk to replicate previous findings in graphical perception [3] and find that it is a valid method of data collection, once controls to eliminate anyone 'gaming' the system by randomly selecting answers to minimise time spent working are implemented. Further details of our use of Amazons’ Mechanical Turk is included in Section 3. The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in Hofmann et al [10], and this work follows that approach.

\section{Experimental Design}

\section{Results}

\section{Conclusion}

\end{document}

### 2015-12-07 paper/Paper_sweave.Rnw
Updated

@@
-\documentclass{article}
+\documentclass[12pt]{article}
+\usepackage{natbib}
+\usepackage{color}
+\usepackage[dvipsnames,svgnames*]{xcolor}
+\usepackage{array}
+\usepackage[colorlinks=TRUE, linkcolor=blue]{hyperref}
+\usepackage{wrapfig,float}
+\usepackage[font=small,skip=5pt]{caption}
+\usepackage{subcaption}
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{dsfont}
+\usepackage{amsthm}
+\usepackage{amsfonts}
+\usepackage{url}
+\usepackage{ulem}
+\usepackage[section]{placeins}
+\usepackage{afterpage}
+
+\graphicspath{{figure/}}
+\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
+
+\newcommand{\blue}[1]{{\color{blue} #1}}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\dc}[1]{{\color{orange} #1}}
+\newcommand{\green}[1]{{\color{cyan} #1}}
+
+\newtheorem{thm}{Theorem}[section]
+\newtheorem{dfn}{Definition}[section]
+\newtheorem{cor}{Corollary}[thm]
+\newtheorem{con}{Conjecture}[thm]
+
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
 
 \begin{document}
+
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
+
 \SweaveOpts{concordance=TRUE}
@@
 
+\if0\blind
+{
 \author{Dianne Cook\\
@@
 Nathaniel Tomasetti\\
-Department of Econometrics and Business Statistics, Monash University\\
+Department of Econometrics and Business Statistics, Monash University\\}
   \maketitle
-}
+} \fi
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Comparing the Power of Plot Designs to Reveal Correlation}
+\end{center}
+  \medskip
+} \fi
+
+\bigskip
 
@@
 
-\begin{multicols}{2}
-\twocolumn
-
-<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
+\newpage
+<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 
-In order to work with data, it first must be understood.
-Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first.
+In order to work with data, it first must be understood. Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first.
 This is especially so today, for vast databases that have been assembled in the big data era, that now need the data scientist to unravel the meaning of the numbers. 
@@
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics{cny_line.png}
-\caption{Figure 1: Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as an overlaid line graph.
-(Both sets of values were standardized to bring them to the same scale for plotting.) Do you think a relationship exists between the two?}
-\end{figure}
+Insert Figure 1
 
@@
 
-\begin{figure}[hbtp]
-\centering
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics{nzd_scatter.png}
-\end{subfigure}
-\begin{subfigure}[t]{0.49\textwidth}
-\includegraphics{nzd_line.png}
-\end{subfigure}
-\caption{Figure 2: Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatter plot (left) and overlaid line graph (right).
-Which one can you read the correlation from more accurately?
-Point-wise linear correlation between the two series, ignoring autodependence, is 0.64.}
-\end{figure}
+Insert Figure 2
 
@@
 
-\begin{figure}[hbtp]
-\centering
-\includegraphics{murders.jpg}
-\caption{Figure 3 [19]: Murders in New York State (Orange) and New York City (Blue).
-The difference between the two represents murders outside of the city.
-What size would you judge the difference to be in 1994?}
-\end{figure}
+Insert Figure 3
 
@@
 If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error.
-p_i, the probability that plot i is chosen in the lineup depends not only on the plot design d, but also the signal strength, q_i, of that plot and of every other competing plot in the lineup.
-It can be defined as some unknown function f_(i,d) (q_1,…,q_20) [11].
+$p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup.
+It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ [11].
 If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution.
@@
 However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test.
-For a lineup of m plots, the probability of selecting any plot when they are all generated by the null distribution is 1/m, setting the Type 1 error rate, α, of a lineup hypothesis test to equal 1/m.
-To control our error rate, it is initially obvious that we can simply change m, with an increase in null lineups having an inverse reduction in m.
+For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
+To control our error rate, it is initially obvious that we can simply change $m$, with an increase in null lineups having an inverse reduction in $\alpha$.
 We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with K observers; the probability that at a particular plot was picked at least x times under the null hypothesis is binomially distributed [2, 11] with:
+A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed [2, 11] with:
 
 $$p-value = P(X \geq x | H_{0}) = 1 - B_{n, 1/m}(x-1)$$
-
 If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
@@
 The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
-
 \begin{figure}[hbtp]
   \centering
-<<lineup.example, dependson='setup', fig.width=6, fig.height=6, outwidth='0.7//textwidth', echo=FALSE>>=
+<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
 set.seed(1234)
@@
 
-ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
+plot <- ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
     theme_bw() + theme(axis.title.x = element_blank(),
@@
                        legend.position="none")
+print(plot)
 @
-\caption{Figure 4: A lineup ($m = 20$) of line graphs. Which plot shows the series with the strongest correlation?}
+  \caption{A lineup ($m = 20$) of line graphs. Which plot shows the series with the strongest correlation?}
+  \label{lineup:example}
 \end{figure}
@@
 The lineup effectively allows us to conduct a hypothesis test on visual features, with the competing hypotheses:
-\begin{indent}
-H_0 : There is no association between the two series (evidence for H0: the associated data is indistinguishable from the nulls).
-H_1: There is an association between the two series (evidence against H0: the associated data can be distinguished from the nulls).
-\end{indent}
+
+\begin{itemize}
+  \item $H_{0}$ : There is no association between the two series (evidence for $H_{0}$: the associated data is indistinguishable from the nulls).
+  \item $H_{1}$: There is an association between the two series (evidence against $H_{0}$: the associated data can be distinguished from the nulls).
+\end{itemize}
 
@@
 
+To allow for a greater control of distributions of the data in the lineups, we use simulated data.
+The data must contain autocorrelation for there to be meaningful information in the time dimension.
+To fulfil this requirement, each plot has a standardised pair of AR(1) models defined by:
+
+$$Y_{1,t} = \beta \times Y_{1,t-1} + e_{1,t}$$
+
+$$Y_{2,t} = \beta \times Y_{2,t-1} + e_{2,t}$$
+
+With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ being generated with from a bivariate normal distribution:
+
+$$~N(\mu= \left| \begin{array}{c}0 \\0 \end{array} \right|,\hspace{4mm}\Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right|)$$
+
+The variation in the covariance parameter  allows for control in the correlation in the pairs of data.
+The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$.
+True data pairs are generated with $\rho = \pm 0.3, 0.5, 0.7, 0.9$.
+Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$.
+Further to this, relationships weaker than this point are often of little interest to analysts.
+Hence we found it unnecessary to test a more full range of correlations.
+Simulating data does not produce correlations exactly equal to $\rho$, so imulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$.
+The null pairs often had correlation generated spuriously through the simulation process.
+This was particularly problematic for generating small sample data, but setting, to a maximum of 0.5, for both true and null data reduced spurious correlation to manageable levels.
+For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications.
+To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component.
+This filter creates a trend by penalising both errors in smoothness and in fit.
+The trend of the series, $\tau_{t}$ is defined by the equation:
+
+$$\min_{\tau}(\sum_{t=1}^{T}(y_{t}-\tau_{t})^2 + \lambda \sum_{t=2}^{T-1}[(\tau_{t+1}-\tau_{t})-(\tau_{t}-\tau_{t-1})]^2)$$
+
+Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation.
+However, for $t = 12, 24$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = 48, 96$.
+Each generated lineup was produced as both a scatter plot and an overlaid line graph.
+
+\begin{figure}[hbtp]
+  \centering
+<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+set.seed(123)
+uns <- gen_true_data(96, 0.8)
+uns_l <- melt(uns, id="t")
+p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
+    theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
+    theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+
+set.seed(123)
+sm <- gen_true_data(96, 0.8, smoothed = TRUE)
+sm_l <- melt(sm, id="t")
+p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
+      theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
+    theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+
+grid.arrange(p1a, p2a, p3a, p4a, ncol=2)
+@
+  \caption{The same AR(1) data both smoothed and unsmoothed. t = 96, correlation = 0.86}
+  \label{smoothing}
+\end{figure}
+
+In this research, we had three basic factors:
+
+\begin{itemize}
+  \item Factor 1: Plot design, levels = scatter plot, overlaid time series
+  \item Factor 2: Sample size, levels = 12, 24, 48, 96
+  \item Factor 3: True plot correlation, levels 
+\end{itemize}
+
+Additionally, there was a fourth factor for  = 48, 96:
+
+\begin{itemize}
+  \item Factor 4: Smoothed, levels = yes, no
+\end{itemize}
+
+Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (2 plots x 8 correlations x 6 sample sizes & smoothness combinations). 
+The position of the actual data plot in the lineup was randomized.
+The order that the factors are presented to a subject were also randomized, and each subject saw 5 scatter plots and 5 overlaid lines at a range of all absolute correlation levels.
+Each subject did not see the same data more than once.
+Subjects were asked to pick the plot in the lineup that has the strongest association, with basic examples of negatively and positively correlated plots.
+They also answered with reasons for choosing the plot, and the confidence that they have that this really is the plot showing the strongest correlation.
+Each subject evaluated 10 lineups, plus a further two trials to eliminate people attempted to game the system.
+If a trial was not answered correctly, the subject could not evaluate the remaining lineups.
+If it was answered correctly, the trial data was discarded and the subject moved onto the rest of the lineups.
+Earlier work [15] has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
+We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
+In total there were 2088 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations leaving us with 1684 evaluated lineups for the analysis.
+
+\begin{figure}[hbtp]
+  \centering
+<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+set.seed(101)
+a <- gen_true_data(24, 0.95)
+b <- gen_true_data(24, -0.95)
+ca <- cor(a$X1, a$X2)
+cb <- cor(b$X1, b$X2)
+a_l <- melt(a, id="t")
+b_l <- melt(b, id="t")
+p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
+      theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
+      theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
+      theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
+      theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.x = element_blank(),
+          axis.text.y = element_blank(),
+          axis.ticks = element_blank(),
+          legend.position="none")
+grid.arrange(p1, p2, p3, p4, ncol=2)
+@
+  \label{designs}
+  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
+\end{figure}
+
+The web site that the Turk workers accessed to complete the task can be seen at http://104.236.245.153:8080/mahbub/turk18/index.html.
+You can also try out the tasks, but your results will not be recorded.
+(The experiment cost $210 to run.)
+
 \section{Results}

### 2015-12-07 paper/Paper_sweave.Rnw
Updated

@@
 library(gridExtra)
+library(splitstackshape)
+correl <- read.csv("correlations.csv")
+full <- read.csv("full.csv") 
+Turk <- read.csv("turk.csv")
 
@@
 
-$$p-value = P(X \geq x | H_{0}) = 1 - B_{n, 1/m}(x-1)$$
+\begin{equation} \label{eq:1}
+  \centering
+  p-value = P(X \geq x | H_{0}) = 1 - B_{n, 1/m}(x-1)
+\end{equation}
+
 If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
@@
 The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
-\begin{figure}[hbtp]
+\begin{figure}[htbp]
   \centering
@@
 
-$$\hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2n-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /n_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /n_{2} }$$
+\begin{equation} \label{eq:2}
+  \centering
+  \hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2n-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /n_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /n_{2} }
+\end{equation}
 
@@
 
-\begin{figure}[hbtp]
+\begin{figure}[htbp]
   \centering
@@
 
-Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (2 plots x 8 correlations x 6 sample sizes & smoothness combinations). 
+Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (2 plots x 8 correlations x 6 sample sizes \& smoothness combinations). 
 The position of the actual data plot in the lineup was randomized.
@@
 You can also try out the tasks, but your results will not be recorded.
-(The experiment cost $210 to run.)
+(The experiment cost \$210 to run.)
 
@@
 
+\begin{figure}[htbp]
+ \centering
+<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
+               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
+#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
+ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
+  scale_colour_discrete(name = "Plot Design") + labs(title="Observed Data", y="Percent Correct", x="Correlation") +
+  theme(legend.position="bottom")
+@
+  \label{results}
+  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size.
+The dots are data points, the lines represent a naive model average fit.
+Scatter plots generally showed superior performance, with the overlaid line graphs being particularly weak at revealing negative correlation}
+
+Figure 7 shows the raw results of the survey and Figure 8 shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
+Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
+Performance between the two plot designs appears to be similar for positive correlation, with the scatter plot power gradually improving relative to the overlaid lines as sample size increases.
+The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered.
+This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure 6).
+The right side shows the performance benefits of smoothing overlaid line time series data.
+Much of the 'noise' in the time series is removed and studying the overall trend shows a greatly improved ability to successfully detect the true data plot when it is negatively correlated.
+The scatter plot still appears to be the superior plot, but much of the gap between the two is reduced for positive correlation.
+
+\begin{figure}[htbp]
+  \centering
+<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+Turk$correct <- Turk$percent*Turk$attempt
+Turk$r <- round(Turk$correlation, 1)
+Turk.line <- subset(Turk, test_param=="line")
+Turk.scatter <- subset(Turk, test_param=="scatter")
+Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c(10, 13, 16, 19, 20)]
+Turk.line <- summarise(group_by(Turk.line, smoothed, n, r), sum(attempt), sum(correct))
+Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c(10, 13, 16, 19, 20)]
+Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
+colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
+colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
+prop <- cbind(Turk.line, Turk.scatter[4:5])
+prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
+               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
+               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
+               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
+ggplot(data=prop) + geom_line(aes(x=r, y=difference), colour="#99CC99", size=2) + 
+                    geom_ribbon(aes(x=r, ymin=lower, ymax=upper), colour="#99CC99", fill="#99CC99", alpha=0.2) + 
+                    labs(title="Difference in Power", x="Correlation") + facet_grid(n~smoothed)
+@
+ \label{powerdiff}
+ \caption{Difference in estimated power to reveal the true plot by smoothness and sample size with a 95\% confidence interval.
+Values above zero indicate the scatter plot is more powerful than the overlaid line graph.}
+\end{figure
+
+\begin{figure}[htbp]
+  \centering
+<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")}
+@
+ \label{samplesize}
+ \caption{Effects of increasing sample size for unsmoothed data. Both plot designs appear to benefit from increased sample sizes}
+\end{figure}
+
+Figure 9 demonstrates the effects of changing sample size for unsmoothed data.
+The scatter plot has increased power to detect correlation almost universally for any increase.
+Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
+Figures 10 and 11 give a breakdown of the interactions between the response variables: plot detection, time and confidence.
+Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
+The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed.
+We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a choice.
+Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to pick the true plot and 32.6 seconds if a null was selected.
+The scatter plots performed much better, with the true plot picked 52.9\% of the time, with an average of 22.7 seconds required to make a decision.
+If the true plot was detected it took only 18.6 seconds, and 29.3 seconds for a null plot selection.
+
+\begin{figure}[htbp]
+  \centering
+<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+full <- mutate(full, sign = ifelse(correlation >= 0, 1, 0), n24 = ifelse(n ==24, 1, 0), n48 = ifelse(n ==48,  1, 0), n96 = ifelse(n ==96, 1, 0), plot = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"))
+time <- data.frame(plot=full$plot, time=full$log_time, correct=full$correct, rsgn=full$sign)
+#ggplot(data=time, aes(x=time, fill=factor(correct))) +geom_bar(stat="bin") + facet_grid(~plot) +labs(x="Log(Time) Taken", y=element_blank()) + scale_fill_discrete(name = "Correct")
+time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
+tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
+t1 <- ggplot(data=subset(time, correct==0), aes(x=time)) + geom_histogram(binwidth=0.5) + 
+  geom_vline(data=subset(tm, correct==0), aes(xintercept = m), color="red") +
+  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
+t2 <- ggplot(data=subset(time, correct==1), aes(x=time)) + geom_histogram(binwidth=0.5) + 
+  geom_vline(data=subset(tm, correct==1), aes(xintercept = m), color="red") +
+  facet_grid(plot~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
+grid.arrange(t1, t2, ncol=2, top = "Log(Time) by plot design, true plot correlation and detection")
+@
+  \label{time} 
+  \caption{Histograms of log time taken to make a decision.
+r+ and r- indicate the sign of true plot correlation.
+The vertical red line is the average time of the particular group.
+On average scatter plots are faster and the true plot is detected more often, with r-.}
+\end{figure}
+
+Subjects were asked to give a number from 1 to 5 to rate how confident they were in there choice (1 - most confident, 5 - least confident, average = 3.34).
+There were no statistically significant differences between plot designs, however subjects were slightly less confident on true plot detections (by $0.25 \pm  0.16, p=0.002$) and every one percent extra time required to make a decision increased confidence (by $0.19 \pm 0.07, p<0.001$).
+One possible explanation is that subjects often focused on a null plot and picked it confidently.
+This would suggest that the null plots often produced visually interesting features subjects assumed were caused by the association between plots.
+
+Subjects spending longer on a plot immediately draws two hypotheses:
+\begin{enumerate}
+  \item The lineup was more difficult and it took longer to determine which plot was best, reducing confidence.
+  \item The subject picked a plot relatively easily but spent extra time going over its details and making comparisons to alternative plots, increasing confidence.
+\end{enumerate}
+
+The result suggests that the latter hypothesis is more likely to be correct.
+We also found a negative relationship between plot detection and time spent on a lineup, supporting the major relationship found: quick decisions are less confident but more likely to successfully detect the true plot.
+
+\begin{figure}[htbp]
+  \centering
+<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+conf <- data.frame(plot=full$plot, conf=full$conf_level, correct=full$correct, rsgn=full$sign)
+conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
+cm <- summarise(group_by(conf, correct, plot, rsgn), m=mean(conf))
+c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf)) + geom_histogram(binwidth=1) + 
+  geom_vline(data=subset(cm, correct==0), aes(xintercept = m), color="red") +
+  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
+c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf)) + geom_histogram(binwidth=1) + 
+  geom_vline(data=subset(cm, correct==1), aes(xintercept = m), color="red") +
+  facet_grid(plot~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
+grid.arrange(c1, c2, ncol=2, top = "Confidence by plot design, true plot correlation and detection")
+@
+  \label{confidence} 
+  \caption{Histograms of reported confidence levels.
+1 - Low, 5 - High.
+Detection had a minor detrimental impact on confidence, whilst time taken increased it.
+There was no statistically significant difference on confidence level between plot designs.}
+\end{figure}
+
+Next we analysed the individual ability of each of our subjects.
+Figure 12 plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
+While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two.
+The accuracy is normally distributed (Jarque-Bera p value = 0.418), with a mean of 42.5\% and a standard deviation of 17.8\%.
+The strongest subject successfully identified the true plot 91\% of the time, whilst the weakest was never able to identify the true plot.
+As all subjects were shown a balanced mix of true plot correlations, the extreme difference in individual visual ability is striking.
+Visual ability will be controlled via a subject specific random effect in the models in the next section.
+
+\begin{figure}[htbp]
+  \centering
+<<confidence2, dependson='confidence', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+full$subj_id <- factor(full$ip_address, labels=1:207)
+y <- data.frame(table(full$subj_id))
+keep2 <- y$Var1[y$Freq>9]
+sub <- filter(full, subj_id %in% keep2)
+sub <- sub[c(7, 15, 16, 17, 20, 27)]
+sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
+sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
+sub <- mutate(sub, prop = correct/Freq)
+ggplot(data=sub, aes(x=conf, y= prop)) + geom_point() + labs(title="Accuracy versus Confidence", y = "Accuracy", x = "Average Confidence")
+@
+  \label{confidence2}
+  \caption{Individual subject accuracy versus confidence. There does not appear to be any relationship between the two.}
+\end{figure}
+
+There are two major reasons why a subject may fail to detect the true plot out of a lineup.
+Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis).
+We next can analyse these kinds of errors with the pin plots found in Figure 13.
+The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot.
+The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis.
+The correlation of the simulated (true) data (r=-0.9, -0.7, -0.3, 0.3, 0.7, 0.9) in the plot, the sample size (t=12, 24, 48, 96), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
+The blue pins indicate the true plots and the red pins indicate the null plots.
+If the plot design has the power to display correlation, we would expect the highest correlation plot to be selected, leading to a Type I error where a plot from the null distribution spuriously had high correlation.
+We find that this is generally not the case.
+The selections appear to have a wide range of correlations, indicating that if a subject selected a null plot it was typically not due to excess spurious correlation.
+Instead the failure to select the highest correlated plot indicates that a Type II error was committed rather than Type I as the plot designs did not have the power to differentiate between medium and low correlation plots.
+However, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected.
+For the overlaid lines this mostly occurred only for plots with strong positive correlation, which is consistent with its general lack of power to reveal negative relationships in data.
+
+\begin{figure}[htbp]
+  \centering
+<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+correl <- read.csv("correlations.csv")
+pin <- full[c(5, 6, 10, 15, 16, 17)]
+pin <- cSplit(pin, "response_0", direction="long")
+counts <- data.frame(table(pin$pic_id, pin$response_0))
+counts$Var1 <- as.integer(counts$Var1)
+counts$Var2 <- as.integer(counts$Var2)
+colnames(counts) <- c("lineup", "plot", "count")
+Turka <- Turk[c(2, 10, 11, 13, 17, 20)]
+pin <- merge(counts, Turka, by.x="lineup", by.y="pic_id")
+pin$r <- factor(pin$r)
+head(pin)
+colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+pin$base <- 0
+pin <- read.csv("pin.csv") #Do this part through R, not excel
+pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
+pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
+pin <- mutate(pin, percent = count/attempt)
+ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
+ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
+ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
+ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
+
+#Code to be fixed
+@
+  \label{pins}
+  \caption{Pin plots of the subject selections of each lineup.
+On the x-axis is the correlation of the plot (there is potentially spurious correlation in the null plots), and on the y-axis is the proportion a plot was chosen.
+Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not appear as it has a frequency of zero and is visually blocked by the red pins.
+The pin plots are split with true plot correlation on the columns and sample size on the rows.
+Each plot contains information from all three lineups with that unique combination of factors: plot design, smoothness, sample size and true correlation.}
+\end{figure}
+
+Occasionally there was an anomaly where a single null plot was selected with a high frequency despite having a weak correlation.
+Occasionally the null distribution cangenerate visually interesting patterns that draw the analyst’s attention.
+We asked subjects for reasons behind their selections and found that choosing on the basis that “the points in the plot were close together”, or had “a unique pattern” more often led to an unsuccessful detection.
+Conversely, decisions based on “matching the peaks and troughs” of the overlaid lines or “seeing a line formed in the dots” of the scatter plot were more likely to be successful in identification of the true plot.
+However, these true detections are, on average, made less confidently than unsuccessful evaluations, which indicates that decisions made on the basis of closeness or patterns are more convincing of an association than the more successful reasons.
+
+\begin{figure}[htbp]
+  \centering
+Insert anomaly lineup 1
+  \label{anom1}
+  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
+The true plot in position 8 was not detected in five evaluations.
+However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
+When asked for a reason of their selections, subjects answered that the lines were closest together.}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+Insert anomaly lineup 2
+  \label{anom2}
+  \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
+The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}).
+The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037.
+This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness.
+Subjects most often cited the clustering pattern formed as the reason for their selection.
+The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
+\end{figure}
+
+Figures 14 and 15 show lineups where a null plot with a strong pattern led to its selection over the data plot.
+Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
+The true data plot is in position 4, but 9/11 subjects chose plot 6.
+Plot 6 is unusual, most of the points are scattered around the vertical median, but there are two outlying points, one with a high y-value and one with a low y-value.
+Other concurrent work has suggested that outliers are pre-attentive patterns and this plot selection by the observers is consistent with that.
+Figure 15 below is another example of where subjects chose a null plot that had smaller correlation than the data plot.
+Four out of five evaluations selected the null plot in position 13,  instead of the data plot in position 8.
+Plot 13 has a correlation of 0.225, smaller (in absolute value) than the -0.707 of plot 8.
+This selection is in-line with Wertheimer’s gestalt law of common fate.
+It appears to be strongly correlated in the beginning of the series but breaks apart towards the end.
+The selection implies that subjects selecting this plot focused on the joint path of the two series early in the plot and ignored the contradictory evidence of no relationship as they diverged later in the plot.
+It also confounded by the seeming preference to find positive correlation over negative correlation, and we suspect that if one of the series in plot 8 had the values flipped, then the strong positive correlation would have been noticed.
+
+\section{Modelling Subject Responses}
+
+A mixed effects logit model is used separately for scatter plot and overlaid line evaluations. The tested model is:
+
+$$Power_{i} = \psi\left\{ \beta_{i,0} + \beta_{i,1} \cdot Correlation + \beta_{i,2}\cdot Sign + \beta_{i,3}\cdot Smoothed + \beta_{i,4}\cdot t + \beta_{i,5}\cdot Correlation \cdot Sign + \gamma_{i,1} \cdot subject_{j} \right\}$$
+
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $r$ is the true plot correlation, Sign is the direction of the true plot correlation, $i$  is the plot design and $j$ is an observer specific random effect.
+
+Table 1: Mixed Effects Logit Model for the predicted power of plot design.
+Sign = 1 if correlation > 0, 0 otherwise.
+Each plot design is estimated separately.
+Results must be transformed via the logit link function to generate a power prediction.
+Subjects with less than ten lineup evaluations were removed from the sample.
+Sample size for scatter plot model: 841, overlaid lines model: 843.
+The model is fitted using the lme4 package in
+<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+full$subj_id <- factor(full$ip_address, labels=1:207)
+
+# Need to keep only evaluations done by subject who completed at least 10
+x <- data.frame(table(full$subj_id))
+keep <- x$Var1[x$Freq>9]
+full.sub <- filter(full, subj_id %in% keep)
+full.sub$r <- round(full.sub$correlation, 1)
+full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
+full.sub$rabs <- abs(full.sub$r)
+full.sub$rabs <- factor(full.sub$rabs)
+full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
+full.sub <- full.sub[,c(1,7, 15,16,17,20:26)]
+full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
+full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
+full.sub$alpha <- log(.05/0.95)
+
+m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "L"), family="binomial")
+m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "S"), family="binomial")
+@
+%Code to make a table to be added
+
+Table 1 summarises the model fits.
+It can be confusing the interpret the model estimates, and at first glance it appears the overlaid lines (-0.660) are substantially more powerful than the scatter plot (-4.403), but this is misleading, and  interpreting the the full model involves combining the estimates for the different interactions with correlation and sign.
+Taking correlation into account, the scatter plot shows superior performance.
+It needs to be noted that we only tested correlations at  0.3, 0.5, 0.7 and 0.9, so the results below an absolute value of 0.3 are unreliable and may be artifacts of regression smoothing methods.
+The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
+Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure 5).
+It did have a significant positive impact on the power of overlaid line graphs.
+Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
+Figure 16 shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
+The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
+Previous studies [15] show that this random visual skill is not impacted by demographics such as age and gender; whilst education increases the power of a graphic display by only approximately one percent, not a practically significant amount.
+We did not include these characteristics of the individual in the modelling due to this lack of significance.
+Figure 16 especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
+
+\begin{figure}[htbp]
+ \centering
+<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
+r <- seq(-0.9,0.9,0.02)
+n <- c(12, 24, 48, 96)
+full.pred8l <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
+                                      smoothed=c(0,1), 
+                                      subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
+full.pred8l$rsgn <- ifelse(full.pred8l$r > 0, 1, 0)
+full.pred8l$rabs <- abs(full.pred8l$r)
+full.pred8l$fitre <- predict(m8l, full.pred8l, type="response", re.form=~(1|subj_id),
+                             allow.new.levels=TRUE)
+full.pred8l$fitfx <- predict(m8l, full.pred8l, type="response", re.form=NA,
+                             allow.new.levels=TRUE)
+full.pred8l$plot <- ifelse(full.pred8l$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
+full.pred8l <- filter(full.pred8l, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
+
+full.pred8s <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
+                                      smoothed=c(0,1), 
+                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
+full.pred8s$rsgn <- ifelse(full.pred8s$r > 0, 1, 0)
+full.pred8s$rabs <- abs(full.pred8s$r)
+full.pred8s$fitre <- predict(m8s, full.pred8s, type="response", re.form=~(1|subj_id),
+                             allow.new.levels=TRUE)
+full.pred8s$fitfx <- predict(m8s, full.pred8s, type="response", re.form=NA,
+                             allow.new.levels=TRUE)
+full.pred8s$plot <- ifelse(full.pred8s$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
+full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
+full.pred2 <- rbind(full.pred8s, full.pred8l)
+
+full.pred2 <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))
+
+qplot(r, fitre, data=full.pred2, color=plot, 
+      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
+  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
+  xlab("Correlation") + ylab("Percent Detected") + labs(title="Predicted Power") +
+  geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
+@
+  \label{powermodel}
+  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
+\end{figure}
+
+\begin{figure}
+  \centering
+<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
+resids <- resid(m8s)
+residl <- resid(m8l)
+resids[842:843] <- NA
+resid <- data.frame(line = residl, scatter = resids)
+p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
+p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
+grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
+@
+  \label{residuals}
+  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
+\end{figure}
+
+Analysing the residuals allows us to test the model fit.
+As the dependent variable is a binary variable, the model will only have a residual close to zero if it predicts either a power of zero or one.
+As the rate of picking the true plot purely by chance is, $1/m$ or 0.05, the model should never predict a power of zero even if the plot has no ability to display correlation.
+This effect is especially notable in the overlaid line residual plot, the model fit is never close to zero or one so the residuals are split with no values around zero.
+The scatter plot residuals are normally distributed with a Jarque-Bera normality test statistic of 0.09 (p-value = 0.95), however the overlaid lines strongly reject normality with a Jarque-Bera statistic of 75.6 (p-value = 0.00).
+The significance tests associated with the overlaid lines are only valid asymptotically.
+Also concerning, the mean value of the overlaid lines residuals is - 0.08, implicating that the model tends to over-predict power, further weakening the performance of this plot design relative to the scatter plot.
+This unrestricted model used in to test the symmetry restriction for the scatter plot design is written formally as:
+
+$$Power_{UR,s} = \psi\left\{ \beta_{s,0} + \beta_{s,1} \cdot |Correlation| + \beta_{s,2}\cdot Sign + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t + \beta_{s,5}\cdot |Correlation| \cdot Sign + \gamma_{s,1} \cdot subject_{j} \right\}$$
+
+Testing if the scatter plot displays positive and negative correlation equally is equivalent to testing the hypothesis:
+
+\begin{itemize}
+  \item $H_{0}: \beta_{2} = \beta_{5} = 0$
+  \item $H_{1}:$ Either equality does not hold.
+\end{itemize}
+
+The restricted model is therefore:
+$$Power_{R,s} = \psi\left\{ \beta_{s,0}+\beta_{s,1}\cdot |Correlation| + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t +\gamma_{s,1} \cdot subject_{j} \right\}$$
+
+The unrestricted model has a log-likelihood value of -346.9, where the restricted model has a log-likelihood of -355.8.
+The test statistic is therefore, with a null distributed as a $\chi^{2}_{2}$ variable.
+The null hypothesis is rejected (p = 0.02).
+We find that the scatter plot has asymmetric performance over negative and positive values of correlation.
+This may be a result of unwittingly introducing bias into the methodology, the first scatter plot shown to subjects in the instruction set always had strong positive correlation.
+A similar setup rejects the symmetric performance of the overlaid line graphs with a log likelihood ratio test statistic of 105.30 (p-value < 0.001).
+Such an extreme result is due to a systematic difference in the overlaid line graph and cannot be attributed simply to bias.
+
+Table 2: Mixed Effects Linear Model with the dependent variable log(time).
+Sign = 1 if correlation > 0, 0 otherwise.
+Both plot designs estimated together with Overlaid Lines as the base design.
+Sample size: 1684.
+The model is fitted using the lme4 package in R.
+<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 
+                  + test_param*r + test_param*n + test_param*rsgn*r + test_param*correct + (1|subj_id), data=full.sub)
+@
+%get table code
+
+While it appears that the overlaid lines are on the verge of statistical significance (at the 5\% level) for being the faster plot for a subject to visually process, this effect only holds for correlation values close to zero.
+The interactions between correlation and sign significantly reduce the time taken to pick the lineup for stronger correlations.
+Unsurprisingly, correct decisions are made faster.
+Surprisingly, neither smoothness nor sample size has an impact on the time for either plot design, despite having statistically significant impacts on power.
+An overlaid line graph having positive correlation increases the time required, eliminating the benefits from the almost significant plot design term.
+This result seems counter-intuitive as the weak performance of the overlaid lines in negative correlation implies that negative would require more time than positive, but the interactions between plot design, correlation and sign result in there being almost no decrease in time for strongly negatively correlated line plots.
+
 \section{Conclusion}
 
+This research showed that the scatterplot is better than the overlaid lines for displaying two time series when the purpose is to examine the association between the two series.
+This is especially true if correlation between the series is negative.
+This comparison was made possible by the use of the lineup protocol for comparing plot designs.
+The power of two plot designs for reading correlation between variables was modelled using a logistic regression incorporating subject’s individual visual ability as a random effect.
+We found that  the scatterplot, in most situations is both more powerful and quicker to process than  the overlaid line graph.
+The overlaid line graph appeared to have stronger performance at extremely low correlations, but this research did not test absolute values of correlation below 0.3.
+Other non-linear associations may exist in data that are not measured by correlation, and it may be of interest to compare plot power with a broader range of true plot features.
+Investigation of the individual selections of the lineup evaluations finds some results in-line with Wertheimer’s Gestalt Law of common fate, as demonstrated in Figure 15.
+The law suggested that people would find relationships between two lines that briefly move together, regardless of whether the relationship actually existed.
+Whilst this is shown, the overlaid line graph did not have a corresponding increase in confidence compared to the scatter plot; so research into similar confidence-boosting visual anomalies that can occur in scatter plots would be necessary to further investigate the law.
+Diaconis's Magical Thinking was supported by the finding on the relationships between time, confidence and correctness.
+A more difficult true plot may be picked rarely and require a longer to make a decision; resulting in decreased subject confidence.
+However, we found that time increases confidence but reduces correctness.
+Highly confident, but incorrect decisions, suggests subjects may have been distracted by strong, spurious patterns in the null plots.
+If a data analyst is required to use overlaid line graphs in their work, substantial improvements to power and hence readability can be made by smoothing time series data if the sample size is long enough for these techniques to become feasible.
+
 \end{document}

### 2015-12-09 paper/Paper_sweave.Rnw
fix author order

@@
 {
-\author{Dianne Cook\\
+\author{Nathaniel Tomasetti and Dianne Cook\\
 Department of Econometrics and Business Statistics, Monash University\\
-and\\
-Nathaniel Tomasetti\\
-Department of Econometrics and Business Statistics, Monash University\\}
+}
   \maketitle

### 2015-12-09 paper/Paper_sweave.Rnw
Updated

--- 2015-12-07 paper/Paper_sweave.Rnw
@@
 pin$base <- 0
-pin <- read.csv("pin.csv") #Do this part through R, not excel
+pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
 pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))

### 2015-12-09 paper/Paper_sweave.Rnw
Updated

--- 2015-12-09 paper/Paper_sweave.Rnw
@@
 colnames(counts) <- c("lineup", "plot", "count")
-Turka <- Turk[c(2, 10, 11, 13, 17, 20)]
-pin <- merge(counts, Turka, by.x="lineup", by.y="pic_id")
+pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
+pin <- pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
-head(pin)
 colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"), smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
-pin <- read.csv("pin.csv") #Do this part through R, not excel
+pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
 pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))

### 2015-12-10 paper/paper.Rnw
renamed Rnw, working on code

--- 2015-12-09 paper/Paper_sweave.Rnw
@@
 
-\SweaveOpts{concordance=TRUE}
-
 \title{\bf Comparing the Power of Plot Designs to Reveal Correlation}
@@
 \newpage
-<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 library(reshape2)
+# Convert reshape2 to tidyr, melt to gather
 library(nullabor)
@@
 library(splitstackshape)
+library(readr)
+library(lubridate)
+# Convert these to read_csv at some point
 correl <- read.csv("correlations.csv")
-full <- read.csv("full.csv") 
+full <- read.csv("full.csv", row.names=1) 
 Turk <- read.csv("turk.csv")
@@
 p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
-    theme(axis.title.x = element_blank(),
+    theme(aspect.ratio=1, axis.title.x = element_blank(),
           axis.title.y = element_blank(),
@@
 p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
-    theme(axis.title.x = element_blank(),
+    theme(aspect.ratio=1, axis.title.x = element_blank(),
           axis.title.y = element_blank(),
@@
 p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(axis.title.x = element_blank(),
+      theme(aspect.ratio=1, axis.title.x = element_blank(),
           axis.title.y = element_blank(),
@@
 p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(axis.title.x = element_blank(),
+      theme(aspect.ratio=1, axis.title.x = element_blank(),
           axis.title.y = element_blank(),
@@
 <<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
-ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")}
+ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
 @
@@
 sub <- filter(full, subj_id %in% keep2)
-sub <- sub[c(7, 15, 16, 17, 20, 27)]
-sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
-sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
-sub <- mutate(sub, prop = correct/Freq)
-ggplot(data=sub, aes(x=conf, y= prop)) + geom_point() + labs(title="Accuracy versus Confidence", y = "Accuracy", x = "Average Confidence")
+sub <- select(sub, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
+#sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
+#sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
+#sub <- mutate(sub, prop = correct/Freq)
+qplot(correlation, 6-conf_level, data=sub, alpha=I(0.1)) + 
+  geom_smooth(method="loess") + 
+  facet_grid(n~plot+smoothed) +
+  labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
 @
@@
 <<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
-correl <- read.csv("correlations.csv")
-pin <- full[c(5, 6, 10, 15, 16, 17)]
+# correl <- read.csv("correlations.csv") # You have already read this
+pin <- select(full, pic_id, response_0, actual_location, 
+              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
+# Need to use the tidyr commands instead of cSplit
+# Here's where we will need to make some changes to incorporate the multiple picks
 pin <- cSplit(pin, "response_0", direction="long")
@@
 colnames(counts) <- c("lineup", "plot", "count")
-Turka <- Turk[c(2, 10, 11, 13, 17, 20)]
-pin <- merge(counts, Turka, by.x="lineup", by.y="pic_id")
+pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
+pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
-head(pin)
-colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+#colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+pin <- mutate(pin, design=ifelse(test_param=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
@@
 The model is fitted using the lme4 package in
-<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+
+<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 
-m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "L"), family="binomial")
-m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=subset(full.sub, plot == "S"), family="binomial")
+m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial")
+m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial")
 @
@@
 The model is fitted using the lme4 package in R.
-<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+
+<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 

### 2015-12-10 paper/paper.Rnw
Figures

@@
 
-Figure 7 shows the raw results of the survey and Figure 8 shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
+Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
 Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
@@
 The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered.
-This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure 6).
+This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure~\ref{designs}).
 The right side shows the performance benefits of smoothing overlaid line time series data.
@@
 
-Figure 9 demonstrates the effects of changing sample size for unsmoothed data.
+Figure~\ref{samplesize} demonstrates the effects of changing sample size for unsmoothed data.
 The scatter plot has increased power to detect correlation almost universally for any increase.
 Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
-Figures 10 and 11 give a breakdown of the interactions between the response variables: plot detection, time and confidence.
+Figure~\ref{time} and Figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
 Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
@@
 Next we analysed the individual ability of each of our subjects.
-Figure 12 plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
+Figure~\ref{confidence2} plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
 While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two.
@@
 Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis).
-We next can analyse these kinds of errors with the pin plots found in Figure 13.
+We next can analyse these kinds of errors with the pin plots found in Figure~\ref{pins}.
 The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot.
@@
 
-Figures 14 and 15 show lineups where a null plot with a strong pattern led to its selection over the data plot.
+Figures~\ref{anom1} and \ref{anom2} show lineups where a null plot with a strong pattern led to its selection over the data plot.
 Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
-The true data plot is in position 4, but 9/11 subjects chose plot 6.
+In Figure~\ref{anom1}, the true data plot is in position 4, but 9/11 subjects chose plot 6.
 Plot 6 is unusual, most of the points are scattered around the vertical median, but there are two outlying points, one with a high y-value and one with a low y-value.
 Other concurrent work has suggested that outliers are pre-attentive patterns and this plot selection by the observers is consistent with that.
-Figure 15 below is another example of where subjects chose a null plot that had smaller correlation than the data plot.
+Figure~\ref{anom2} above is another example of where subjects chose a null plot that had smaller correlation than the data plot.
 Four out of five evaluations selected the null plot in position 13,  instead of the data plot in position 8.
@@
 The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure 5).
+Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure~\ref{smoothing}).
 It did have a significant positive impact on the power of overlaid line graphs.
 Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
-Figure 16 shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
+Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
 The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
@@
 We did not include these characteristics of the individual in the modelling due to this lack of significance.
-Figure 16 especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
+Figure~\ref{powermodel} especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
 

### 2015-12-10 paper/paper.Rnw
Figures

@@
 
-\graphicspath{{figure/}}
+\graphicspath{{images/}}
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
@@
 
-Insert Figure 1
+\begin{figure}[htbp]
+  \centering
+  \includegraphics{cny line}
+  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
+(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
+  \label{cny}
+\end{figure}
 
@@
 
-Insert Figure 2
+\begin{figure}[htbp]
+  \centering
+  \begin{subfigure}
+    \includegraphics{nzd scatter}
+  \end{subfigure}
+  \begin{subfigure}
+    \includegraphics{nzd line}
+  \end{subfigure}
+  \caption{ Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
+Which one can you read the correlation from more accurately?
+Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
+  \label{nzd}
+\end{figure}
 
 To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26].
-This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure 1), even if no relationship exists.
+This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists.
 Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18].
@@
 Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot.
-Figure 1 highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
+Figure~\ref{cny} highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
 However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
 Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm.
-Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure 3).
+Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
 When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small.
@@
 
-Insert Figure 3
+\begin{figure}[htbp]
+  \centering
+  \includegraphics{murders}
+  \caption{[19] Murders in New York State (Orange) and New York City (Blue).
+The difference between the two represents murders outside of the city.
+What size would you judge the difference to be in 1994?}
+  \label{murder}
+\end{figure}
 
@@
   \centering
-Insert anomaly lineup 1
+  \includegraphics{anom1}
   \label{anom1}
@@
   \centering
-Insert anomaly lineup 2
+  \includegraphics{anom2}
   \label{anom2}

### 2015-12-10 paper/paper.Rnw
Removed melt

@@
 library(mnormt)
-library(reshape2)
-# Convert reshape2 to tidyr, melt to gather
+library(tidyr)
+# Convert reshape2 to tidyr, melt to gather - done
 library(nullabor)
@@
 a <- gen_true_data(24, 0.95)
-td <- melt(a, id="t") 
+td <- gather(a, variable, value, c(X1, X2))
 b <- gen_null(24)
 b$.n <- NULL
-nd_l <- melt(b, id="t")
+nd_l <- gather(b, variable, value, c(X1, X2))
 nd_l$.n <- rep(1:19, rep(24, 19))
@@
 uns <- gen_true_data(96, 0.8)
-uns_l <- melt(uns, id="t")
+uns_l <- gather(uns, variable, value, -t)
 p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
@@
 sm <- gen_true_data(96, 0.8, smoothed = TRUE)
-sm_l <- melt(sm, id="t")
+sm_l <- gather(sm, variable, value, -t)
 p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
@@
 cb <- cor(b$X1, b$X2)
-a_l <- melt(a, id="t")
-b_l <- melt(b, id="t")
+a_l <- gather(a, variable, value, -t)
+b_l <- gather(b, variable, value, -t)
 p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +

### 2015-12-10 paper/paper.Rnw
Removed CSplit

@@
 # Convert reshape2 to tidyr, melt to gather - done
+#library(reshape2)
 library(nullabor)
@@
 library(gridExtra)
-library(splitstackshape)
+#library(splitstackshape) - changed cSplit to unnest
 library(readr)
@@
 # Here's where we will need to make some changes to incorporate the multiple picks
-pin <- cSplit(pin, "response_0", direction="long")
+pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
 counts <- data.frame(table(pin$pic_id, pin$response_0))

### 2015-12-10 paper/paper.Rnw
Fixed Pins

--- 2015-12-10 paper/paper.Rnw
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 \newpage
-<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 library(gridExtra)
-library(splitstackshape)
-library(readr)
-library(lubridate)
+#library(splitstackshape) - changed cSplit to unnest
+#library(readr)
+#library(lubridate)
 # Convert these to read_csv at some point
@@
 # Here's where we will need to make some changes to incorporate the multiple picks
-pin <- cSplit(pin, "response_0", direction="long")
+pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+pin$response_0 <- as.integer(pin$response_0)
 counts <- data.frame(table(pin$pic_id, pin$response_0))
@@
 pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
@@
 pin <- mutate(pin, percent = count/attempt)
-ggplot(data=subset(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed=="no" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
        geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed=="yes" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed=="no" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed=="yes" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
@@
 
-<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 
-<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 

### 2015-12-10 paper/paper.Rnw
Updates

@@
   \centering
-  \includegraphics{cny line}
+  \includegraphics{cnyline}
   \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
@@
   \begin{subfigure}
-    \includegraphics{nzd scatter}
+    \includegraphics{nzdscatter}
   \end{subfigure}
   \begin{subfigure}
-    \includegraphics{nzd line}
+    \includegraphics{nzdline}
   \end{subfigure}
-  \caption{ Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
+  \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
 Which one can you read the correlation from more accurately?
@@
   \centering
-<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+\iffalse
 # correl <- read.csv("correlations.csv") # You have already read this
@@
 #Code to be fixed
-@
+\fi
+%@
+
   \label{pins}
@@
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- full.sub[,c(1,7, 15,16,17,20:26)]
-full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
-full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
+full.sub <- select(full.sub, correct, r, rsgn, smoothed, n, subj_id, plot, log_time)
 full.sub$alpha <- log(.05/0.95)
@@
 <<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
-timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 
-                  + test_param*r + test_param*n + test_param*rsgn*r + test_param*correct + (1|subj_id), data=full.sub)
+timelmer <-  lmer(log_time ~ plot*smoothed + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
 @

### 2015-12-10 paper/paper.Rnw
got it to compile, almost its complaining about latex commands now, will clean this up on plane

--- 2015-12-10 paper/paper.Rnw
@@
 \usepackage{amsmath}
-\usepackage{dsfont}
+%\usepackage{dsfont}
 \usepackage{amsthm}
@@
 \usepackage{ulem}
-\usepackage[section]{placeins}
+%\usepackage[section]{placeins}
 \usepackage{afterpage}
 
-\graphicspath{{images/}}
+\graphicspath{{figure/}}
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
@@
 \newpage
+
 <<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+opts_chunk$set(cache=FALSE)
 library(lme4)
@@
 library(mnormt)
-library(tidyr)
-# Convert reshape2 to tidyr, melt to gather - done
-#library(reshape2)
+library(reshape2)
+# Convert reshape2 to tidyr, melt to gather
 library(nullabor)
@@
 library(gridExtra)
-#library(splitstackshape) - changed cSplit to unnest
+library(splitstackshape)
 library(readr)
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics{cny line}
-  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
-(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
-  \label{cny}
-\end{figure}
+Insert Figure 1
 
@@
 
-\begin{figure}[htbp]
-  \centering
-  \begin{subfigure}
-    \includegraphics{nzd scatter}
-  \end{subfigure}
-  \begin{subfigure}
-    \includegraphics{nzd line}
-  \end{subfigure}
-  \caption{ Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
-Which one can you read the correlation from more accurately?
-Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
-  \label{nzd}
-\end{figure}
+Insert Figure 2
 
 To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26].
-This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists.
+This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure 1), even if no relationship exists.
 Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18].
@@
 Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot.
-Figure~\ref{cny} highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
+Figure 1 highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
 However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
 Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm.
-Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
+Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure 3).
 When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small.
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics{murders}
-  \caption{[19] Murders in New York State (Orange) and New York City (Blue).
-The difference between the two represents murders outside of the city.
-What size would you judge the difference to be in 1994?}
-  \label{murder}
-\end{figure}
+Insert Figure 3
 
@@
 a <- gen_true_data(24, 0.95)
-td <- gather(a, variable, value, c(X1, X2))
+td <- melt(a, id="t") 
 b <- gen_null(24)
 b$.n <- NULL
-nd_l <- gather(b, variable, value, c(X1, X2))
+nd_l <- melt(b, id="t")
 nd_l$.n <- rep(1:19, rep(24, 19))
@@
 uns <- gen_true_data(96, 0.8)
-uns_l <- gather(uns, variable, value, -t)
+uns_l <- melt(uns, id="t")
 p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
@@
 sm <- gen_true_data(96, 0.8, smoothed = TRUE)
-sm_l <- gather(sm, variable, value, -t)
+sm_l <- melt(sm, id="t")
 p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
@@
 cb <- cor(b$X1, b$X2)
-a_l <- gather(a, variable, value, -t)
-b_l <- gather(b, variable, value, -t)
+a_l <- melt(a, id="t")
+b_l <- melt(b, id="t")
 p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
@@
 
-Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
+Figure 7 shows the raw results of the survey and Figure 8 shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
 Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
@@
 The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered.
-This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure~\ref{designs}).
+This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure 6).
 The right side shows the performance benefits of smoothing overlaid line time series data.
@@
 
-Figure~\ref{samplesize} demonstrates the effects of changing sample size for unsmoothed data.
+Figure 9 demonstrates the effects of changing sample size for unsmoothed data.
 The scatter plot has increased power to detect correlation almost universally for any increase.
 Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
-Figure~\ref{time} and Figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
+Figures 10 and 11 give a breakdown of the interactions between the response variables: plot detection, time and confidence.
 Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
@@
 Next we analysed the individual ability of each of our subjects.
-Figure~\ref{confidence2} plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
+Figure 12 plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
 While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two.
@@
 Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis).
-We next can analyse these kinds of errors with the pin plots found in Figure~\ref{pins}.
+We next can analyse these kinds of errors with the pin plots found in Figure 13.
 The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot.
@@
 # Here's where we will need to make some changes to incorporate the multiple picks
-pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+pin <- cSplit(pin, "response_0", direction="long")
 counts <- data.frame(table(pin$pic_id, pin$response_0))
@@
 pin$r <- factor(pin$r)
-#colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
-pin <- mutate(pin, design=ifelse(test_param=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
+colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
@@
   \centering
-  \includegraphics{anom1}
+Insert anomaly lineup 1
   \label{anom1}
@@
   \centering
-  \includegraphics{anom2}
+Insert anomaly lineup 2
   \label{anom2}
@@
 
-Figures~\ref{anom1} and \ref{anom2} show lineups where a null plot with a strong pattern led to its selection over the data plot.
+Figures 14 and 15 show lineups where a null plot with a strong pattern led to its selection over the data plot.
 Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
-In Figure~\ref{anom1}, the true data plot is in position 4, but 9/11 subjects chose plot 6.
+The true data plot is in position 4, but 9/11 subjects chose plot 6.
 Plot 6 is unusual, most of the points are scattered around the vertical median, but there are two outlying points, one with a high y-value and one with a low y-value.
 Other concurrent work has suggested that outliers are pre-attentive patterns and this plot selection by the observers is consistent with that.
-Figure~\ref{anom2} above is another example of where subjects chose a null plot that had smaller correlation than the data plot.
+Figure 15 below is another example of where subjects chose a null plot that had smaller correlation than the data plot.
 Four out of five evaluations selected the null plot in position 13,  instead of the data plot in position 8.
@@
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- full.sub[,c(1,7, 15,16,17,20:26)]
-full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
-full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
+full.sub <- select(full.sub, subj_id, correct, r, rsgn, smoothed, n, plot, log_time)#[,c(1,7, 15,16,17,20:26)]
+#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
+#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
 full.sub$alpha <- log(.05/0.95)
@@
 The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure~\ref{smoothing}).
+Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure 5).
 It did have a significant positive impact on the power of overlaid line graphs.
 Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
-Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
+Figure 16 shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
 The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
@@
 We did not include these characteristics of the individual in the modelling due to this lack of significance.
-Figure~\ref{powermodel} especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
+Figure 16 especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
 
@@
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
-timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + test_param + test_param*smoothed 
-                  + test_param*r + test_param*n + test_param*rsgn*r + test_param*correct + (1|subj_id), data=full.sub)
+timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
+                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
 @

### 2015-12-10 paper/paper.Rnw

--- 2015-12-10 paper/paper.Rnw
@@
 \usepackage{ulem}
-\usepackage[section]{placeins}
+%\usepackage[section]{placeins}
 \usepackage{afterpage}
@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
 \newpage
-<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+
+<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 
-<<powerfit, dependson='setup', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- select(full.sub, correct, r, rsgn, smoothed, n, subj_id, plot, log_time)
+full.sub <- select(full.sub, subj_id, correct, r, rsgn, smoothed, n, plot, log_time)#[,c(1,7, 15,16,17,20:26)]
+#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
+#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
 full.sub$alpha <- log(.05/0.95)
@@
 
-<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
-timelmer <-  lmer(log_time ~ plot*smoothed + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
+<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
+                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
 @

### 2015-12-10 paper/paper.Rnw
getting closer to fixing latex errors
tweaking pin plot code

@@
 \usepackage{amsmath}
-%\usepackage{dsfont}
+\usepackage{dsfont}
 \usepackage{amsthm}
@@
 \usepackage{ulem}
-%\usepackage[section]{placeins}
+\usepackage[section]{placeins}
 \usepackage{afterpage}
@@
 \DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
-
-\newcommand{\blue}[1]{{\color{blue} #1}}
-\newcommand{\hh}[1]{{\color{magenta} #1}}
-\newcommand{\dc}[1]{{\color{orange} #1}}
-\newcommand{\green}[1]{{\color{cyan} #1}}
 
@@
 # Convert these to read_csv at some point
-correl <- read.csv("correlations.csv")
-full <- read.csv("full.csv", row.names=1) 
-Turk <- read.csv("turk.csv")
+correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
+full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
+Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)
 
@@
   \centering
-  \includegraphics{cnyline}
+  \includegraphics[width=0.5\textwidth]{cnyline}
   \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
@@
   \centering
-  \begin{subfigure}
-    \includegraphics{nzdscatter}
-  \end{subfigure}
-  \begin{subfigure}
-    \includegraphics{nzdline}
-  \end{subfigure}
+    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
+    \includegraphics[width=0.45\textwidth]{nzdline}}
   \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
@@
   \centering
-  \includegraphics{murders}
+  \includegraphics[width=0.5\textwidth]{murders}
   \caption{[19] Murders in New York State (Orange) and New York City (Blue).
@@
   \centering
-<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
 set.seed(1234)
@@
   \centering
-<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
 set.seed(123)
@@
   \centering
-<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
 set.seed(101)
@@
   \centering
-%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
-\iffalse
+<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+#\iffalse
 # correl <- read.csv("correlations.csv") # You have already read this
@@
 # Here's where we will need to make some changes to incorporate the multiple picks
-pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
-pin$response_0 <- as.integer(pin$response_0)
-counts <- data.frame(table(pin$pic_id, pin$response_0))
+#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
+pin$choice <- separate(pin, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
+pin$choice[str_detect(pin$actual_location, pin$response_0)] <- 
+  pin$actual_location[str_detect(pin$actual_location, pin$response_0)]
+                         #as.integer(pin$response_0)
+counts <- data.frame(table(pin$pic_id, pin$choice))
 counts$Var1 <- as.integer(counts$Var1)
@@
 pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
 colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
-pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
+pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines")) #, smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
@@
 pin <- mutate(pin, percent = count/attempt)
-ggplot(data=subset(pin, smoothed=="no" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
        geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed=="yes" & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed=="no" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed=="yes" & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-
 #Code to be fixed
-\fi
-%@
+#\fi
+@
 
@@
   \centering
-  \includegraphics{anom1}
+  \includegraphics[width=0.5\textwidth]{anom1}
   \label{anom1}
@@
   \centering
-  \includegraphics{anom2}
+  \includegraphics[width=0.5\textwidth]{anom2}
   \label{anom2}

### 2015-12-10 paper/paper.Rnw
its working, with references

@@
 library(gridExtra)
+library(stringr)
 #library(splitstackshape) - changed cSplit to unnest
@@
 Scatter plots generally showed superior performance, with the overlaid line graphs being particularly weak at revealing negative correlation}
+\end{figure}
 
@@
 Values above zero indicate the scatter plot is more powerful than the overlaid line graph.}
-\end{figure
+\end{figure}
 
@@
 
+Here's how to do references \citet{hadley:2009} and \citep{hadley:2009}.
+
+\bibliographystyle{asa}
+\bibliography{references}
+
 \end{document}

### 2015-12-11 paper/paper.Rnw
Tweaked Pins

@@
 counts <- data.frame(table(pin$pic_id, pin$choice))
-counts$Var1 <- as.integer(counts$Var1)
-counts$Var2 <- as.integer(counts$Var2)
+counts$Var1 <- as.numeric(as.character(counts$Var1))
+counts$Var2 <- as.numeric(as.character(counts$Var2))
 colnames(counts) <- c("lineup", "plot", "count")
 pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
 colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
-pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines")) #, smoothed=ifelse(smoothed=="no", 0, 1))
+pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"), smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0

### 2015-12-11 paper/paper.Rnw
Bibliography

--- 2015-12-10 paper/paper.Rnw
@@
 \begin{abstract}
-Visual inference in EDA is prone to type 1 errors from the over-interpretation of randomness [4, 6].
+Visual inference in EDA is prone to type 1 errors from the over-interpretation of randomness \citep{Daniel, Diaconis}.
 Two competing plot designs, the scatter plot and overlaid line graph are both popular in the analysis of time series data.
-Lineups [2, 10, 11, 16, 27] allow the visual inference power of a graphic display to be evaluated, and were used to compare the plot designs.
-We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the decision for 2088 lineup evaluations.
+Lineups \citep{Majumder:2013, Wickham} allow the visual inference power of a graphic display to be evaluated, and were used to compare the plot designs.
+We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the decision for 2089 lineup evaluations.
 The results show that the scatter plot is both the faster and more powerful plot design, despite its inability to display the time dimension.
@@
 
-<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 This idea was formed as early as the 18th century, when William Playfair institutionalized the then revolutionary idea of graphing government and economic data.
-Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information [20].
-Since then, advances in computing power have allowed statistical graphics to flourish, spearheaded by Tukey in 1965 into the new domain of exploratory data analysis (EDA) [7, 23].
-EDA can be thought of as a well understood [3, 25] set of tools and techniques required to visualise information, to physically see what the data contains.
+Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information \citep{Sachs}.
+Since then, advances in computing power have allowed statistical graphics to flourish, spearheaded by \citet{Tukey} in 1965 into the new domain of exploratory data analysis (EDA) \citep{Friendly}.
+EDA can be thought of as a well understood \citep{Cleveland, Vanderplas} set of tools and techniques required to visualise information, to physically see what the data contains.
 In particular, it incorporates a free roaming approach, where the analyst is able to explore structure to find whatever relationships and structure exists within.
@@
 Both EDA and MD follow the same framework: Visualise the data, look for patterns that suggest an underlying relationship, and if one is found implement it into the model then continue the exploration of the data.
-It has long been thought that EDA and statistical inference were worlds apart, but recent work [2, 16] bridges this chasm.
+It has long been thought that EDA and statistical inference were worlds apart, but recent work in \citet{Buja} and \citet{Majumder:2013} bridges this chasm.
 Framing a data plot as a test statistic, which when compared to plots of null data, places EDA into the statistical inference recipe.
@@
 Visual skills of an observer in EDA is prone to Type I error, where the null hypothesis is rejected when it is actually true, caused by the inherently random formation of patterns when visualised which are attributed to structure rather than chance.
-Daniel [4] warns against this, by providing 40 pages of plots from the null distribution, he encourages data analysts to understand the patterns that can be created from data without inherent structure.
-(This is what Buja et al [2] call the Rorschach protocol.) By being aware of what can appear in this type of data, the analyst should be more wary of claiming any visual feature they find is a true relationship.
+\citet{Daniel} warns against this, by providing 40 pages of plots from the null distribution, he encourages data analysts to understand the patterns that can be created from data without inherent structure.
+(This is what \citet{Buja} call the Rorschach protocol.) By being aware of what can appear in this type of data, the analyst should be more wary of claiming any visual feature they find is a true relationship.
 But this is not enough, even seasoned data analysts can be misled.
-Diaconis [6] introduced the notion of 'magical thinking' which argues that people commonly suffer from the over-interpretation of randomness, particularly if it matches a pre-conception.
+\citet{Diaconis} introduced the notion of 'magical thinking' which argues that people commonly suffer from the over-interpretation of randomness, particularly if it matches a pre-conception.
 If it suits their particular bias, an analyst may make false discoveries of structure and false rejections of the null hypothesis.
@@
 In this research we utilise the methods in visual inference to study plot design to investigate the relationship between two temporal variables.
-There is a substantial literature suggesting that scatter plots are the appropriate display to read association [3, 8, 14], however, when the two variables are temporal it is common to display them as time series, drawn on the same plot.
-Temporal dependence is present in many macroeconomic variables [16, 17] and there is a strong argument for its presence in financial data [21, 24].
+There is a substantial literature suggesting that scatter plots are the appropriate display to read association \citep{Cleveland, Harrison, Li}, however, when the two variables are temporal it is common to display them as time series, drawn on the same plot.
+Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and there is a strong argument for its presence in financial data \citep{Sher, Turtle}.
 It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data.
@@
 
-To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by Wertheimer in 1923 [22, 26].
+To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by \citet{Wertheimer}.
 This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists.
-Lee and Blake's [13] work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot [18].
-However, keep in mind magical thinking [6], the desire to find any pattern may lead to a false discovery of positive correlation.
+\citet{Lee} work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot.
+However, keep in mind magical thinking \citep{Diaconis}, the desire to find any pattern may lead to a false discovery of positive correlation.
 The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier.
@@
 Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm.
-Work by Robbins [19] is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
+\citet{Robbins} is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
 When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small.
@@
 It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993.
-Vanderplas [25] examines this effect in more detail.
+\citet[Vanderplas] examines this effect in more detail.
 It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope.
@@
   \includegraphics[width=0.5\textwidth]{murders}
-  \caption{[19] Murders in New York State (Orange) and New York City (Blue).
+  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
 The difference between the two represents murders outside of the city.
@@
 
-There has been other research on the perception of  temporal displays, e.g.
-Javed et al.
-[12], but they do not examine perception of association.
+There has been other research on the perception of  temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
 Javed et al.
@@
 We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within.
-This can be found by using the lineup protocol developed in 2009 by Buja et al.
-[2], which is easily implemented with the nullabor R package.
-They created a statistically rigorous framework with properties explored primarily in Majumder et.
-al [16] and further in Hofmann et.
-al [11] that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
-The lineup protocol is inspired in part by the police lineup [27].
+This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
+They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hoffman:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
+The lineup protocol is inspired in part by the police lineup \citep{Wickham}.
 We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people).
@@
 $p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup.
-It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ [11].
+It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hoffman:2014}.
 If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution.
@@
 We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed [2, 11] with:
+A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hoffman:2014} with:
 
@@
 We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ correctly detected the true plot out of the lineup.
-We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as [10]:
+We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as \citep{Hoffman:2012}:
 
@@
 Some individuals may have a better natural ability at detecting the correct correlation pattern in the graphics, but by having each participant viewing multiple lineups and each lineup being viewed by multiple different participants, this can be controlled via a random effect variable in the model.
-We recruited participants through Amazon's Mechanical Turk, where the lineup protocol has been applied to a number of problems in prior papers [9, 16, 27].
-Amazon’s Mechanical Turk [1] (MTurk) is a labour crowd-sourcing platform developed to give easy access to workers with basic tasks paid in line with the United States minimum wage.
+We recruited participants through Amazon's Mechanical Turk, where the lineup protocol has been applied to a number of problems in prior papers \citep{Heer, Majumder:2013, Wickham}.
+Amazon’s Mechanical Turk \citep{MTurk} is a labour crowd-sourcing platform developed to give easy access to workers with basic tasks paid in line with the United States minimum wage.
 MTurk can be used to recruit subjects to read a variety of graphics and report on particular visual tasks.
 The time taken to decide, confidence in the decision and the reason for that decision can also be recorded.
-Heer and Bostock [9] use MTurk to replicate previous findings in graphical perception [3] and find that it is a valid method of data collection, once controls to eliminate anyone 'gaming' the system by randomly selecting answers to minimise time spent working are implemented.
+\citet{Heer} use MTurk to replicate previous findings in graphical perception from \citet{Cleveland} and find that it is a valid method of data collection, once controls to eliminate anyone 'gaming' the system by randomly selecting answers to minimise time spent working are implemented.
 Further details of our use of Amazons’ Mechanical Turk is included in Section 3.
-The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in Hofmann et al [10], and this work follows that approach.
+The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in \citet{Hoffman:2012}, and this work follows that approach.
 
@@
 If it was answered correctly, the trial data was discarded and the subject moved onto the rest of the lineups.
-Earlier work [15] has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
+Earlier work by \citet{Majumder:2014} has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
 We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
@@
   \centering
-<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
-#\iffalse
+%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+\iffalse
 # correl <- read.csv("correlations.csv") # You have already read this
@@
 #Code to be fixed
-#\fi
-@
+\fi
+%@
 
@@
 The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
-Previous studies [15] show that this random visual skill is not impacted by demographics such as age and gender; whilst education increases the power of a graphic display by only approximately one percent, not a practically significant amount.
+\citet{Majumder:2014} show that this random visual skill is not impacted by demographics such as age and gender; whilst education increases the power of a graphic display by only approximately one percent, not a practically significant amount.
 We did not include these characteristics of the individual in the modelling due to this lack of significance.
@@
 
-<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
 timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
@@
 
+\bibliographystyle{asa}
+\bibliography{references}
+
 \end{document}

### 2015-12-15 paper/paper.Rnw
few code fixes

@@
 
-<<setup, echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 library(lme4)
@@
 library(stringr)
+library(pander)
 #library(splitstackshape) - changed cSplit to unnest
@@
   \centering
-%<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
-\iffalse
+<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
 # correl <- read.csv("correlations.csv") # You have already read this
@@
 pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
 colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
-pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"), smoothed=ifelse(smoothed=="no", 0, 1))
+pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
@@
 #Code to be fixed
-\fi
-%@
+@
 
@@
 full.sub$rabs <- abs(full.sub$r)
-full.sub$rabs <- factor(full.sub$rabs)
+#full.sub$rabs <- factor(full.sub$rabs)
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- select(full.sub, subj_id, correct, r, rsgn, smoothed, n, plot, log_time)#[,c(1,7, 15,16,17,20:26)]
+full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
+full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
+#full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
+#full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+#full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
 #full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
@@
 full.sub$alpha <- log(.05/0.95)
+full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
 
-m8l <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial")
-m8s <- glmer(correct~offset(alpha) + r + rsgn + r*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial")
+m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
+m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
+pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
+pander(summary(m8s), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for scatter plots"))
 @
-%Code to make a table to be added
 
@@
 full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
-full.pred2 <- rbind(full.pred8s, full.pred8l)
+full.pred <- rbind(full.pred8s, full.pred8l)
 
-full.pred2 <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))
+#full.pred <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))
 
-qplot(r, fitre, data=full.pred2, color=plot, 
+qplot(r, fitre, data=full.pred, color=plot, 
       group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
   scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
-  xlab("Correlation") + ylab("Percent Detected") + labs(title="Predicted Power") +
+  xlab("Correlation") + ylab("Estimated Power") + labs(title="Predicted Power") +
   geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
@@
 
-<<timefit, dependson='powerfit', echo=FALSE, results=hide, message=FALSE, error=FALSE, warning=FALSE>>=
+<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                   + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
+pander(summary(timelmer), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for time taken"))
 @
-%get table code
 

### 2015-12-15 paper/paper.Rnw
few code fixes

@@
 \newpage
+
+\tableofcontents
 
@@
 
-Table 1: Mixed Effects Logit Model for the predicted power of plot design.
+\begin{table}[htp]
+\caption{Mixed Effects Logit Model for the predicted power of plot design.
 Sign = 1 if correlation > 0, 0 otherwise.
@@
 Sample size for scatter plot model: 841, overlaid lines model: 843.
-The model is fitted using the lme4 package in
-
-<<powerfit, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+The model is fitted using the lme4 package in R.}
+\label{tab1}
+<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 
-m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
+#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
 m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
-pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
-pander(summary(m8s), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for scatter plots"))
+#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
+#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
+#library(xtable)
+kable(summary(m8s)$coefficients, format="latex")
 @
+\end{table}
 
-Table 1 summarises the model fits.
+Table \ref{tab1} summarises the model fits.
 It can be confusing the interpret the model estimates, and at first glance it appears the overlaid lines (-0.660) are substantially more powerful than the scatter plot (-4.403), but this is misleading, and  interpreting the the full model involves combining the estimates for the different interactions with correlation and sign.

### 2015-12-15 paper/paper.Rnw
Cleanup

@@
 library(stringr)
-library(pander)
+#library(pander)
 #library(splitstackshape) - changed cSplit to unnest
@@
 b <- gen_null(24)
-b$.n <- NULL
 nd_l <- gather(b, variable, value, c(X1, X2))
-nd_l$.n <- rep(1:19, rep(24, 19))
 pos <- 7
@@
 
-plot <- ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
+ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
     theme_bw() + theme(axis.title.x = element_blank(),
@@
                        legend.position="none")
-print(plot)
 @
@@
 b <- gen_true_data(24, -0.95)
-ca <- cor(a$X1, a$X2)
-cb <- cor(b$X1, b$X2)
 a_l <- gather(a, variable, value, -t)
 b_l <- gather(b, variable, value, -t)
-p1 <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
+p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
       theme(axis.title.x = element_blank(),
@@
           legend.position="none")
-p2 <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
+p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
       theme(axis.title.x = element_blank(),
@@
           legend.position="none")
-p3 <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
+p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
       theme(aspect.ratio=1, axis.title.x = element_blank(),
@@
           legend.position="none")
-p4 <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
+p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
       theme(aspect.ratio=1, axis.title.x = element_blank(),
@@
           legend.position="none")
-grid.arrange(p1, p2, p3, p4, ncol=2)
+grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
 @
@@
 \section{Results}
+
+\subsection{Detection, Time and Confidence}
 
@@
 Turk$r <- round(Turk$correlation, 1)
-Turk.line <- subset(Turk, test_param=="line")
-Turk.scatter <- subset(Turk, test_param=="scatter")
-Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c(10, 13, 16, 19, 20)]
-Turk.line <- summarise(group_by(Turk.line, smoothed, n, r), sum(attempt), sum(correct))
-Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c(10, 13, 16, 19, 20)]
-Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
-colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
-colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
-prop <- cbind(Turk.line, Turk.scatter[4:5])
+#Turk.line <- subset(Turk, test_param=="line")
+#Turk.scatter <- subset(Turk, test_param=="scatter")
+prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "plot")]
+prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(plot=="Overlaid Lines", attempt, 0)),
+                                                  l.correct=sum(ifelse(plot=="Overlaid Lines", correct, 0)),
+                                                  s.attempt=sum(ifelse(plot=="Scatterplot", attempt, 0)),
+                                                  s.correct=sum(ifelse(plot=="Scatterplot", correct, 0)))
+#Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
+#Turk.line <- summarise(group_by(Turk.line, smoothed, n, r, test_param), sum(attempt), sum(correct))
+#Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
+#Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
+#colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
+#colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
+#prop <- cbind(Turk.line, Turk.scatter[4:5])
 prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
@@
                lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
-ggplot(data=prop) + geom_line(aes(x=r, y=difference), colour="#99CC99", size=2) + 
-                    geom_ribbon(aes(x=r, ymin=lower, ymax=upper), colour="#99CC99", fill="#99CC99", alpha=0.2) + 
-                    labs(title="Difference in Power", x="Correlation") + facet_grid(n~smoothed)
+ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
+        geom_hline(aes(yint=0, colour="red")) + labs(title="Difference in Power", x="Correlation", y="Scatterplot Advantage") + facet_grid(n~smoothed)
 @
@@
 Figure~\ref{time} and Figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
-Figure 8 is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
+Figure~\ref{time} is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
 The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed.
@@
 
+\subsection{Selection Analysis}
+
 Next we analysed the individual ability of each of our subjects.
@@
   \centering
-<<confidence2, dependson='confidence', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 # correl <- read.csv("correlations.csv") # You have already read this
-pin <- select(full, pic_id, response_0, actual_location, 
+count.setup <- select(full, pic_id, response_0, actual_location, 
               test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
@@
 #pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
-pin$choice <- separate(pin, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
-pin$choice[str_detect(pin$actual_location, pin$response_0)] <- 
-  pin$actual_location[str_detect(pin$actual_location, pin$response_0)]
+count.setup$choice <- separate(count.setup, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
+count.setup$choice[str_detect(count.setup$actual_location, count.setup$response_0)] <- 
+  count.setup$actual_location[str_detect(count.setup$actual_location, count.setup$response_0)]
                          #as.integer(pin$response_0)
-counts <- data.frame(table(pin$pic_id, pin$choice))
+counts <- data.frame(table(count.setup$pic_id, count.setup$choice))
 counts$Var1 <- as.numeric(as.character(counts$Var1))
@@
 pin <- mutate(pin, percent = count/attempt)
-ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
        geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
+pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
   geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-#Code to be fixed
+grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
 @
@@
 
-\section{Modelling Subject Responses}
+\subsection{Modelling Subject Responses}
 
@@
 
-$$Power_{i} = \psi\left\{ \beta_{i,0} + \beta_{i,1} \cdot Correlation + \beta_{i,2}\cdot Sign + \beta_{i,3}\cdot Smoothed + \beta_{i,4}\cdot t + \beta_{i,5}\cdot Correlation \cdot Sign + \gamma_{i,1} \cdot subject_{j} \right\}$$
+$$Power_{i} = \psi\left\{ \beta_{i,0} + \beta_{i,1} \cdot |\alpha| + \beta_{i,2}\cdot Sign + \beta_{i,3}\cdot Smoothed + \beta_{i,4}\cdot t + \beta_{i,5}\cdot |\alpha| \cdot Sign + \gamma_{i,1} \cdot subject_{j} \right\}$$
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $r$ is the true plot correlation, Sign is the direction of the true plot correlation, $i$  is the plot design and $j$ is an observer specific random effect.
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $\alpha$ is the true plot correlation, Sign is the direction of $\alpha$, $i$  is the plot design and $j$ is an observer specific random effect.
 
@@
 \label{tab1}
+\centering
 <<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
@@
 
-#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
+m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
 m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
@@
 #library(xtable)
+kable(summary(m8l)$coefficients, format="latex")
 kable(summary(m8s)$coefficients, format="latex")
@@
 
-$$Power_{UR,s} = \psi\left\{ \beta_{s,0} + \beta_{s,1} \cdot |Correlation| + \beta_{s,2}\cdot Sign + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t + \beta_{s,5}\cdot |Correlation| \cdot Sign + \gamma_{s,1} \cdot subject_{j} \right\}$$
+$$Power_{UR,s} = \psi\left\{ \beta_{s,0} + \beta_{s,1} \cdot |\alpha| + \beta_{s,2}\cdot Sign + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t + \beta_{s,5}\cdot |\alpha| \cdot Sign + \gamma_{s,1} \cdot subject_{j} \right\}$$
 
@@
 The restricted model is therefore:
-$$Power_{R,s} = \psi\left\{ \beta_{s,0}+\beta_{s,1}\cdot |Correlation| + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t +\gamma_{s,1} \cdot subject_{j} \right\}$$
+$$Power_{R,s} = \psi\left\{ \beta_{s,0}+\beta_{s,1}\cdot |\alpha| + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t +\gamma_{s,1} \cdot subject_{j} \right\}$$
 
@@
                   + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
-pander(summary(timelmer), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for time taken"))
+xable(summary(timelmer)$coefficients, format="latex")
 @

### 2015-12-16 paper/paper.Rnw
few fixes

@@
 This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
-They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hoffman:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
+They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
 The lineup protocol is inspired in part by the police lineup \citep{Wickham}.
@@
 $p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup.
-It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hoffman:2014}.
+It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hofmann:2014}.
 If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution.
@@
 #library(xtable)
-kable(summary(m8l)$coefficients, format="latex")
-kable(summary(m8s)$coefficients, format="latex")
+x <- summary(m8l)$coefficients
+y <- summary(m8s)$coefficients
+colnames(y)[4] <- "p-value"
+rownames(y) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
+kable(x, format="latex", digits=2)
+kable(y, format="latex", digits=2)
 @
@@
                   + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
-xable(summary(timelmer)$coefficients, format="latex")
+kable(summary(timelmer)$coefficients, format="latex")
 @

### 2015-12-16 paper/paper.Rnw
Updated

@@
 library(tidyr)
+library(knitr)
 # Convert reshape2 to tidyr, melt to gather - done
@@
   \centering
-<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
+<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
 # correl <- read.csv("correlations.csv") # You have already read this
 count.setup <- select(full, pic_id, response_0, actual_location, 
-              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
+              test_param, smoothed, n)
+count.setup$response_0 <- as.character(count.setup$response_0) #full[,c(5, 6, 10, 15, 16, 17)]
 # Need to use the tidyr commands instead of cSplit
@@
 colnames(counts) <- c("lineup", "plot", "count")
-pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- merge(counts, Turk[-19], by.x="lineup", by.y="pic_id")
+pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
@@
 pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
-pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
-pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
-pin <- mutate(pin, percent = count/attempt)
+pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"), percent = count/attempt)
+#pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
 pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
@@
 full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
-#full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
-#full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
-#full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
+full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
+full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
 #full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
@@
 full.sub$alpha <- log(.05/0.95)
-full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
-
+#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
+#weirw.csv(full.sub, "stata.csv") #exporting data to other programs
 m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
@@
 #library(xtable)
-x <- summary(m8l)$coefficients
-y <- summary(m8s)$coefficients
-colnames(y)[4] <- "p-value"
-rownames(y) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
-kable(x, format="latex", digits=2)
-kable(y, format="latex", digits=2)
+line.fit <- summary(m8l)$coefficients
+scatter.fit <- summary(m8s)$coefficients
+colnames(scatter.fit)[4] <- "p-value"
+rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
+#colnames(line.fit)[4] <- "p-value"
+#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
+kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
+kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
 @
@@
 n <- c(12, 24, 48, 96)
-full.pred8l <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
+predl <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                       smoothed=c(0,1), 
                                       subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
-full.pred8l$rsgn <- ifelse(full.pred8l$r > 0, 1, 0)
-full.pred8l$rabs <- abs(full.pred8l$r)
-full.pred8l$fitre <- predict(m8l, full.pred8l, type="response", re.form=~(1|subj_id),
+predl$rsgn <- ifelse(predl$r > 0, 1, 0)
+predl$rabs <- abs(predl$r)
+predl$fitre <- predict(m8l, predl, type="response", re.form=~(1|subj_id),
                              allow.new.levels=TRUE)
-full.pred8l$fitfx <- predict(m8l, full.pred8l, type="response", re.form=NA,
+predl$fitfx <- predict(m8l, predl, type="response", re.form=NA,
                              allow.new.levels=TRUE)
-full.pred8l$plot <- ifelse(full.pred8l$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
-full.pred8l <- filter(full.pred8l, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
+predl$plot <- ifelse(predl$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
+predl <- filter(predl, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
 
-full.pred8s <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
+preds <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                       smoothed=c(0,1), 
                                       subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
-full.pred8s$rsgn <- ifelse(full.pred8s$r > 0, 1, 0)
-full.pred8s$rabs <- abs(full.pred8s$r)
-full.pred8s$fitre <- predict(m8s, full.pred8s, type="response", re.form=~(1|subj_id),
+preds$rsgn <- ifelse(preds$r > 0, 1, 0)
+preds$rabs <- abs(preds$r)
+preds$fitre <- predict(m8s, preds, type="response", re.form=~(1|subj_id),
                              allow.new.levels=TRUE)
-full.pred8s$fitfx <- predict(m8s, full.pred8s, type="response", re.form=NA,
+preds$fitfx <- predict(m8s, preds, type="response", re.form=NA,
                              allow.new.levels=TRUE)
-full.pred8s$plot <- ifelse(full.pred8s$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
-full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
-full.pred <- rbind(full.pred8s, full.pred8l)
+preds$plot <- ifelse(preds$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
+preds <- filter(preds, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
+full.pred <- rbind(preds, predl)
 
@@
   \centering
-<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
 resids <- resid(m8s)
@@
 
+\begin{table}
+\label{tab2}
+\centering
+\caption{Mixed Linear Model fit for time taken to make a selection}
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
@@
                   + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
-kable(summary(timelmer)$coefficients, format="latex")
+time.fit <- summary(timelmer)$coefficients
+pv <- function(x){
+  p <- 2*(1-pt(x, 1835))
+}
+p.value <- sapply(time.fit[,3], pv)
+time.fit <- cbind(time.fit, p.value)
+time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
+kable(time.fit[1:7,], format="latex", caption="Main Effects")
+kable(time.fit[8:14,], format="latex", caption="Interaction Effects")
 @
+\end{table}
 

### 2015-12-16 paper/paper.Rnw
Updates

@@
   \item Factor 1: Plot design, levels = scatter plot, overlaid time series
-  \item Factor 2: Sample size, levels = 12, 24, 48, 96
+  \item Factor 2: Sample size (t), levels = 12, 24, 48, 96
   \item Factor 3: True plot correlation, levels 
@@
 colnames(scatter.fit)[4] <- "p-value"
-rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
+rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
 #colnames(line.fit)[4] <- "p-value"
@@
 time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
-kable(time.fit[1:7,], format="latex", caption="Main Effects")
-kable(time.fit[8:14,], format="latex", caption="Interaction Effects")
+rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
+kable(time.fit[1:7,], format="latex", caption="Main Effects (Overlaid Lines)")
+kable(time.fit[8:14,], format="latex", caption="Interaction Effects (Scatter plot)")
 @

### 2015-12-17 paper/paper.Rnw
Updates

@@
 
-In this research, we had three basic factors:
+The lineups were generated with three basic factors:
 
@@
 
-Additionally, there was a fourth factor for  = 48, 96:
+Additionally, there was a fourth factor for $t = 48, 96$:
 
@@
 
-$$Power_{i} = \psi\left\{ \beta_{i,0} + \beta_{i,1} \cdot |\alpha| + \beta_{i,2}\cdot Sign + \beta_{i,3}\cdot Smoothed + \beta_{i,4}\cdot t + \beta_{i,5}\cdot |\alpha| \cdot Sign + \gamma_{i,1} \cdot subject_{j} \right\}$$
+$$Power_{ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma_{1} \cdot z_{1,s} \right\}$$
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $\alpha$ is the true plot correlation, Sign is the direction of $\alpha$, $i$  is the plot design and $j$ is an observer specific random effect.
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$ $x_{2,j} is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
 
 \begin{table}[htp]
-\caption{Mixed Effects Logit Model for the predicted power of plot design.
-Sign = 1 if correlation > 0, 0 otherwise.
+\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design.
+Sample size for scatter plot model: 900, overlaid lines model: 949, total: 1849.
+Default plot design is an overlaid lines type.
+The model is fitted using the glimmix procedure in SAS.
+Subjects with less than ten lineup evaluations were removed from the sample.
+%Sign = 1 if correlation > 0, 0 otherwise.
 Each plot design is estimated separately.
 Results must be transformed via the logit link function to generate a power prediction.
-Subjects with less than ten lineup evaluations were removed from the sample.
-Sample size for scatter plot model: 841, overlaid lines model: 843.
-The model is fitted using the lme4 package in R.}
+Subjects with less than ten lineup evaluations were removed from the sample..}
 \label{tab1}
@@
 #full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
-#weirw.csv(full.sub, "stata.csv") #exporting data to other programs
-m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
-m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
+#weirw.csv(full.sub, "choices.csv") #exporting data to be estimated in SAS.
+#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
+#m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
 #pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
@@
 #library(xtable)
-line.fit <- summary(m8l)$coefficients
-scatter.fit <- summary(m8s)$coefficients
-colnames(scatter.fit)[4] <- "p-value"
-rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
+#line.fit <- summary(m8l)$coefficients
+#scatter.fit <- summary(m8s)$coefficients
+#colnames(scatter.fit)[4] <- "p-value"
+#rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
 #colnames(line.fit)[4] <- "p-value"
 #rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
-kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
-kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
+#kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
+#kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
 @

### 2015-12-17 paper/paper.Rnw
small typos

@@
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$ $x_{2,j} is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$ $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
 
@@
 The model is fitted using the glimmix procedure in SAS.
-Subjects with less than ten lineup evaluations were removed from the sample.
-%Sign = 1 if correlation > 0, 0 otherwise.
+Subjects with less than ten lineup evaluations were removed from the sample.} %Sign = 1 if correlation > 0, 0 otherwise.
 Each plot design is estimated separately.
@@
 \centering
-\caption{Mixed Linear Model fit for time taken to make a selection}
+\caption{Linear mixed model fit for time taken to make a selection}
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=

### 2015-12-17 paper/paper.Rnw
Rewrite model predict

--- 2015-12-17 paper/paper.Rnw
@@
 \caption{Weighted Mixed Effects Logit Model for the predicted power of plot design.
-Sample size for scatter plot model: 900, overlaid lines model: 949, total: 1849.
+Sample size for scatter plot evaluations: 900, overlaid lines: 949, total: 1849.
 Default plot design is an overlaid lines type.
@@
 full.sub$rabs <- abs(full.sub$r)
-#full.sub$rabs <- factor(full.sub$rabs)
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
@@
 full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
-#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
-#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
 full.sub$alpha <- log(.05/0.95)
-#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
-#weirw.csv(full.sub, "choices.csv") #exporting data to be estimated in SAS.
-#m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
-#m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
-#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
-#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
-#library(xtable)
-#line.fit <- summary(m8l)$coefficients
-#scatter.fit <- summary(m8s)$coefficients
-#colnames(scatter.fit)[4] <- "p-value"
-#rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smoothed", "t", "Interaction(r+, r-)")
-#colnames(line.fit)[4] <- "p-value"
-#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
-#kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
-#kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
+full.sub$subj_id <- as.numeric(full.sub$subj_id)
+subsetted <- select(full.sub, rabs, n, alpha, rsgn, smoothed, plot, subj_id, weights, correct)
+#write.csv(subsetted, "sas/choices.csv")
+
+powercoef <- read.csv("sas/PowerEstimates.csv")
+powercoef <- powercoef[,-1]
+rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
+kable(powercoef[1:6,], format="latex", digits=2, caption="Main Effects")
+kable(powercoef[7:12,], format="latex", digits=2, caption="Interaction Effects")
 @
@@
 <<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
-r <- seq(-0.9,0.9,0.02)
-n <- c(12, 24, 48, 96)
-predl <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
-                                      smoothed=c(0,1), 
-                                      subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
-predl$rsgn <- ifelse(predl$r > 0, 1, 0)
-predl$rabs <- abs(predl$r)
-predl$fitre <- predict(m8l, predl, type="response", re.form=~(1|subj_id),
-                             allow.new.levels=TRUE)
-predl$fitfx <- predict(m8l, predl, type="response", re.form=NA,
-                             allow.new.levels=TRUE)
-predl$plot <- ifelse(predl$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
-predl <- filter(predl, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
+rabs <- seq(0,0.9,0.02)
+t <- c(12, 24, 48, 96)
+#pred <- data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
+#                                      smoothed=c(0,1), plot=c(0,1),
+#                                      subj_id=full.sub$subj_id))
+#write.csv(pred, "sas/predgrid.csv") # these lines are done, but its a large file so commented out for knitr
+#Code exported to sas, see sas/model\ fit.sas
+prediction <- read.csv("sas/predfit.csv") #Large file, very slow
+rownames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "RandomEf", "FixedEf", "resid")
+resid <- subset(prediction, detected==0 | detected==1)
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==1, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
+prediction <- select(prediction, rabs, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
+prediction <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
 
-preds <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
-                                      smoothed=c(0,1), 
-                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
-preds$rsgn <- ifelse(preds$r > 0, 1, 0)
-preds$rabs <- abs(preds$r)
-preds$fitre <- predict(m8s, preds, type="response", re.form=~(1|subj_id),
-                             allow.new.levels=TRUE)
-preds$fitfx <- predict(m8s, preds, type="response", re.form=NA,
-                             allow.new.levels=TRUE)
-preds$plot <- ifelse(preds$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
-preds <- filter(preds, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
-full.pred <- rbind(preds, predl)
-
-#full.pred <- mutate(full.pred2, plot2 = ifelse(substr(plot, 1, 1) =="S", "Scatterplot", "Overlaid Lines"))
-
-qplot(r, fitre, data=full.pred, color=plot, 
+qplot(r, RandomEf, data=prediction, color=plot, 
       group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
   scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
-  xlab("Correlation") + ylab("Estimated Power") + labs(title="Predicted Power") +
-  geom_line(aes(y=fitfx), size=2, alpha=1) + theme_bw() + facet_wrap(~n, ncol = 2) + theme(aspect.ratio=1.2)
+  xlab("Correlation") + ylab("Estimated Power") +
+  geom_line(aes(y=FixedEf), size=2, alpha=1) + theme_bw() + facet_wrap(~t, ncol = 2) + theme(aspect.ratio=1.2)
 @
@@
 <<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
-resids <- resid(m8s)
-residl <- resid(m8l)
-resids[842:843] <- NA
-resid <- data.frame(line = residl, scatter = resids)
-p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
-p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
-grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
+resids <- subset(resid, plot==1)
+residl <- subset(resid, plot==0)
+#resids[842:843] <- NA  #fix residual stuff later
+#resid <- data.frame(line = residl, scatter = resids)
+#p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
+#p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
+#grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
 @
@@
 \centering
-\caption{Mixed Linear Model fit for time taken to make a selection}
+\caption{Mixed Linear Model fit for time taken to make a selection, default plot style is overlaid lines.}
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
@@
 rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
-kable(time.fit[1:7,], format="latex", caption="Main Effects (Overlaid Lines)")
-kable(time.fit[8:14,], format="latex", caption="Interaction Effects (Scatter plot)")
+kable(time.fit[1:7,], format="latex", caption="Main Effects", digits=2)
+kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
 @

### 2015-12-17 paper/paper.Rnw
Integrated typo fixes

@@
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$ $x_{2,j} is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
 
@@
 The model is fitted using the glimmix procedure in SAS.
-Subjects with less than ten lineup evaluations were removed from the sample.
-%Sign = 1 if correlation > 0, 0 otherwise.
+Subjects with less than ten lineup evaluations were removed from the sample.} %Sign = 1 if correlation > 0, 0 otherwise.
 Each plot design is estimated separately.
 Results must be transformed via the logit link function to generate a power prediction.
-Subjects with less than ten lineup evaluations were removed from the sample..}
+Subjects with less than ten lineup evaluations were removed from the sample.
 \label{tab1}
@@
 \centering
-\caption{Mixed Linear Model fit for time taken to make a selection, default plot style is overlaid lines.}
+\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines.}
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=

### 2015-12-18 paper/paper.Rnw
Fixed

@@
 Subjects with less than ten lineup evaluations were removed from the sample.
+
 \label{tab1}
@@
 prediction <- read.csv("sas/predfit.csv") #Large file, very slow
-rownames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "RandomEf", "FixedEf", "resid")
-resid <- subset(prediction, detected==0 | detected==1)
+colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "RandomEf", "FixedEf", "residual")
+glmm.resid <- subset(prediction, detected==0 | detected==1)
 prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==1, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
@@
 <<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
-resids <- subset(resid, plot==1)
-residl <- subset(resid, plot==0)
-#resids[842:843] <- NA  #fix residual stuff later
-#resid <- data.frame(line = residl, scatter = resids)
-#p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
-#p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
-#grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
+p1b <- ggplot(data=subset(glmm.resid, plot==1)) + geom_histogram(aes(y=..density.., x=residual)) + labs(title="Scatter plot", x="residual", y=element_blank())
+p2b <- ggplot(data=subset(glmm.resid, plot==0)) + geom_histogram(aes(y=..density.., x=residual))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
+grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
 @
@@
 \caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines.}
+
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=

### 2015-12-18 paper/paper.Rnw
fixes to syntax, update TODO

@@
 rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
-kable(powercoef[1:6,], format="latex", digits=2, caption="Main Effects")
-kable(powercoef[7:12,], format="latex", digits=2, caption="Interaction Effects")
+kable(powercoef[1:6,], format="latex", digits=2)
+kable(powercoef[7:12,], format="latex", digits=2)
 @
@@
 prediction <- read.csv("sas/predfit.csv") #Large file, very slow
-colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "RandomEf", "FixedEf", "residual")
+colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "Included", "RandomEf", "FixedEf", "residual")
 glmm.resid <- subset(prediction, detected==0 | detected==1)
-prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==1, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
-prediction <- select(prediction, rabs, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==0, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
+prediction <- select(prediction, r, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
 prediction <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))

### 2015-12-18 paper/paper.Rnw
More Work

@@
  \centering
-<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
 Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
@@
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
+full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id)#[,c(1,7, 15,16,17,20:26)]
 full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
@@
   \centering
-<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
+<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 p1b <- ggplot(data=subset(glmm.resid, plot==1)) + geom_histogram(aes(y=..density.., x=residual)) + labs(title="Scatter plot", x="residual", y=element_blank())

### 2015-12-18 paper/Extra Material.R

@@ -1,0 +1,24 @@
+#Residuals#
+
+extreme <- subset(glmm.resid, residual <= -10 | residual >= 10)
+extreme <- merge(extreme, full.sub[c(9,11)], by=0)
+
+##sources of extreme residuals:
+#multiple selections - one was correct, others are predicted to be correct but aren't
+#only person to get an easy one wrong
+#only person to get a hard one right
+#investigate pic_id = 61 name = s459.... => true plot in 18 looks alright, but most people picked plot 17 with cor = 0.24. 17 looks very good
+#if you ignore the outliers. Two chose reason 4, one chose reason 1. Last person selected 7 & 20, neither plot is especially interesting.
+
+#Subject evaluations#
+subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights*attempts), attempt = sum(attempts), detect = sum(correct))
+subjplot <- mutate(subjplot, byeval = detect/eval, byattempt = detect/attempt)
+#one subject has no successful evaluations: Saw ten lineups and made ten selections. Harder lineups: low correl scatter, negative lines.
+#did see 0.7 scatter and two 0.9 lines which should be reasonable chance to get correct (pic_id = 181, 140, 44)
+#was only person to get 0.7 scatter wrong (sd4V..., true = 12, chose 1, reason 1), but 0.7 is not high enough to get really large residuals.
+#11/13 detected line 140 (levq...., true = 4, chose 17, reason 2, this one is really easy) and 4/7 deteced in line 44. (lYbEZ, picked 19, true = 3, reason=2)
+#subjects with one successful detection tended to have negative lines and low correlation scatters
+ggplot(data=subjplot, aes(x=byattempt)) + geom_histogram(binwidth=0.05) + ylim(0,30) #proportion: detection to attempts
+#mean 0.37
+ggplot(data=subjplot, aes(x=byeval)) + geom_histogram(binwidth=0.05) + ylim(0,30) # proportion: detection to lineup evals.
+#mean 0.40

### 2016-01-04 paper/paper.Rnw
More work

@@
 The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
+
+If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability equal to $s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
+In the event that a lineup has observers making different numbers of selections, we have a binomial distribution with different success probabilities, and a Poisson Binomial distribution must be used to calculate the p-value.
+The probability mass function of a Poisson Binomial distribution can be written as:
+
+$$P(X=x) = \sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})$$
+
+Where $T_{x}$ is all possible sets of $x$ trials out of $K$ resulting in a success, $A \in T_{x}$ is a specific set of $x$ trials resulting in a success and $A^{c}$ is the complement to $A$, the set of $K-x$ trials resulting in a failure.
+
 \begin{figure}[htbp]
@@
 @
-  \caption{The same AR(1) data both smoothed and unsmoothed. t = 96, correlation = 0.86}
+  \caption{The same AR(1) data both unsmoothed and smoothed for both plot designs. t = 96, correlation = 0.86}
   \label{smoothing}
@@
 We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
-In total there were 2088 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations leaving us with 1684 evaluated lineups for the analysis.
+In total there were 2088 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations leaving us with 1684 lineup evaluations for the analysis.
+
+Subjects were allowed to make multiple selections, which is handled by probability weighting in the model. This increases our data from 1684 to 1849 evaluations.
 
@@
 The dots are data points, the lines represent a naive model average fit.
-Scatter plots generally showed superior performance, with the overlaid line graphs being particularly weak at revealing negative correlation}
+Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are particularly weak at revealing negative correlation.}
 \end{figure}
@@
  \label{powerdiff}
- \caption{Difference in estimated power to reveal the true plot by smoothness and sample size with a 95\% confidence interval.
-Values above zero indicate the scatter plot is more powerful than the overlaid line graph.}
+ \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a 95\% confidence interval.
+The red line indicates the point where both plot designs are equally powerful while values above this line indicate that the scatter plot is more powerful than the overlaid line graph.
+Again, the largest advantage difference appears in negative correlation, where the scatter plot is more powerful.}
 \end{figure}
@@
  \label{samplesize}
- \caption{Effects of increasing sample size for unsmoothed data. Both plot designs appear to benefit from increased sample sizes}
+ \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data.
+The scatter plot appears to universaly benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
 \end{figure}
@@
   \label{confidence2}
-  \caption{Individual subject accuracy versus confidence. There does not appear to be any relationship between the two.}
+  \caption{Individual subject accuracy versus subject average confidence. There does not appear to be any relationship between the two.}
 \end{figure}
@@
   \label{pins}
-  \caption{Pin plots of the subject selections of each lineup.
+  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
 On the x-axis is the correlation of the plot (there is potentially spurious correlation in the null plots), and on the y-axis is the proportion a plot was chosen.
-Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not appear as it has a frequency of zero and is visually blocked by the red pins.
-The pin plots are split with true plot correlation on the columns and sample size on the rows.
-Each plot contains information from all three lineups with that unique combination of factors: plot design, smoothness, sample size and true correlation.}
+Each plot contains data from the three lineups with the requisite characteristics.
+Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins.
+Strongly correlated plot seem to be selected regardless of if the relationship is spurious (a null plot) or real (a true plot).}
 \end{figure}
@@
   \label{anom1}
-  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
-The true plot in position 8 was not detected in five evaluations.
-However, plot 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
-When asked for a reason of their selections, subjects answered that the lines were closest together.}
+  \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
+The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33).
+The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037.
+Subjects most often cited the clustering effect shown as the reason for their selection, an example of the over-interpretation of randomness.}
 \end{figure}
@@
   \label{anom2}
-  \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
-The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33, see formula \ref{eq:1}).
-The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037.
-This null plot shows a unique clustering effect that draws in the observer despite not having any linear relationship in the data, an example of the over-interpretation of randomness.
-Subjects most often cited the clustering pattern formed as the reason for their selection.
-The final selection was position 8, with a correlation of -0.226 (p = 0.33).}
+  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
+The true plot in position 8 was not detected in five evaluations.
+However, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
+When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.}
 \end{figure}
@@
 Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
-In Figure~\ref{anom1}, the true data plot is in position 4, but 9/11 subjects chose plot 6.
+In Figure~\ref{anom1}, the true data plot is in position 4, but 9 out of 11 subjects chose plot 6.
 Plot 6 is unusual, most of the points are scattered around the vertical median, but there are two outlying points, one with a high y-value and one with a low y-value.
@@
 
-$$Power_{ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma_{1} \cdot z_{1,s} \right\}$$
+$$Power_{ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$.
+It is assumed that $\epsilon_{ijkt} \sim \mathcal{N} (0, \Sigma_{ijkt})$ and $\gamma \sim \mathcal{N} (0, \delta)$
+
+Where a subject made multiple selections, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
 
@@
 Sample size for scatter plot evaluations: 900, overlaid lines: 949, total: 1849.
-Default plot design is an overlaid lines type.
-The model is fitted using the glimmix procedure in SAS.
-Subjects with less than ten lineup evaluations were removed from the sample.} %Sign = 1 if correlation > 0, 0 otherwise.
+Default plot design is overlaid lines.
+The model is fitted using the GLIMMIX procedure in SAS.} %Sign = 1 if correlation > 0, 0 otherwise.
 Each plot design is estimated separately.
@@
 rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
-kable(powercoef[1:6,], format="latex", digits=2)
-kable(powercoef[7:12,], format="latex", digits=2)
+kable(powercoef, format="latex", digits=2)
+#kable(powercoef[7:12,], format="latex", digits=2)
 @
@@
 <<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
-rabs <- seq(0,0.9,0.02)
+rabs <- seq(0,0.9,0.03)
 t <- c(12, 24, 48, 96)
@@
 #                                      subj_id=full.sub$subj_id))
+#pred <- subset(pred, smoothed==0 | smoothed == 1 & n ==48 | smoothed ==1 & n ==96)
 #write.csv(pred, "sas/predgrid.csv") # these lines are done, but its a large file so commented out for knitr
@@
   \label{powermodel}
-  \caption{Predicted power fit from the mixed effects logit power model. The weakness of the overlaid line graph in negative correlation and  the scatter plot's gain from sample size are highlighted.}
+  \caption{Predicted power fit from the mixed effects logit power model.
+The thicker lines shows the fixed effects prediction while the thinner lines show the range of random effects, the observer's visual ability.
+The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
 \end{figure}
@@
   \label{residuals}
-  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
+  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal}
 \end{figure}
@@
 Also concerning, the mean value of the overlaid lines residuals is - 0.08, implicating that the model tends to over-predict power, further weakening the performance of this plot design relative to the scatter plot.
-This unrestricted model used in to test the symmetry restriction for the scatter plot design is written formally as:
+This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
 
-$$Power_{UR,s} = \psi\left\{ \beta_{s,0} + \beta_{s,1} \cdot |\alpha| + \beta_{s,2}\cdot Sign + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t + \beta_{s,5}\cdot |\alpha| \cdot Sign + \gamma_{s,1} \cdot subject_{j} \right\}$$
+$$Power_{UR,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
 
@@
 The restricted model is therefore:
-$$Power_{R,s} = \psi\left\{ \beta_{s,0}+\beta_{s,1}\cdot |\alpha| + \beta_{s,3}\cdot Smoothed + \beta_{s,4}\cdot t +\gamma_{s,1} \cdot subject_{j} \right\}$$
+
+$$Power_{R,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i} + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
 
@@
 rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
-kable(time.fit[1:7,], format="latex", caption="Main Effects", digits=2)
-kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
+kable(time.fit[1:14,], format="latex", caption="Main Effects", digits=2)
+#kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
 @

### 2016-01-05 paper/paper.Rnw
Changes

@@
   \centering
-<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
+<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(1234)
@@
   \centering
-<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
+<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(123)
@@
   \centering
-<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE>>=
+<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(101)
@@
  \centering
-<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
+<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
 Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
@@
   \centering
-<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 Turk$correct <- Turk$percent*Turk$attempt
@@
   \centering
-<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
+<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
@@
   \centering
-<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 full <- mutate(full, sign = ifelse(correlation >= 0, 1, 0), n24 = ifelse(n ==24, 1, 0), n48 = ifelse(n ==48,  1, 0), n96 = ifelse(n ==96, 1, 0), plot = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"))
@@
   \centering
-<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 conf <- data.frame(plot=full$plot, conf=full$conf_level, correct=full$correct, rsgn=full$sign)
@@
   \centering
-<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
+<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
   \centering
-<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
+<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 # correl <- read.csv("correlations.csv") # You have already read this
@@
 pin <- merge(counts, Turk[-19], by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r, attempt)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
-colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
+colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r", "attempt")
 pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
@@
 Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins.
-Strongly correlated plot seem to be selected regardless of if the relationship is spurious (a null plot) or real (a true plot).}
+Strongly correlated plot seem to be selected regardless of if the relationship is real or spurious.}
 \end{figure}
@@
 \centering
-<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
+<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 full$subj_id <- factor(full$ip_address, labels=1:207)
@@
 full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id)#[,c(1,7, 15,16,17,20:26)]
+full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id, choice_reason)#[,c(1,7, 15,16,17,20:26)]
 full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
@@
  \centering
-<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
+<<powermodel, dependson='powerfit', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 rabs <- seq(0,0.9,0.03)

### 2016-01-05 paper/Extra Material.R

@@ -22,3 +22,22 @@ ggplot(data=subjplot, aes(x=byattempt)) + geom_histogram(binwidth=0.05) + ylim(0
 #mean 0.37
 ggplot(data=subjplot, aes(x=byeval)) + geom_histogram(binwidth=0.05) + ylim(0,30) # proportion: detection to lineup evals.
 #mean 0.40
+
+#Choice plots
+full.sub$choice_reason <-as.numeric(unlist(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+")))
+#Multiple selections#
+m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
+##112 subject/lineup combinations
+##53 different subjects choose multiple plots
+##84 different lineups had multiple selections, 61 with 1, 19 with 2, 3 with 2, 1 with  
+##55 lines, 29 scatter (1 person selecting multiple - 40 lines, 21 scatter
+#                       2 people selecting multiple - 11 lines, 8 scatter
+#                       3 people selecting multiple - 3 lines
+#                       4 people selecting multiple - 1 line
+)
+
+m.freq <- data.frame(table(m.select$pic_id))
+m.lineup <- merge(m.freq, Turk, by.x="Var1", by.y="pic_id")
+ggplot(data=subset(m.lineup, Freq==2)) + geom_bar(aes(x=r, fill=plot)) + facet_wrap(~n)
+#high sample size increases chance of multiple selection, even spread across correlations except -0.5/-0.3 lines t=96
+

### 2016-01-06 paper/paper.Rnw
SAS plus more

@@
 full.sub$subj_id <- as.numeric(full.sub$subj_id)
-subsetted <- select(full.sub, rabs, n, alpha, rsgn, smoothed, plot, subj_id, weights, correct)
-#write.csv(subsetted, "sas/choices.csv")
+subsetted <- select(full.sub, rabs, n, alpha, rsgn, smoothed, plot, subj_id, weights, correct, log_time)
+subsetted$rabs <- as.numeric(subsetted$rabs)
+subsetted$subj_id <- as.numeric(subsetted$subj_id)
+subsetted <- mutate(subsetted, plot = ifelse(plot=="S", 1, 0))
+#write.csv(subsetted, "sas/choices.csv", row.names = FALSE)
 
@@
 rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
+colnames(powercoef) <- c("Estimate", "standard error", "df", "t-value", "p-value")
 kable(powercoef, format="latex", digits=2)
@@
 <<powermodel, dependson='powerfit', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-rabs <- seq(0,0.9,0.03)
-t <- c(12, 24, 48, 96)
-#pred <- data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
+# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
 #                                      smoothed=c(0,1), plot=c(0,1),
 #                                      subj_id=full.sub$subj_id))
-#pred <- subset(pred, smoothed==0 | smoothed == 1 & n ==48 | smoothed ==1 & n ==96)
-#write.csv(pred, "sas/predgrid.csv") # these lines are done, but its a large file so commented out for knitr
+#pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)
+#write.csv(pred, "sas/predgrid.csv", row.names=FALSE) # these lines are done, but its a large file so commented out for knitr
 #Code exported to sas, see sas/model\ fit.sas
 prediction <- read.csv("sas/predfit.csv") #Large file, very slow
-colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "Included", "RandomEf", "FixedEf", "residual")
+colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
 glmm.resid <- subset(prediction, detected==0 | detected==1)
-prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==0, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot))
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==0, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot),  diff = round(RandomEf - FixedEf,5))
+randomeffects <- summarise(group_by(prediction, subj_id), re=max(diff)) #all values of diff are identical for a given subject, max chosen just as a function to select one of them
 prediction <- select(prediction, r, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
@@
   \label{residuals}
-  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal}
+  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
 \end{figure}
@@
 Also concerning, the mean value of the overlaid lines residuals is - 0.08, implicating that the model tends to over-predict power, further weakening the performance of this plot design relative to the scatter plot.
+
+\begin{figure}
+  \centering
+<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..)) + xlab("Random Effects") + ylab(NULL)
+@
+  \label{random}
+  \caption{Histograms of estimated subject random effects, $\hat{\gamma}$}
+\end{figure}
 This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
@@
 \centering
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines.}
+\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS}
 
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
-timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
+#timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                   + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
-time.fit <- summary(timelmer)$coefficients
-pv <- function(x){
-  p <- 2*(1-pt(x, 1835))
-}
-p.value <- sapply(time.fit[,3], pv)
-time.fit <- cbind(time.fit, p.value)
-time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
-rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "correct", "Plot", "Plot*Smoothed", "Plot*t", "Plot*r", "Plot*sign(r)", "Plot*r*sign(r)", "Plot*correct")
-kable(time.fit[1:14,], format="latex", caption="Main Effects", digits=2)
+#time.fit <- summary(timelmer)$coefficients
+#pv <- function(x){
+#  p <- 2*(1-pt(x, 1835))
+#}
+#p.value <- sapply(time.fit[,3], pv)
+#time.fit <- cbind(time.fit, p.value)
+time.fit <- read.csv("sas/time.csv")
+time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
+rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "detected", "Plot", "Plot*Smoothed", "Plot*t", "Plot*abs(r)", "Plot*sign(r)", "Plot*abs(r)*sign(r)", "Plot*detected")
+time.fit$Effect <- NULL
+colnames(time.fit) <- c("Estimate", "standard error", "df", "t-value", "p-value")
+kable(time.fit[1:14,], format="latex", digits=2)
 #kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)

### 2016-01-06 paper/Extra Material.R

@@ -41,3 +41,9 @@ m.lineup <- merge(m.freq, Turk, by.x="Var1", by.y="pic_id")
 ggplot(data=subset(m.lineup, Freq==2)) + geom_bar(aes(x=r, fill=plot)) + facet_wrap(~n)
 #high sample size increases chance of multiple selection, even spread across correlations except -0.5/-0.3 lines t=96
 
+#Subject ability#
+ability <- merge(full.sub[,c("subj_id", "pic_id", "correct")], Turk[,c("pic_id", "percent")], by="pic_id")
+ability <- summarise(group_by(ability, subj_id), actual=sum(correct), estimated=sum(percent))
+ability$diff <- ability$actual - ability$estimated
+ability.re <- merge(ability, randomeffects, by="subj_id")
+ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point()

### 2016-01-06 paper/paper.Rnw
Updates

@@
 keep <- x$Var1[x$Freq>9]
-full.sub <- filter(full, subj_id %in% keep)
-full.sub$r <- round(full.sub$correlation, 1)
-full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
-full.sub$rabs <- abs(full.sub$r)
-full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
-full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id, choice_reason)#[,c(1,7, 15,16,17,20:26)]
-full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
-full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
-full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+full.subs <- filter(full, subj_id %in% keep)
+full.subs$r <- round(full.subs$correlation, 1)
+full.subs$rsgn <- ifelse(full.subs$r >= 0, 1, 0)
+full.subs$rabs <- abs(full.subs$r)
+full.subs$plot <- ifelse(full.subs$test_param == 0, "S", "L")
+full.subs <- select(full.subs, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id, choice_reason)#[,c(1,7, 15,16,17,20:26)]
+full.subs$attempts <- sapply(strsplit(as.character(full.subs$response_0), ","), length)
+full.subs$weights = 1/(full.subs$attempts) #weights are actually number of attempts in the glm
+full.sub <- full.subs %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
 full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses

### 2016-01-06 paper/Extra Material.R

@@ -23,12 +23,22 @@ ggplot(data=subjplot, aes(x=byattempt)) + geom_histogram(binwidth=0.05) + ylim(0
 ggplot(data=subjplot, aes(x=byeval)) + geom_histogram(binwidth=0.05) + ylim(0,30) # proportion: detection to lineup evals.
 #mean 0.40
 
-#Choice plots
-full.sub$choice_reason <-as.numeric(unlist(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+")))
+#Choice plots#
+full.subs$choice_reason <-sapply(strsplit(as.character(unlist(full.subs$choice_reason)), "[^0-9]+"), "[[", 1)
+ggplot(data=full.subs) + geom_bar(aes(x=choice_reason))
+choice <- full.subs %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
+choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
+#tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
+c1 <- ggplot(data=subset(choice, correct==0), aes(x=choice_reason)) + geom_bar() + 
+  facet_grid(plot~rsgn) + labs(title="Did not detect", x=element_blank(), y=element_blank())  
+c2 <- ggplot(data=subset(choice, correct==1), aes(x=choice_reason)) + geom_bar() + 
+  facet_grid(plot~rsgn) + labs(title="Detected", x=element_blank, y=element_blank())  
+grid.arrange(c1, c2, ncol=2, top = "Choice by plot design, true plot correlation and detection")
+
 #Multiple selections#
 m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
 ##112 subject/lineup combinations
-##53 different subjects choose multiple plots
+##53 different subjects chose multiple plots
 ##84 different lineups had multiple selections, 61 with 1, 19 with 2, 3 with 2, 1 with  
 ##55 lines, 29 scatter (1 person selecting multiple - 40 lines, 21 scatter
 #                       2 people selecting multiple - 11 lines, 8 scatter
@@ -42,8 +52,12 @@ ggplot(data=subset(m.lineup, Freq==2)) + geom_bar(aes(x=r, fill=plot)) + facet_w
 #high sample size increases chance of multiple selection, even spread across correlations except -0.5/-0.3 lines t=96
 
[…]

### 2016-01-07 paper/paper.Rnw
Appendix

@@
 library(stringr)
+library(ipapi)
+library(maps)
+library(moments)
 #library(pander)
@@
 
-$$Power_{ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
+$$Power_{ijklt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijkt} \right\}$$
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$.
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type.
+Finally, $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$.
 It is assumed that $\epsilon_{ijkt} \sim \mathcal{N} (0, \Sigma_{ijkt})$ and $\gamma \sim \mathcal{N} (0, \delta)$
@@
 grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
+jarque.test(subset(glmm.resid), plot==1)$residual
+jarque.test(subset(glmm.resid), plot==0)$residual
 @
@@
 ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..)) + xlab("Random Effects") + ylab(NULL)
+jarque.test(randomeffects$re)
 @
@@
 
+\section{Appendix}
+
+\begin{table}
+\lab{tab3}
+\centering
+<<answertables, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+turk.tbl <- select(Turk, test_param, smoothed, r, n, correct, attempt)
+turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$attempt, sep="/")
+turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
+scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="scatter")$result, ncol=8))
+scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="scatter")$result, ncol=8))
+line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="line")$result, ncol=8))
+line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="line")$result, ncol=8))
+colnames(scatter.uns) <- c("", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(scatter.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.uns) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+kable(scatter.uns, format="latex", caption="Unsmoothed scatter plot results")
+kable(scatter.sm, format="latex", caption="Smoothed scatter plot results")
+kable(line.uns, format="latex", caption="Unsmoothed overlaid line results")
+kable(line.sm, format="latex", caption="Smoothed overlaid line results")
+@
+\end{table}
+
+Table~\ref{tab3} contains the true plot detections / attempts for each lineup in the data collection. True plot correlation is on the x axis and t is on the y axis.
+
+\begin{figure}[htbp]
+\lab{map}
+\centering
+<<map, , dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+location <- read.csv("location.csv")
+map.dat <- map_data("world")
+ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
+@
+\end{figure}
+
+Figure~\ref{map} shows the location of the recruited subjects. The majority reside in the United States but some are spread around other countries in Europe and Asia.
+
 \end{document}

### 2016-01-07 paper/Extra Material.R

@@ -28,7 +28,6 @@ full.subs$choice_reason <-sapply(strsplit(as.character(unlist(full.subs$choice_r
 ggplot(data=full.subs) + geom_bar(aes(x=choice_reason))
 choice <- full.subs %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
 choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
-#tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
 c1 <- ggplot(data=subset(choice, correct==0), aes(x=choice_reason)) + geom_bar() + 
   facet_grid(plot~rsgn) + labs(title="Did not detect", x=element_blank(), y=element_blank())  
 c2 <- ggplot(data=subset(choice, correct==1), aes(x=choice_reason)) + geom_bar() + 
@@ -60,4 +59,16 @@ ability$diff <- ability$actual - ability$estimated
 ability.re <- merge(ability, randomeffects, by="subj_id")
 ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point()
 
+#Results Tables#
+turk.tbl <- select(Turk, test_param, smoothed, r, n, correct, attempt)
+turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$attempt, sep="/")
+turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
+scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="scatter")$result, ncol=8))
+scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="scatter")$result, ncol=8))
+line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="line")$result, ncol=8))
+line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="line")$result, ncol=8))
[…]

### 2016-01-08 paper/paper.Rnw
Changes

@@
   \item Factor 2: Sample size (t), levels = 12, 24, 48, 96
-  \item Factor 3: True plot correlation, levels 
+  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$
 \end{itemize}
@@
 counts$Var2 <- as.numeric(as.character(counts$Var2))
-colnames(counts) <- c("lineup", "plot", "count")
-pin <- merge(counts, Turk[-19], by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r, attempt)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+colnames(counts) <- c("lineup", "plots", "count")
+pin <- merge(counts, Turk[-17], by.x="lineup", by.y="pic_id")
+pin <- select(pin, lineup, plots, count, smoothed, true_pos, n, test_param, r, attempt)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
 pin$r <- factor(pin$r)
@@
 
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type.
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, $\rho$, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type.
 Finally, $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$.
@@
 
-Table \ref{tab1} summarises the model fits.
-It can be confusing the interpret the model estimates, and at first glance it appears the overlaid lines (-0.660) are substantially more powerful than the scatter plot (-4.403), but this is misleading, and  interpreting the the full model involves combining the estimates for the different interactions with correlation and sign.
-Taking correlation into account, the scatter plot shows superior performance.
-It needs to be noted that we only tested correlations at  0.3, 0.5, 0.7 and 0.9, so the results below an absolute value of 0.3 are unreliable and may be artifacts of regression smoothing methods.
-The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Surprisingly, smoothing out the excess 'noise' from a time series had a negative effect on scatter plots, despite a minimal visual impact of smoothed (Figure~\ref{smoothing}).
-It did have a significant positive impact on the power of overlaid line graphs.
+Table \ref{tab1} summarises the model estimates.
+It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overliad lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with abs($\rho$) and sign($\rho$).
+abs($\rho$) models the power difference between low and high correlations, and abs($\rho)*sign($\rho$) models an increase in power appearing in positive but not negative correlation.
+Taking these into account, the scatter plot shows superior performance and higher values of abs($\rho$), It needs to be restated that we only tested correlations at  0.3, 0.5, 0.7 and 0.9, so the results below an absolute value of 0.3 are unreliable and may be artifacts of regression smoothing methods.
+
+%The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
+Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots(...), it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (Figure~\ref{smoothing}.
 Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
+
 Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
@@
   \centering
-<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-p1b <- ggplot(data=subset(glmm.resid, plot==1)) + geom_histogram(aes(y=..density.., x=residual)) + labs(title="Scatter plot", x="residual", y=element_blank())
-p2b <- ggplot(data=subset(glmm.resid, plot==0)) + geom_histogram(aes(y=..density.., x=residual))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
-grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
-jarque.test(subset(glmm.resid), plot==1)$residual
-jarque.test(subset(glmm.resid), plot==0)$residual
+<<residuals, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=glmm.resid) + geom_histogram(aes(y=..density.., x=residual), binwidth=2) + labs(x=NULL, y=NULL)
 @
@@
 
-Analysing the residuals allows us to test the model fit.
-As the dependent variable is a binary variable, the model will only have a residual close to zero if it predicts either a power of zero or one.
-As the rate of picking the true plot purely by chance is, $1/m$ or 0.05, the model should never predict a power of zero even if the plot has no ability to display correlation.
-This effect is especially notable in the overlaid line residual plot, the model fit is never close to zero or one so the residuals are split with no values around zero.
-The scatter plot residuals are normally distributed with a Jarque-Bera normality test statistic of 0.09 (p-value = 0.95), however the overlaid lines strongly reject normality with a Jarque-Bera statistic of 75.6 (p-value = 0.00).
-The significance tests associated with the overlaid lines are only valid asymptotically.
-Also concerning, the mean value of the overlaid lines residuals is - 0.08, implicating that the model tends to over-predict power, further weakening the performance of this plot design relative to the scatter plot.
-
-\begin{figure}
-  \centering
-<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..)) + xlab("Random Effects") + ylab(NULL)
-jarque.test(randomeffects$re)
+Analysing the level one residuals, $\hat{\epsilon_{i}}$ and level two residuals, $\hat{\gamma_{i}}$ allows us to test the model fit as described \citet{Loy}.
+There are some negative extreme outliers but the distributional assumption appears to hold.
+
+\begin{figure}[htbp]
+  \centering
+<<random, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
 @
@@
 \end{figure}
+
+In Figure~\ref{random}, the level two residuals, $\hat{\gamma}$, generated appear to be normally distributed validating the assumptions of the modelling process.
+For each lineup, we rate the difficulty by the proportion of true plot detections to evaluations.
+Then, for each subject the difficulty of each lineup they evaluated is summed to give a measure of the expected number of detections.
+The visual ability of a subject is defined as the difference between the actual and expected number of detections, and is compared to the random effects in Figure~\ref{random2} to further validate the model.
+
+\begin{figure}[htbp]
+\centering
+\label{random2}
+<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+turk.correct <- summarise(group_by(full.sub, pic_id), correct=sum(correct), attempts=sum(weights*attempts))
+turk.correct$percent <- turk.correct$correct/turk.correct$attempts
+ability <- merge(full.sub[,c("subj_id", "pic_id", "correct")], turk.correct[c("pic_id", "percent")], by="pic_id")
+ability <- summarise(group_by(ability, subj_id), actual=sum(correct), estimated=sum(percent))
+ability$diff <- ability$actual - ability$estimated
+ability.re <- merge(ability, randomeffects, by="subj_id")
+ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs{x="Visual Ability", y="Random Effects"}
+@
+\caption{Visual ability plotted against random effects.
+The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
+\end{figure}
+
+\iffalse
 This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
@@
 Such an extreme result is due to a systematic difference in the overlaid line graph and cannot be attributed simply to bias.
+\fi
 

### 2016-01-08 paper/paper.Rnw
equations and little things

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+%\SweaveOpts{concordance=TRUE}
 
@@
 
-$$Power_{ijklt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijkt} \right\}$$
+\begin{eqnarray*} 
+Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
+ & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \\
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijkt} \}
+\end{eqnarray*}
 
@@
 It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overliad lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with abs($\rho$) and sign($\rho$).
-abs($\rho$) models the power difference between low and high correlations, and abs($\rho)*sign($\rho$) models an increase in power appearing in positive but not negative correlation.
+abs($\rho$) models the power difference between low and high correlations, and $\mbox{abs}(\rho)*\mbox{sign}(\rho$) models an increase in power appearing in positive but not negative correlation.
 Taking these into account, the scatter plot shows superior performance and higher values of abs($\rho$), It needs to be restated that we only tested correlations at  0.3, 0.5, 0.7 and 0.9, so the results below an absolute value of 0.3 are unreliable and may be artifacts of regression smoothing methods.
@@
 ability.re <- merge(ability, randomeffects, by="subj_id")
-ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs{x="Visual Ability", y="Random Effects"}
+ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs(x="Visual Ability", y="Random Effects")
 @
@@
 #timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
-                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
+#                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
 #time.fit <- summary(timelmer)$coefficients
@@
 \begin{table}
-\lab{tab3}
+\label{tab3}
 \centering
@@
 \begin{figure}[htbp]
-\lab{map}
+\label{map}
 \centering

### 2016-01-08 paper/paper.Rnw
Cleanup

@@
 library(stringr)
-library(ipapi)
+#library(ipapi)
 library(maps)
@@
 It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993.
-\citet[Vanderplas] examines this effect in more detail.
+\citet{Vanderplas} examines this effect in more detail.
 It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope.
@@
 
-\begin{equation} \label{eq:1}
+\begin{equation}
+\label{pvalue}
   \centering
@@
 If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
-This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $α = 0.05$.
+This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $\alpha = 0.05$.
 The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
 
-If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability equal to $s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
+If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability $p_{j} = s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
 In the event that a lineup has observers making different numbers of selections, we have a binomial distribution with different success probabilities, and a Poisson Binomial distribution must be used to calculate the p-value.
@@
 
-$$P(X=x) = \sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})$$
+\begin{equation} 
+\label{pvalue2}
+\centering
+P(X=x) = \sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})
+\end{equation}
 
@@
 We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ correctly detected the true plot out of the lineup.
-We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as \citep{Hoffman:2012}:
-
-\begin{equation} \label{eq:2}
+We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as \citep{Hofmann:2012}:
+
+\begin{equation} \label{propci}
   \centering
@@
 Further details of our use of Amazons’ Mechanical Turk is included in Section 3.
-The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in \citet{Hoffman:2012}, and this work follows that approach.
+The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in \citet{Hofmann:2012}, and this work follows that approach.
 
@@
 
-$$~N(\mu= \left| \begin{array}{c}0 \\0 \end{array} \right|,\hspace{4mm}\Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right|)$$
+$$~\mathcal{N}\left\(\mu= \left| \begin{array}{c}0 \\0 \end{array} \right|,\hspace{4mm}\Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right|\right\)$$
 
@@
 The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$.
-True data pairs are generated with $\rho = \pm 0.3, 0.5, 0.7, 0.9$.
+True data pairs are generated with $\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$.
 Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$.
@@
 The null pairs often had correlation generated spuriously through the simulation process.
-This was particularly problematic for generating small sample data, but setting, to a maximum of 0.5, for both true and null data reduced spurious correlation to manageable levels.
+This was particularly problematic for generating small sample data, but setting $\beta = max(0.5,100t)$ for both true and null data reduced spurious correlation to manageable levels.
 For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications.
@@
 
-$$\min_{\tau}(\sum_{t=1}^{T}(y_{t}-\tau_{t})^2 + \lambda \sum_{t=2}^{T-1}[(\tau_{t+1}-\tau_{t})-(\tau_{t}-\tau_{t-1})]^2)$$
+\begin{equation}
+\centering
+\label{hpfilter}
+\min_{\tau}(\sum_{t=1}^{T}(y_{t}-\tau_{t})^2 + \lambda \sum_{t=2}^{T-1}[(\tau_{t+1}-\tau_{t})-(\tau_{t}-\tau_{t-1})]^2)
+\end{equation}
 
 Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation.
-However, for $t = 12, 24$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = 48, 96$.
+However, for $t = \{12, 24\}$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = \{48, 96\}$.
 Each generated lineup was produced as both a scatter plot and an overlaid line graph.
@@
   \item Factor 1: Plot design, levels = scatter plot, overlaid time series
-  \item Factor 2: Sample size (t), levels = 12, 24, 48, 96
+  \item Factor 2: Sample size (t), levels = \{12, 24, 48, 96\}
   \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$
@@
 
-Additionally, there was a fourth factor for $t = 48, 96$:
+Additionally, there was a fourth factor for $t = \{48, 96\}$:
 
@@
 
-Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (2 plots x 8 correlations x 6 sample sizes \& smoothness combinations). 
+Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (3 replications x 2 plots x 8 correlations x 6 sample size and smoothness combinations). 
 The position of the actual data plot in the lineup was randomized.
@@
 ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
-  scale_colour_discrete(name = "Plot Design") + labs(title="Observed Data", y="Percent Correct", x="Correlation") +
+  scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
   theme(legend.position="bottom")
@@
 
-Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to the formula in \ref{eq:2}.
+Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to  formula~\ref{propci}.
 Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
@@
 ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-        geom_hline(aes(yint=0, colour="red")) + labs(title="Difference in Power", x="Correlation", y="Scatterplot Advantage") + facet_grid(n~smoothed)
+        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
 @
@@
 The red line indicates the point where both plot designs are equally powerful while values above this line indicate that the scatter plot is more powerful than the overlaid line graph.
-Again, the largest advantage difference appears in negative correlation, where the scatter plot is more powerful.}
+The largest differences appear in negative correlation, where the scatter plot is more powerful.}
 \end{figure}
@@
 <<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(title="Effect of sample size, with unsmoothed data", y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
+ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
 @
@@
   facet_grid(plot~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
-grid.arrange(t1, t2, ncol=2, top = "Log(Time) by plot design, true plot correlation and detection")
+grid.arrange(t1, t2, ncol=2)
 @
@@
 This would suggest that the null plots often produced visually interesting features subjects assumed were caused by the association between plots.
-
 Subjects spending longer on a plot immediately draws two hypotheses:
+
 \begin{enumerate}
@@
   facet_grid(plot~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
-grid.arrange(c1, c2, ncol=2, top = "Confidence by plot design, true plot correlation and detection")
+grid.arrange(c1, c2, ncol=2)
 @
@@
 
-\subsection{Selection Analysis}
-
-Next we analysed the individual ability of each of our subjects.
+%Next we analysed the individual ability of each of our subjects.
 Figure~\ref{confidence2} plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
@@
 
+Figure~\ref{confidence2} compares the reported confidence for each selection against the true plot correlation.
+The results are mixed with no consistent increase or decrease in confidence for changes in correlation across different plot designs and sample size.
+Strong correlation appears to decrease confidence more often than it is increased.
+
 \begin{figure}[htbp]
@@
 keep2 <- y$Var1[y$Freq>9]
-sub <- filter(full, subj_id %in% keep2)
-sub <- select(sub, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
+conf2 <- filter(full, subj_id %in% keep2)
+conf2 <- select(conf2, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
 #sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
@@
 #sub <- mutate(sub, prop = correct/Freq)
-qplot(correlation, 6-conf_level, data=sub, alpha=I(0.1)) + 
+qplot(correlation, 6-conf_level, data=conf2, alpha=I(0.1)) + 
   geom_smooth(method="loess") + 
@@
   \label{confidence2}
-  \caption{Individual subject accuracy versus subject average confidence. There does not appear to be any relationship between the two.}
-\end{figure}
+  \caption{Reported confidence against true plot correlation.
+Darker points indicate a confidence level selected more regularly.
+Strong correlation is more often associated with a lower reported confidence.}
+\end{figure}
+
+\subsection{Selection Analysis}
 
@@
 The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis.
-The correlation of the simulated (true) data (r=-0.9, -0.7, -0.3, 0.3, 0.7, 0.9) in the plot, the sample size (t=12, 24, 48, 96), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
+The correlation of the simulated (true) data ($\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$ in the plot, the sample size ($t= \{12, 24, 48, 96\}$), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
 The blue pins indicate the true plots and the red pins indicate the null plots.
@@
 
-\begin{eqnarray*} 
+\begin{eqnarray} 
+\label{modelpower}
 Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijkt} \}
-\end{eqnarray*}
-
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$, $x_{1}$ is the absolute value of the true plot correlation, $\rho$, with $i \in {0.3, 0.5, 0.7, 0.9}$, $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, $x_{4,t}$ is the sample size, with $t \in {12, 24, 48, 96}$ and $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type.
-Finally, $z_{s}$ is an observer specific random effect, with $s = 1, 2, \cdot , 157$.
-It is assumed that $\epsilon_{ijkt} \sim \mathcal{N} (0, \Sigma_{ijkt})$ and $\gamma \sim \mathcal{N} (0, \delta)$
-
-Where a subject made multiple selections, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
+ & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
+\end{eqnarray}
+
+Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$. Additionally:
+
+\begin{enumerate}
+\item $x_{1,i}$ is the absolute value of the true plot correlation, $\rho$, with $i \in \{0.3, 0.5, 0.7, 0.9\}$,
+\item $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, 
+\item $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, 
+\item $x_{4,t}$ is the sample size, with $t \in \{12, 24, 48, 96\}$ ,
+\item $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type,
+\item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 157$.
+\end{enumerate}
+
+It is assumed that $\epsilon_{ijkt} \sim \mathcal{N} (0, \Sigma_{ijkt})$ and $\gamma \sim \mathcal{N} (0, \delta)$.
+
+Where a subject made multiple selections from the same lineup, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
 
@@
 Table \ref{tab1} summarises the model estimates.
-It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overliad lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with abs($\rho$) and sign($\rho$).
-abs($\rho$) models the power difference between low and high correlations, and $\mbox{abs}(\rho)*\mbox{sign}(\rho$) models an increase in power appearing in positive but not negative correlation.
-Taking these into account, the scatter plot shows superior performance and higher values of abs($\rho$), It needs to be restated that we only tested correlations at  0.3, 0.5, 0.7 and 0.9, so the results below an absolute value of 0.3 are unreliable and may be artifacts of regression smoothing methods.
+It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overliad lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with $abs(\rho)$ and $sign(\rho)$.
+$abs(\rho)$ models the power difference between low and high correlations, and $abs(\rho)*sign(\rho)$ models an increase in power appearing in positive but not negative correlation.
+Taking these into account, the scatter plot shows superior performance and higher values of abs($\rho$), It needs to be restated that we only tested correlations at  $0.3, 0.5, 0.7$ and $0.9$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
 
@@
 \begin{figure}[htbp]
+\label{random}
   \centering
@@
 @
-  \label{random}
     \caption{Histograms of estimated subject random effects, $\hat{\gamma}$}
@@
 
-Table 2: Mixed Effects Linear Model with the dependent variable log(time).
-Sign = 1 if correlation > 0, 0 otherwise.
-Both plot designs estimated together with Overlaid Lines as the base design.
-Sample size: 1684.
-The model is fitted using the lme4 package in R.
+A linear mixed model is used to model the time to make a selection.
+
+\begin{eqnarray} 
+\label{timemodel}
+Time_{ijklmt} &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
+ & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \beta_{13} x_{6,m} + \beta_{14} x_{5_l} \* x_{6,m} \gamma z_{s} + \epsilon_{ijklmt} \nonumber 
+\end{eqnarray}
+
+The structure of the model is similar to that in Equation~\ref{modelpower} except the link function is removed and a new variable, $x_{6,m}$ is introduced, with $m = \{0,1\}$ indicating if the subject detected the true plot.
 
@@
 \centering
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS}
+\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
 
@@
 time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
-rownames(time.fit) <- c("Intercept", "Smoothed", "t", "r", "sign(r)", "r*sign(r)", "detected", "Plot", "Plot*Smoothed", "Plot*t", "Plot*abs(r)", "Plot*sign(r)", "Plot*abs(r)*sign(r)", "Plot*detected")
+rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "Plot", "Plot*Smoothed", "Plot*t", "Plot*abs(r)", "Plot*sign(r)", "Plot*abs(r)*sign(r)", "Plot*detected")
 time.fit$Effect <- NULL
@@
 
-While it appears that the overlaid lines are on the verge of statistical significance (at the 5\% level) for being the faster plot for a subject to visually process, this effect only holds for correlation values close to zero.
-The interactions between correlation and sign significantly reduce the time taken to pick the lineup for stronger correlations.
-Unsurprisingly, correct decisions are made faster.
-Surprisingly, neither smoothness nor sample size has an impact on the time for either plot design, despite having statistically significant impacts on power.
-An overlaid line graph having positive correlation increases the time required, eliminating the benefits from the almost significant plot design term.
+Table~\ref{tab2} shows a similar story for time as Table~\ref{tab1} did for power, that the overlaid lines have better performance close to zero correlation, but the scatter plot is much faster at high correlations.
+Postivie correlation reduces time for both plot designs, but only the scatter plot sees improvements in negative correlation.
+Detecing the true plot was also faster than a null.
+Neither smoothness nor sample size has an impact on the time for either plot design, despite having statistically significant impacts on power.
+%An overlaid line graph having positive correlation increases the time required, eliminating the benefits from the almost significant plot design term.
 This result seems counter-intuitive as the weak performance of the overlaid lines in negative correlation implies that negative would require more time than positive, but the interactions between plot design, correlation and sign result in there being almost no decrease in time for strongly negatively correlated line plots.
@@
 We found that the scatterplot, in most situations is both more powerful and quicker to process than the overlaid line graph.
-The overlaid line graph appeared to have stronger performance at extremely low correlations, but this research did not test absolute values of correlation below 0.3.
+The overlaid line graph appeared to have stronger performance at extremely low correlations, but this work did not test absolute values of correlation below 0.3.
 Other non-linear associations may exist in data that are not measured by correlation, and it may be of interest to compare plot power with a broader range of true plot features.

### 2016-01-08 paper/paper.Rnw
Cleanup

@@
 We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hoffman:2014} with:
+A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:
 
@@
 
-$$~\mathcal{N}\left\(\mu= \left| \begin{array}{c}0 \\0 \end{array} \right|,\hspace{4mm}\Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right|\right\)$$
+$$e_{t} \sim \mathcal{N} \left( \mu = \left| \begin{array}{c}0 \\ 0 \end{array} \right| , \hspace{4mm} \Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right| \right)$$
 
@@
 Hence we found it unnecessary to test a more full range of correlations.
-Simulating data does not produce correlations exactly equal to $\rho$, so imulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$.
+Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$.
 The null pairs often had correlation generated spuriously through the simulation process.

### 2016-01-11 paper/paper.Rnw
changes
TODO and fig refs

@@
 @
+  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
     \label{designs}
-  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
 \end{figure}
@@
 @
-  \label{results}
   \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size.
@@
 Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are particularly weak at revealing negative correlation.}
+    \label{results}
 \end{figure}
@@
 @
- \label{powerdiff}
  \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a 95\% confidence interval.
@@
 The largest differences appear in negative correlation, where the scatter plot is more powerful.}
+ \label{powerdiff}
 \end{figure}
@@
 @
- \label{samplesize}
  \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data.
 The scatter plot appears to universaly benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
+ \label{samplesize}
 \end{figure}
@@
 @
-  \label{time} 
   \caption{Histograms of log time taken to make a decision.
@@
 On average scatter plots are faster and the true plot is detected more often, with r-.}
+  \label{time} 
 \end{figure}
@@
 @
-  \label{confidence} 
   \caption{Histograms of reported confidence levels.
@@
 There was no statistically significant difference on confidence level between plot designs.}
+  \label{confidence} 
 \end{figure}
@@
 @
-  \label{confidence2}
   \caption{Reported confidence against true plot correlation.
@@
 Strong correlation is more often associated with a lower reported confidence.}
+  \label{confidence2}
 \end{figure}
@@
 
-  \label{pins}
   \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
@@
 Strongly correlated plot seem to be selected regardless of if the relationship is real or spurious.}
+  \label{pins}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{anom1}
-  \label{anom1}
   \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
@@
 Subjects most often cited the clustering effect shown as the reason for their selection, an example of the over-interpretation of randomness.}
+  \label{anom1}
 \end{figure}
@@
   \includegraphics[width=0.5\textwidth]{anom2}
-  \label{anom2}
   \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
@@
 When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.}
+  \label{anom2}
 \end{figure}
@@
  & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
 \end{eqnarray}
@@
 @
-  \label{powermodel}
   \caption{Predicted power fit from the mixed effects logit power model.
@@
 The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
+  \label{powermodel}
 \end{figure}
@@
 @
+  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
   \label{residuals}
-  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
 \end{figure}
@@
 \begin{figure}[htbp]
-\label{random}
   \centering
@@
 @
-    \caption{Histograms of estimated subject random effects, $\hat{\gamma}$}
+    \caption{Histograms of estimated subject random effects, $\hat{\gamma}$. They appear to be normally distributed with mean zero.}
+\label{random}
 \end{figure}
@@
 \centering
-\label{random2}
 <<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
@@
 The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
-\end{figure}
-
-\iffalse
-This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
-
-$$Power_{UR,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
-
-Testing if the scatter plot displays positive and negative correlation equally is equivalent to testing the hypothesis:
-
-\begin{itemize}
-  \item $H_{0}: \beta_{2} = \beta_{5} = 0$
-  \item $H_{1}:$ Either equality does not hold.
-\end{itemize}
-
-The restricted model is therefore:
-
-$$Power_{R,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i} + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
-
-The unrestricted model has a log-likelihood value of -346.9, where the restricted model has a log-likelihood of -355.8.
-The test statistic is therefore, with a null distributed as a $\chi^{2}_{2}$ variable.
-The null hypothesis is rejected (p = 0.02).
-We find that the scatter plot has asymmetric performance over negative and positive values of correlation.
-This may be a result of unwittingly introducing bias into the methodology, the first scatter plot shown to subjects in the instruction set always had strong positive correlation.
-A similar setup rejects the symmetric performance of the overlaid line graphs with a log likelihood ratio test statistic of 105.30 (p-value < 0.001).
-Such an extreme result is due to a systematic difference in the overlaid line graph and cannot be attributed simply to bias.
-\fi
-
-A linear mixed model is used to model the time to make a selection.
-
-\begin{eqnarray} 
-\label{timemodel}
-Time_{ijklmt} &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \beta_{13} x_{6,m} + \beta_{14} x_{5_l} \* x_{6,m} \gamma z_{s} + \epsilon_{ijklmt} \nonumber 
-\end{eqnarray}
-
-The structure of the model is similar to that in Equation~\ref{modelpower} except the link function is removed and a new variable, $x_{6,m}$ is introduced, with $m = \{0,1\}$ indicating if the subject detected the true plot.
-
-\begin{table}
+\label{random2}
+\end{figure}
+
+\begin{table}[htbp]
+\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
+\centering
 \label{tab2}
-\centering
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
-
 <<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
@@
 
+\iffalse
+This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
+
+$$Power_{UR,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
+
+Testing if the scatter plot displays positive and negative correlation equally is equivalent to testing the hypothesis:
+
+\begin{itemize}
+  \item $H_{0}: \beta_{2} = \beta_{5} = 0$
+  \item $H_{1}:$ Either equality does not hold.
+\end{itemize}
+
+The restricted model is therefore:
+
+$$Power_{R,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i} + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
+
+The unrestricted model has a log-likelihood value of -346.9, where the restricted model has a log-likelihood of -355.8.
+The test statistic is therefore, with a null distributed as a $\chi^{2}_{2}$ variable.
+The null hypothesis is rejected (p = 0.02).
+We find that the scatter plot has asymmetric performance over negative and positive values of correlation.
+This may be a result of unwittingly introducing bias into the methodology, the first scatter plot shown to subjects in the instruction set always had strong positive correlation.
+A similar setup rejects the symmetric performance of the overlaid line graphs with a log likelihood ratio test statistic of 105.30 (p-value < 0.001).
+Such an extreme result is due to a systematic difference in the overlaid line graph and cannot be attributed simply to bias.
+\fi
+
+A linear mixed model is used to model the time to make a selection.
+
+\begin{eqnarray} 
+\label{timemodel}
+Time_{ijklmt} &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
+ & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \beta_{13} x_{6,m} + \beta_{14} x_{5_l} \* x_{6,m} \gamma z_{s} + \epsilon_{ijklmt} \nonumber 
+\end{eqnarray}
+
+The structure of the model is similar to that in Equation~\ref{modelpower} except the link function is removed and a new variable, $x_{6,m}$ is introduced, with $m = \{0,1\}$ indicating if the subject detected the true plot.
+
 Table~\ref{tab2} shows a similar story for time as Table~\ref{tab1} did for power, that the overlaid lines have better performance close to zero correlation, but the scatter plot is much faster at high correlations.
@@
 
-\begin{table}
+\begin{table}[htbp]
 \label{tab3}
@@
 map.dat <- map_data("world")
-ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
+ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + 
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) + scale_colour_manual(values=c("#FF5500", "#00FFFF")) + 
+  theme_bw() + theme(aspect.ratio=0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
 @

### 2016-01-11 paper/paper.Rnw
Code cleanup

@@
 full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
-Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)
+prediction <- read.csv("sas/predfit.csv") #Large file, very slow
+time.fit <- read.csv("sas/time.csv")
+powercoef <- read.csv("sas/PowerEstimates.csv")
+
+colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "design", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
+
+full$subj_id <- factor(full$ip_address, labels=1:207)
+y <- data.frame(table(full$subj_id))
+keep <- y$Var1[y$Freq>9]
+
+full <- mutate(full, design = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"),
+               r = round(correlation, 1),
+               rsgn = ifelse(r >= 0, 1, 0),
+               rabs=abs(r),
+               alpha = log(0.05/0.95),
+               attempts = sapply(strsplit(as.character(full$response_0), ","), length), #count number of selections in response_0 column
+               weights = 1/attempts)
+full <- full %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
+#split multiple selections in response_0 into different rows
+full <- mutate(full, correct = ifelse(response_0 == actual_location, 1, 0),
+                   score = ifelse(correct==1, (20-attempts)/19, 0))
+full <- select(full, subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0, actual_location, pic_id, choice_reason, attempts, weights, alpha)
+
+full.sub <- filter(full, subj_id %in% keep) #keeping only >9 selections in actual modelling to better estimate random effects
+
+bylineup <- summarise(group_by(full.sub, pic_id, n, smoothed, design, r, actual_location), total=sum(score), correct=sum(correct), evals=sum(attempts*weights))
+bylineup <- mutate(bylineup, percent = total/evals)
+bylineup.sub <- summarise(group_by(full.sub, pic_id), total=sum(score), evals=sum(weights*attempts))
 
 gen_true_data <- function(n, r, dep=TRUE, smoothed = FALSE) {
-  d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2))) ##True plot, 0.5 for sigma off-diagonals requires ~300 iterations for all lineups.
+  d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2))) 
   if (dep) {
@@
 <<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
-Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
-               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
-#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
-ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
+ggplot(data=bylineup, aes(x=r, y=percent, colour=design)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
   scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
@@
 The dots are data points, the lines represent a naive model average fit.
-Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are particularly weak at revealing negative correlation.}
+Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at revealing negative correlation.}
     \label{results}
@@
 
-\begin{figure}[htbp]
-  \centering
-<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-Turk$correct <- Turk$percent*Turk$attempt
-Turk$r <- round(Turk$correlation, 1)
-#Turk.line <- subset(Turk, test_param=="line")
-#Turk.scatter <- subset(Turk, test_param=="scatter")
-prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "plot")]
-prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(plot=="Overlaid Lines", attempt, 0)),
-                                                  l.correct=sum(ifelse(plot=="Overlaid Lines", correct, 0)),
-                                                  s.attempt=sum(ifelse(plot=="Scatterplot", attempt, 0)),
-                                                  s.correct=sum(ifelse(plot=="Scatterplot", correct, 0)))
-#Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
-#Turk.line <- summarise(group_by(Turk.line, smoothed, n, r, test_param), sum(attempt), sum(correct))
-#Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
-#Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
-#colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
-#colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
-#prop <- cbind(Turk.line, Turk.scatter[4:5])
+\iffalse
+\begin{figure}[htbp]
+  \centering
+<<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "design")]
+prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", attempt, 0)),
+                                                  l.correct=sum(ifelse(design=="Overlaid Lines", correct, 0)),
+                                                  s.attempt=sum(ifelse(design=="Scatterplot", attempt, 0)),
+                                                  s.correct=sum(ifelse(design=="Scatterplot", correct, 0)))
 prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
@@
 \end{figure}
-
-\begin{figure}[htbp]
-  \centering
-<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
+\fi
+
+\begin{figure}[htbp]
+  \centering
+<<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
 @
  \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data.
-The scatter plot appears to universaly benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
+The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
  \label{samplesize}
@@
 <<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-full <- mutate(full, sign = ifelse(correlation >= 0, 1, 0), n24 = ifelse(n ==24, 1, 0), n48 = ifelse(n ==48,  1, 0), n96 = ifelse(n ==96, 1, 0), plot = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"))
-time <- data.frame(plot=full$plot, time=full$log_time, correct=full$correct, rsgn=full$sign)
-#ggplot(data=time, aes(x=time, fill=factor(correct))) +geom_bar(stat="bin") + facet_grid(~plot) +labs(x="Log(Time) Taken", y=element_blank()) + scale_fill_discrete(name = "Correct")
+time <- select(full, design, log_time, correct,, sign)
 time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
-tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
-t1 <- ggplot(data=subset(time, correct==0), aes(x=time)) + geom_histogram(binwidth=0.5) + 
+tm <- summarise(group_by(time, correct, design, rsgn), m=mean(log_time))
+t1 <- ggplot(data=subset(time, correct==0), aes(x=log_time)) + geom_histogram(binwidth=0.5) + 
   geom_vline(data=subset(tm, correct==0), aes(xintercept = m), color="red") +
-  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
-t2 <- ggplot(data=subset(time, correct==1), aes(x=time)) + geom_histogram(binwidth=0.5) + 
+  facet_grid(design~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
+t2 <- ggplot(data=subset(time, correct==1), aes(x=log_time)) + geom_histogram(binwidth=0.5) + 
   geom_vline(data=subset(tm, correct==1), aes(xintercept = m), color="red") +
-  facet_grid(plot~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
+  facet_grid(design~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
 grid.arrange(t1, t2, ncol=2)
@@
   \centering
-<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-conf <- data.frame(plot=full$plot, conf=full$conf_level, correct=full$correct, rsgn=full$sign)
+<<confidence, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+conf <- select(full, design, conf_level, correct, sign)
 conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
-cm <- summarise(group_by(conf, correct, plot, rsgn), m=mean(conf))
-c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf)) + geom_histogram(binwidth=1) + 
+cm <- summarise(group_by(conf, correct, design, rsgn), m=mean(conf_level))
+c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf_level)) + geom_histogram(binwidth=1) + 
   geom_vline(data=subset(cm, correct==0), aes(xintercept = m), color="red") +
-  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
-c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf)) + geom_histogram(binwidth=1) + 
+  facet_grid(design~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
+c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf_level)) + geom_histogram(binwidth=1) + 
   geom_vline(data=subset(cm, correct==1), aes(xintercept = m), color="red") +
-  facet_grid(plot~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
+  facet_grid(design~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
 grid.arrange(c1, c2, ncol=2)
@@
   \centering
-<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-full$subj_id <- factor(full$ip_address, labels=1:207)
-y <- data.frame(table(full$subj_id))
-keep2 <- y$Var1[y$Freq>9]
-conf2 <- filter(full, subj_id %in% keep2)
-conf2 <- select(conf2, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
-#sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
-#sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
-#sub <- mutate(sub, prop = correct/Freq)
+<<confidence2, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+conf2 <- select(full, conf_level, correlation, smoothed, n, design) #c(7, 15
 qplot(correlation, 6-conf_level, data=conf2, alpha=I(0.1)) + 
   geom_smooth(method="loess") + 
-  facet_grid(n~plot+smoothed) +
+  facet_grid(n~design+smoothed) +
   labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
@@
   \centering
-<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-# correl <- read.csv("correlations.csv") # You have already read this
+<<pins, dependson='setup', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 count.setup <- select(full, pic_id, response_0, actual_location, 
-              test_param, smoothed, n)
-count.setup$response_0 <- as.character(count.setup$response_0) #full[,c(5, 6, 10, 15, 16, 17)]
-# Need to use the tidyr commands instead of cSplit
-# Here's where we will need to make some changes to incorporate the multiple picks
-#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
-#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
-count.setup$choice <- separate(count.setup, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
-count.setup$choice[str_detect(count.setup$actual_location, count.setup$response_0)] <- 
-  count.setup$actual_location[str_detect(count.setup$actual_location, count.setup$response_0)]
-                         #as.integer(pin$response_0)
-counts <- data.frame(table(count.setup$pic_id, count.setup$choice))
+              design, smoothed, n)
+count.setup$response_0 <- as.character(count.setup$response_0)
+counts <- data.frame(table(count.setup$pic_id, count.setup$response_0))
 counts$Var1 <- as.numeric(as.character(counts$Var1))
 counts$Var2 <- as.numeric(as.character(counts$Var2))
-colnames(counts) <- c("lineup", "plots", "count")
-pin <- merge(counts, Turk[-17], by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plots, count, smoothed, true_pos, n, test_param, r, attempt)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
+colnames(counts) <- c("lineup", "plot", "count")
+pin <- merge(counts, bylineup, by.x="lineup", by.y="pic_id")
+pin <- select(pin, lineup, plot, count, smoothed, actual_location, n, design, r, evals)
 pin$r <- factor(pin$r)
-colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r", "attempt")
-pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
 pin$base <- 0
 pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
-pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"), percent = count/attempt)
-#pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
+pin <- mutate(pin, true = ifelse(plot==actual_location, "True plot", "Null plot"), percent = count/evals)
 pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
@@
 @
-
   \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
@@
 Default plot design is overlaid lines.
-The model is fitted using the GLIMMIX procedure in SAS.} %Sign = 1 if correlation > 0, 0 otherwise.
-Each plot design is estimated separately.
-Results must be transformed via the logit link function to generate a power prediction.
-Subjects with less than ten lineup evaluations were removed from the sample.
-
+The model is fitted using the GLIMMIX procedure in SAS.}
 \label{tab1}
@@
 <<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-full$subj_id <- factor(full$ip_address, labels=1:207)
-
-# Need to keep only evaluations done by subject who completed at least 10
-x <- data.frame(table(full$subj_id))
-keep <- x$Var1[x$Freq>9]
-full.subs <- filter(full, subj_id %in% keep)
-full.subs$r <- round(full.subs$correlation, 1)
-full.subs$rsgn <- ifelse(full.subs$r >= 0, 1, 0)
-full.subs$rabs <- abs(full.subs$r)
-full.subs$plot <- ifelse(full.subs$test_param == 0, "S", "L")
-full.subs <- select(full.subs, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location, pic_id, choice_reason)#[,c(1,7, 15,16,17,20:26)]
-full.subs$attempts <- sapply(strsplit(as.character(full.subs$response_0), ","), length)
-full.subs$weights = 1/(full.subs$attempts) #weights are actually number of attempts in the glm
-full.sub <- full.subs %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
-full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
-full.sub$alpha <- log(.05/0.95)
-full.sub$subj_id <- as.numeric(full.sub$subj_id)
-subsetted <- select(full.sub, rabs, n, alpha, rsgn, smoothed, plot, subj_id, weights, correct, log_time)
-subsetted$rabs <- as.numeric(subsetted$rabs)
-subsetted$subj_id <- as.numeric(subsetted$subj_id)
-subsetted <- mutate(subsetted, plot = ifelse(plot=="S", 1, 0))
+sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
+sasdata$rabs <- as.numeric(sasdata$rabs)
+sasdata$subj_id <- as.numeric(sasdata$subj_id)
+sasadata <- mutate(sasdata, design = ifelse(design=="Scatterplot", 1, 0))
 #write.csv(subsetted, "sas/choices.csv", row.names = FALSE)
-
-powercoef <- read.csv("sas/PowerEstimates.csv")
 powercoef <- powercoef[,-1]
@@
 kable(powercoef, format="latex", digits=2)
-#kable(powercoef[7:12,], format="latex", digits=2)
 @
@@
  \centering
-<<powermodel, dependson='powerfit', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+<<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 # data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
-#                                      smoothed=c(0,1), plot=c(0,1),
+#                                      smoothed=c(0,1), design=c(0,1),
 #                                      subj_id=full.sub$subj_id))
@@
 #Code exported to sas, see sas/model\ fit.sas
-prediction <- read.csv("sas/predfit.csv") #Large file, very slow
-colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "plot", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
-glmm.resid <- subset(prediction, detected==0 | detected==1)
-prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), plot = ifelse(plot==0, "Overlaid Lines", "Scatterplot"), plot = ifelse(smoothed==1, paste(plot, "(Smoothed)", sep=" "), plot),  diff = round(RandomEf - FixedEf,5))
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), design = ifelse(design==0, "Overlaid Lines", "Scatterplot"), design = ifelse(smoothed==1, paste(design, "(Smoothed)", sep=" "), design),  diff = round(RandomEf - FixedEf,5))
 randomeffects <- summarise(group_by(prediction, subj_id), re=max(diff)) #all values of diff are identical for a given subject, max chosen just as a function to select one of them
-prediction <- select(prediction, r, t, rsgn, smoothed, plot, subj_id, RandomEf, FixedEf)
+prediction <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
 prediction <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
 
-qplot(r, RandomEf, data=prediction, color=plot, 
-      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
+qplot(r, RandomEf, data=prediction, color=design, 
+      group=interaction(subj_id, design), geom="line", alpha=I(0.1)) + 
   scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
@@
   \centering
-<<residuals, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+<<residuals, dependson='setup', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+glmm.resid <- subset(prediction, detected==0 | detected==1)
 ggplot(data=glmm.resid) + geom_histogram(aes(y=..density.., x=residual), binwidth=2) + labs(x=NULL, y=NULL)
@@
   \centering
-<<random, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+<<random, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
@@
 \centering
-<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-turk.correct <- summarise(group_by(full.sub, pic_id), correct=sum(correct), attempts=sum(weights*attempts))
-turk.correct$percent <- turk.correct$correct/turk.correct$attempts
-ability <- merge(full.sub[,c("subj_id", "pic_id", "correct")], turk.correct[c("pic_id", "percent")], by="pic_id")
-ability <- summarise(group_by(ability, subj_id), actual=sum(correct), estimated=sum(percent))
+<<random2, dependson='setup', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ability <- merge(full.sub[,c("subj_id", "pic_id", "correct", "score")], bylineup[c("pic_id", "percent")], by="pic_id")
+ability <- summarise(group_by(ability, subj_id), actual=sum(score), estimated=sum(percent))
 ability$diff <- ability$actual - ability$estimated
@@
 \end{figure}
-
-\begin{table}[htbp]
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
-\centering
-\label{tab2}
-<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
-#timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
-#                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
-#time.fit <- summary(timelmer)$coefficients
-#pv <- function(x){
-#  p <- 2*(1-pt(x, 1835))
-#}
-#p.value <- sapply(time.fit[,3], pv)
-#time.fit <- cbind(time.fit, p.value)
-time.fit <- read.csv("sas/time.csv")
-time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
-rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "Plot", "Plot*Smoothed", "Plot*t", "Plot*abs(r)", "Plot*sign(r)", "Plot*abs(r)*sign(r)", "Plot*detected")
-time.fit$Effect <- NULL
-colnames(time.fit) <- c("Estimate", "standard error", "df", "t-value", "p-value")
-kable(time.fit[1:14,], format="latex", digits=2)
-#kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
-@
-\end{table}
 
@@
 The structure of the model is similar to that in Equation~\ref{modelpower} except the link function is removed and a new variable, $x_{6,m}$ is introduced, with $m = \{0,1\}$ indicating if the subject detected the true plot.
+
+\begin{table}[htp]
+\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
+\label{tab2}
+\centering
+<<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
+rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "design", "design*Smoothed", "design*t", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*detected")
+time.fit$Effect <- NULL
+colnames(time.fit) <- c("Estimate", "standard error", "df", "t-value", "p-value")
+kable(time.fit[1:14,], format="latex", digits=2)
+@
+\end{table}
 
@@
 \centering
-<<answertables, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
-turk.tbl <- select(Turk, test_param, smoothed, r, n, correct, attempt)
-turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$attempt, sep="/")
+<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+turk.tbl <- select(bylineup, design, smoothed, r, n, correct, evals)
+turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$evals, sep="/")
 turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
-scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="scatter")$result, ncol=8))
-scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="scatter")$result, ncol=8))
-line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="line")$result, ncol=8))
-line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="line")$result, ncol=8))
+scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Scatterplot")$result, ncol=8))
+scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Scatterplot")$result, ncol=8))
+line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Overlaid Lines")$result, ncol=8))
+line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Overlaid Lines")$result, ncol=8))
 colnames(scatter.uns) <- c("", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
@@
 
-Table~\ref{tab3} contains the true plot detections / attempts for each lineup in the data collection. True plot correlation is on the x axis and t is on the y axis.
-
-\begin{figure}[htbp]
-\label{map}
-\centering
-<<map, , dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+Table~\ref{tab3} contains the true plot detections / attempts for each lineup in the data collection.
+Results ignore multiple selection weighting.
+True plot correlation is on the x axis and t is on the y axis.
+
+\begin{figure}[htbp]
+\centering
+<<map, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
 location <- read.csv("location.csv")
 map.dat <- map_data("world")
-ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + 
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) + scale_colour_manual(values=c("#FF5500", "#00FFFF")) + 
-  theme_bw() + theme(aspect.ratio=0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
-@
+ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
+@
+\label{map}
 \end{figure}

### 2016-01-11 paper/paper.Rnw
Updates

@@
 \centering
-<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
 sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
@@
 \centering
-<<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+<<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
+<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
@@
 \centering
-<<map, dependson='setup', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 location <- read.csv("location.csv")

### 2016-01-11 paper/paper.Rnw
Tweaks

@@
 colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "design", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), design = ifelse(design==0, "Overlaid Lines", "Scatterplot"), design = ifelse(smoothed==1, paste(design, "(Smoothed)", sep=" "), design),  diff = round(RandomEf - FixedEf,5))
 
@@
                    score = ifelse(correct==1, (20-attempts)/19, 0))
-full <- select(full, subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0, actual_location, pic_id, choice_reason, attempts, weights, alpha)
+full <- select(full, subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0, actual_location, pic_id, choice_reason, attempts, weights, alpha, conf_level)
 
@@
 <<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "design")]
-prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", attempt, 0)),
-                                                  l.correct=sum(ifelse(design=="Overlaid Lines", correct, 0)),
-                                                  s.attempt=sum(ifelse(design=="Scatterplot", attempt, 0)),
-                                                  s.correct=sum(ifelse(design=="Scatterplot", correct, 0)))
-prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
-               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
-               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
-               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
-ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
+#prop <- bylineup[order(bylineup$r, bylineup$n, bylineup$smoothed),c("smoothed", "n", "evals", "total", "r", #"design")]
+#prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", total, 0)),
+#                                                  l.correct=sum(ifelse(design=="Overlaid Lines", score, 0)),
+#                                                  s.attempt=sum(ifelse(design=="Scatterplot", total, 0)),
+#                                                  s.correct=sum(ifelse(design=="Scatterplot", score, 0)))
+#prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
+#               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
+#               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
+##               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
+#ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
+#        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
 @
@@
 <<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-time <- select(full, design, log_time, correct,, sign)
+time <- select(full, design, log_time, correct, rsgn)
 time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
@@
 <<confidence, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-conf <- select(full, design, conf_level, correct, sign)
+conf <- select(full, design, conf_level, correct, rsgn)
 conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
@@
 <<confidence2, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-conf2 <- select(full, conf_level, correlation, smoothed, n, design) #c(7, 15
-qplot(correlation, 6-conf_level, data=conf2, alpha=I(0.1)) + 
+conf2 <- select(full, conf_level, r, smoothed, n, design) #c(7, 15
+qplot(r, 6-conf_level, data=conf2, alpha=I(0.1)) + 
   geom_smooth(method="loess") + 
@@
 
-\begin{enumerate}
+\begin{itemize}
 \item $x_{1,i}$ is the absolute value of the true plot correlation, $\rho$, with $i \in \{0.3, 0.5, 0.7, 0.9\}$,
@@
 \item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 157$.
-\end{enumerate}
+\end{itemize}
 
@@
 
-Where a subject made multiple selections from the same lineup, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
+If a subject made multiple selections from the same lineup, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
 
@@
 #Code exported to sas, see sas/model\ fit.sas
-prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), design = ifelse(design==0, "Overlaid Lines", "Scatterplot"), design = ifelse(smoothed==1, paste(design, "(Smoothed)", sep=" "), design),  diff = round(RandomEf - FixedEf,5))
-randomeffects <- summarise(group_by(prediction, subj_id), re=max(diff)) #all values of diff are identical for a given subject, max chosen just as a function to select one of them
-prediction <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
-prediction <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
-
-qplot(r, RandomEf, data=prediction, color=design, 
+#all values of diff are identical for a given subject, max chosen just as a function to select one of them
+predplot <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
+predplot <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
+
+qplot(r, RandomEf, data=predplot, color=design, 
       group=interaction(subj_id, design), geom="line", alpha=I(0.1)) + 
@@
 <<random, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+randomeffects <- summarise(group_by(prediction, subj_id), re=max(diff))
 ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
@@
 
-The structure of the model is similar to that in Equation~\ref{modelpower} except the link function is removed and a new variable, $x_{6,m}$ is introduced, with $m = \{0,1\}$ indicating if the subject detected the true plot.
+The structure of the model ~\ref{timemodel} is similar to that in Equation~\ref{modelpower} except the link function is replace with the identity function, $\psi(x) = x$, and a new variable, $x_{6,m}$, is introduced with $m = \{0,1\}$ indicating if the subject detected the true plot, where $m = 0$ is a failure to detect the true plot.
 
@@
 <<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
-<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
@@
 \begin{table}[htbp]
+\caption{The true plot detections / attempts for each lineup in the data collection. Results ignore multiple selection weighting. True plot correlation is on the x axis and t is on the y axis. }
 \label{tab3}
@@
 
-Table~\ref{tab3} contains the true plot detections / attempts for each lineup in the data collection.
-Results ignore multiple selection weighting.
-True plot correlation is on the x axis and t is on the y axis.
-
-\begin{figure}[htbp]
-\centering
-<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+\begin{figure}[htbp]
+\centering
+<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 location <- read.csv("location.csv")
@@
 ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
-@
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
+@
+\caption{The location of the recruited subjects. The majority reside in the United States but some are spread around other countries in Europe and Asia.}
 \label{map}
@@
 
-Figure~\ref{map} shows the location of the recruited subjects. The majority reside in the United States but some are spread around other countries in Europe and Asia.
-
 \end{document}

### 2016-01-11 paper/paper.Rnw
tweaks

@@
 powercoef <- powercoef[,-1]
-rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "plot", "plot*abs(r)", "plot*sign(r)", "plot*abs(r)*sign(r)", "plot*smoothed", "plot*t")
+rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "design", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*smoothed", "design*t")
 colnames(powercoef) <- c("Estimate", "standard error", "df", "t-value", "p-value")
@@
 Table \ref{tab1} summarises the model estimates.
-It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overliad lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with $abs(\rho)$ and $sign(\rho)$.
-$abs(\rho)$ models the power difference between low and high correlations, and $abs(\rho)*sign(\rho)$ models an increase in power appearing in positive but not negative correlation.
-Taking these into account, the scatter plot shows superior performance and higher values of abs($\rho$), It needs to be restated that we only tested correlations at  $0.3, 0.5, 0.7$ and $0.9$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
+It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overlaid lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with $\rho$.
+$\beta_{2}$ and $\beta_{8}$ model the power difference between low and high correlations, while $\beta_{3}, \beta_{4}, beta_{9}$ and $\beta_{10}$ model any increase in power appearing only in positive correlation.
+Taking these into account, the scatter plot shows superior performance relative to the overlaid lines at higher values of abs($\rho$).
+It needs to be restated that we only tested correlations at  $0.3, 0.5, 0.7$ and $0.9$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
 

### 2016-01-12 paper/paper.Rnw
Cleanup

@@
 \usepackage{afterpage}
+\usepackage{bbm}
 
@@
 library(knitr)
-# Convert reshape2 to tidyr, melt to gather - done
-#library(reshape2)
 library(nullabor)
@@
 library(stringr)
-#library(ipapi)
 library(maps)
-library(moments)
-#library(pander)
-#library(splitstackshape) - changed cSplit to unnest
 #library(readr)
@@
 # Convert these to read_csv at some point
-correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
-full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
-prediction <- read.csv("sas/predfit.csv") #Large file, very slow
-time.fit <- read.csv("sas/time.csv")
-powercoef <- read.csv("sas/PowerEstimates.csv")
-
-colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "design", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
-prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs), design = ifelse(design==0, "Overlaid Lines", "Scatterplot"), design = ifelse(smoothed==1, paste(design, "(Smoothed)", sep=" "), design),  diff = round(RandomEf - FixedEf,5))
+correl <- read.csv("correlations.csv", stringsAsFactors = FALSE) ##Correlations of every plot in every lineup
+full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
+prediction <- read.csv("sas/predfit.csv") ##Large file, very slow, fitted data from SAS
+time.fit <- read.csv("sas/time.csv") ##coefficients for time model from SAS
+powercoef <- read.csv("sas/PowerEstimates.csv") ##coefficients for power model from SAS
 
@@
 keep <- y$Var1[y$Freq>9]
-
-full <- mutate(full, design = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"),
+##Recording who evaluated at least ten lineups. Done prior to splitting multiple selections
+
+full <- full %>% mutate(design = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"),
                          r = round(correlation, 1),
                          rsgn = ifelse(r >= 0, 1, 0),
-               rabs=abs(r),
+                         rabs = abs(r),
                          alpha = log(0.05/0.95),
-               attempts = sapply(strsplit(as.character(full$response_0), ","), length), #count number of selections in response_0 column
-               weights = 1/attempts)
-full <- full %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
-#split multiple selections in response_0 into different rows
-full <- mutate(full, correct = ifelse(response_0 == actual_location, 1, 0),
-                   score = ifelse(correct==1, (20-attempts)/19, 0))
-full <- select(full, subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0, actual_location, pic_id, choice_reason, attempts, weights, alpha, conf_level)
-
-full.sub <- filter(full, subj_id %in% keep) #keeping only >9 selections in actual modelling to better estimate random effects
-
-bylineup <- summarise(group_by(full.sub, pic_id, n, smoothed, design, r, actual_location), total=sum(score), correct=sum(correct), evals=sum(attempts*weights))
+                         attempts = sapply(strsplit(as.character(full$response_0), ","), length), 
+                         ##count number of selections in response_0 column
+                         weights = 1/attempts) %>%
+  transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0) %>%
+  ##split multiple selections in response_0 into different rows
+  mutate(correct = ifelse(response_0 == actual_location, 1, 0), 
+         score = ifelse(correct==1, (20-attempts)/19, 0)) %>%
+  select(subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0,
+         actual_location, pic_id, choice_reason, attempts, weights, alpha, conf_level)
+
+full.sub <- filter(full, subj_id %in% keep) 
+##keeping only subjects with >0 evaluations in actual modelling to better estimate random effects
+
+bylineup <- summarise(group_by(full, pic_id, n, smoothed, design, r, actual_location), total=sum(score), correct=sum(correct), evals=sum(weights))
 bylineup <- mutate(bylineup, percent = total/evals)
-bylineup.sub <- summarise(group_by(full.sub, pic_id), total=sum(score), evals=sum(weights*attempts))
-
-gen_true_data <- function(n, r, dep=TRUE, smoothed = FALSE) {
+bylineup.sub <- summarise(group_by(full.sub, pic_id), total=sum(score), evals=sum(weights))
+bylineup.sub <- mutate(bylineup.sub, percent = total/evals)
+
+##Summarise data by lineup
+
+#sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
+#sasadata <- mutate(sasdata, rabs = as.numeric(rabs),
+#                   subj_id = as.numeric(subj_id)
+#                   design = ifelse(design=="Scatterplot", 1, 0))
+#write.csv(subsetted, "sas/sasdata.csv", row.names = FALSE)
+##The code that is exported to fit the model into SAS, see sas/model fit.sas
+
+# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
+#                                      smoothed=c(0,1), design=c(0,1),
+#                                      subj_id=full.sub$subj_id))
+#pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)
+#write.csv(pred, "sas/predgrid.csv", row.names=FALSE) # these lines are done, but its a large file so commented out for knitr
+#Code exported to SAS to predict new values
+
+colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "design", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
+prediction <- mutate(prediction, r = ifelse(rsgn==1, rabs, -rabs),
+                     design = ifelse(design==0, "Overlaid Lines", "Scatterplot"), 
+                     design = ifelse(smoothed==1, paste0(design, " (Smoothed)"), design),  
+                     diff = round(RandomEf - FixedEf,5))
+##Data from SAS
+
+gen_true_data <- function(n, r, smoothed = FALSE) {
   d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2))) 
-  if (dep) {
   d$X1 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X1))
   d$X2 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X2))
-  }
   if (smoothed) {
@@
 }
-
-gen_null <- function(n, m=20, dep=TRUE, smoothed = FALSE){
+##Generates a pair of data, length n, correlation r
+
+gen_null <- function(n, m=20, smoothed = FALSE){
   nd <- NULL
   for(i in 1:(m-1)) {
-    d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, 0, 0, 1), 2, 2)))
-    if (dep) {
-      d$X1 <- as.vector(arima.sim(list(ar=n/100), n))
-      d$X2 <- as.vector(arima.sim(list(ar=n/100), n))
-    }
+    d <- data.frame(X1 = as.vector(arima.sim(list(ar=n/100), n)),
+                    X2 = as.vector(arima.sim(list(ar=n/100), n)))
     if (smoothed) {
@@
 }
+##Generates m-1 pairs of uncorrelated data, length n
 @
@@
 
-Where $\hat\pi_{i} = (x_{i}+1)/(n_{i}+1)$, where $x$ is the number of times a true lineup was correctly identified and $n$ is the Welch-Satterwaite estimate for the degrees of freedom.
+Where $\hat\pi_{i} = (x_{i}+1)/(n_{i}+1)$, $x$ is the number of times a true lineup was correctly identified and $n$ is the Welch-Satterwaite estimate for the degrees of freedom.
 
@@
 The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$.
-True data pairs are generated with $\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$.
+True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$.
 Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$.
@@
 The null pairs often had correlation generated spuriously through the simulation process.
-This was particularly problematic for generating small sample data, but setting $\beta = max(0.5,100t)$ for both true and null data reduced spurious correlation to manageable levels.
+This was particularly problematic for generating small sample data, but setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
 For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications.
@@
 
-Subjects were allowed to make multiple selections, which is handled by probability weighting in the model. This increases our data from 1684 to 1849 evaluations.
+Subjects were allowed to make multiple selections from each lineup evaluated, additional selections on lineup $i$ by subject $j$ are penalised according to a score defined by:
+\begin{equation}
+\label{score}
+{score}_{ij} =\frac{m-s_{ij}}{m-1} \mathbbm{I}_{ij}
+\end{equation}
+
+where $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise, $s_{ij}$ is the number of selections made and $m = 20$ is the number of plots in the lineup.
+
+However, for the model fit to power, multiple selections are seperated and given a weighting equal to to inverse of the number of selections made in that lineup evaluation.
+This increases our data from 1684 to 1849 evaluations.
 
@@
 <<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-#prop <- bylineup[order(bylineup$r, bylineup$n, bylineup$smoothed),c("smoothed", "n", "evals", "total", "r", #"design")]
-#prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", total, 0)),
-#                                                  l.correct=sum(ifelse(design=="Overlaid Lines", score, 0)),
-#                                                  s.attempt=sum(ifelse(design=="Scatterplot", total, 0)),
-#                                                  s.correct=sum(ifelse(design=="Scatterplot", score, 0)))
-#prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
-#               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
-#               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
-##               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
-#ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-#        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
+prop <- select(bylineup, smoothed, n, evals, total, r, design)
+prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", evals, 0)),
+                                                  l.correct=sum(ifelse(design=="Overlaid Lines", total, 0)),
+                                                  s.attempt=sum(ifelse(design=="Scatterplot", evals, 0)),
+                                                  s.correct=sum(ifelse(design=="Scatterplot", total, 0)))
+prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
+               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
+               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
+               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
+ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
+        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
 @
@@
 Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
-Figure~\ref{time} and Figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
+Figure~\ref{time} and figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
 Figure~\ref{time} is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
@@
 The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis.
-The correlation of the simulated (true) data ($\rho = \pm \{0.3, 0.5, 0.7, 0.9\}$ in the plot, the sample size ($t= \{12, 24, 48, 96\}$), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
+The correlation of the simulated (true) data ($\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$ in the plot, the sample size ($t \in \{12, 24, 48, 96\}$), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
 The blue pins indicate the true plots and the red pins indicate the null plots.
@@
 <<pins, dependson='setup', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-count.setup <- select(full, pic_id, response_0, actual_location, 
-              design, smoothed, n)
+count.setup <- select(full, pic_id, response_0, actual_location, design, smoothed, n)
 count.setup$response_0 <- as.character(count.setup$response_0)
@@
 pin <- select(pin, lineup, plot, count, smoothed, actual_location, n, design, r, evals)
-pin$r <- factor(pin$r)
-pin$base <- 0
-pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
-pin <- mutate(pin, true = ifelse(plot==actual_location, "True plot", "Null plot"), percent = count/evals)
-pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
-       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
-  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
-  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
-  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
+pin <- mutate(pin, r = factor(r),
+              true = ifelse(plot==actual_location, "True plot", "Null plot"), 
+              percent = count/evals,
+              base = 0,
+              correlation = diag(as.matrix(correl[pin$lineup, (pin$plot+2)])))
+pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
+         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + 
+         labs(title ="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
+pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
+         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
+         labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
+pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
+         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) +
+         labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
+pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + 
+         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
+         labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
 grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
@@
   \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
-On the x-axis is the correlation of the plot (there is potentially spurious correlation in the null plots), and on the y-axis is the proportion a plot was chosen.
-Each plot contains data from the three lineups with the requisite characteristics.
+On the x-axis is the correlation of the plot (there typically is spurious correlation in the null plots), and on the y-axis is the proportion of times a plot was chosen.
+Each plot contains data from the three lineups with the requisite factor combination.
 Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins.
@@
 
-\subsection{Modelling Subject Responses}
+\subsection{Modelling Graphical Power}
 
@@
 <<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
-sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
-sasdata$rabs <- as.numeric(sasdata$rabs)
-sasdata$subj_id <- as.numeric(sasdata$subj_id)
-sasadata <- mutate(sasdata, design = ifelse(design=="Scatterplot", 1, 0))
-#write.csv(subsetted, "sas/choices.csv", row.names = FALSE)
 powercoef <- powercoef[,-1]
 rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "design", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*smoothed", "design*t")
-colnames(powercoef) <- c("Estimate", "standard error", "df", "t-value", "p-value")
+colnames(powercoef) <- c("Estimate", "Standard Error", "df", "t-value", "p-value")
 kable(powercoef, format="latex", digits=2)
@@
 It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overlaid lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with $\rho$.
-$\beta_{2}$ and $\beta_{8}$ model the power difference between low and high correlations, while $\beta_{3}, \beta_{4}, beta_{9}$ and $\beta_{10}$ model any increase in power appearing only in positive correlation.
+$\beta_{2}$ and $\beta_{8}$ model the power difference between low and high correlations, while $\beta_{3}, \beta_{4}, \beta_{9}$ and $\beta_{10}$ model any increase in power appearing only in positive correlation.
 Taking these into account, the scatter plot shows superior performance relative to the overlaid lines at higher values of abs($\rho$).
@@
 %The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots(...), it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (Figure~\ref{smoothing}.
+Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (Figure~\ref{smoothing}.
 Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
@@
 <<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
-#                                      smoothed=c(0,1), design=c(0,1),
-#                                      subj_id=full.sub$subj_id))
-#pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)
-#write.csv(pred, "sas/predgrid.csv", row.names=FALSE) # these lines are done, but its a large file so commented out for knitr
-#Code exported to sas, see sas/model\ fit.sas
-#all values of diff are identical for a given subject, max chosen just as a function to select one of them
 predplot <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
-predplot <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
+predplot <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), 
+                   FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
+##Logit link function transform
 
@@
 In Figure~\ref{random}, the level two residuals, $\hat{\gamma}$, generated appear to be normally distributed validating the assumptions of the modelling process.
-For each lineup, we rate the difficulty by the proportion of true plot detections to evaluations.
-Then, for each subject the difficulty of each lineup they evaluated is summed to give a measure of the expected number of detections.
-The visual ability of a subject is defined as the difference between the actual and expected number of detections, and is compared to the random effects in Figure~\ref{random2} to further validate the model.
+
+The scores of each subject given by equation~\ref{score} can be summed to give a total score per subject $j$:
+
+$${score}_{j} = \sum_{i}\{{score}_{ij} | i \in L_{j}\}$$
+
+where $L_{j}$ is the set of lineups evaluated by subject $j$.
+
+The lineup $i$ with $K$ evaluations has an expected score:
+
+$${E[score]}_{i} = \frac{1}{K}\sum_{j}{score}_{ij}$$
+
+summing these per subject gives an expected total score per subject:
+
+$${E[score]}_{j} = \sum_{i}\{{E[score]}_{i} | i \in L_{j}\}$$
+
+Finally, we measure visual ability by:
+
+$${Visual Ability}_{j} = {score}_{j}-{E[score]}_{j}$$.
+
+Comparing these to the estimated random effects $\hat\gamma$ in Figure~\ref{random2} shows that there is a strong correlation between visual ability and the random effects, further validating the model.
 
@@
 <<random2, dependson='setup', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-ability <- merge(full.sub[,c("subj_id", "pic_id", "correct", "score")], bylineup[c("pic_id", "percent")], by="pic_id")
-ability <- summarise(group_by(ability, subj_id), actual=sum(score), estimated=sum(percent))
+ability <- merge(full.sub[,c("subj_id", "pic_id", "score", "weights")], bylineup.sub[c("pic_id", "percent")], by="pic_id")
+ability <- summarise(group_by(ability, subj_id), actual=sum(score), estimated=sum(percent*weights))
 ability$diff <- ability$actual - ability$estimated
@@
 
-A linear mixed model is used to model the time to make a selection.
+\subsection{Modelling time taken}
+
+The time taken to evaluate a lineup is recorded and log transformed to reduce the impact of several large ($> 500$ seconds) outliers.
 
@@
 \label{timemodel}
-Time_{ijklmt} &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
+ln(time_{ijklmt}) &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
  & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
@@
 <<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
-time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
+time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] 
+##separate into main effects then interaction effects in same order
 rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "design", "design*Smoothed", "design*t", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*detected")
@@
 
-\begin{table}[htbp]
-\caption{The true plot detections / attempts for each lineup in the data collection. Results ignore multiple selection weighting. True plot correlation is on the x axis and t is on the y axis. }
+\begin{table}[htp]
+\caption{The true plot detections / attempts for each lineup in the data collection. Results ignore multiple selection weighting. True plot correlation is on the x axis and t is on the y axis.}
 \label{tab3}
@@
 line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Overlaid Lines")$result, ncol=8))
-colnames(scatter.uns) <- c("", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(scatter.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.uns) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-kable(scatter.uns, format="latex", caption="Unsmoothed scatter plot results")
-kable(scatter.sm, format="latex", caption="Smoothed scatter plot results")
-kable(line.uns, format="latex", caption="Unsmoothed overlaid line results")
-kable(line.sm, format="latex", caption="Smoothed overlaid line results")
+colnames(scatter.uns) <- c("Unsmoothed Scatterplot", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(scatter.sm) <- c("Smoothed Scatterplot","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.uns) <- c("Unsmoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.sm) <- c("Smoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+kable(scatter.uns, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables2, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(scatter.sm, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables3, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(line.uns, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables4, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(line.sm, format="latex")
 @
@@
 @
-\caption{The location of the recruited subjects. The majority reside in the United States but some are spread around other countries in Europe and Asia.}
+\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
 \label{map}

### 2016-01-12 paper/Extra Material.R

@@ -59,16 +59,7 @@ ability$diff <- ability$actual - ability$estimated
 ability.re <- merge(ability, randomeffects, by="subj_id")
 ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point()
 
-#Results Tables#
-turk.tbl <- select(Turk, test_param, smoothed, r, n, correct, attempt)
-turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$attempt, sep="/")
-turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
-scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="scatter")$result, ncol=8))
-scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="scatter")$result, ncol=8))
-line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & test_param=="line")$result, ncol=8))
-line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & test_param=="line")$result, ncol=8))
-colnames(scatter.uns) <- c("", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(scatter.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.uns) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.sm) <- c("","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-
+#Attempts histogram#
+attempts <- summarise(group_by(full.sub, subj_id, pic_id), attempts = max(attempts))
+ggplot(data=attempts) + geom_bar(aes(x=as.factor(attempts))) + labs(x="Number of selections", y=NULL)
+summary(as.factor(attempts$attempts))

### 2016-01-13 paper/paper.Rnw
Conf Interval

@@
 <<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
-library(lme4)
 library(ggplot2)
@@
 
-In this research we utilise the methods in visual inference to study plot design to investigate the relationship between two temporal variables.
+In this paper we utilised the methods in visual inference to study plot design to investigate the relationship between two temporal variables.
 There is a substantial literature suggesting that scatter plots are the appropriate display to read association \citep{Cleveland, Harrison, Li}, however, when the two variables are temporal it is common to display them as time series, drawn on the same plot.
@@
 If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
-This research used $m= 20$, giving us a significance level (and Type 1 error rate) of $\alpha = 0.05$.
+We used $m= 20$, giving us a significance level (and Type 1 error rate) of $\alpha = 0.05$.
 The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
@@
   \centering
-  \hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2n-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /n_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /n_{2} }
+  \hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2K-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /K_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /K_{2} }
 \end{equation}
-
-Where $\hat\pi_{i} = (x_{i}+1)/(n_{i}+1)$, $x$ is the number of times a true lineup was correctly identified and $n$ is the Welch-Satterwaite estimate for the degrees of freedom.
 
@@
 
-Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a $95\%$ confidence interval calculated according to  formula~\ref{propci}.
+Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a joint $95\%$ confidence interval calculated according to  formula~\ref{propci}.
 Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
@@
 
-\iffalse
 \begin{figure}[htbp]
@@
 prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", evals, 0)),
-                                                  l.correct=sum(ifelse(design=="Overlaid Lines", total, 0)),
+                                                  l.detect=sum(ifelse(design=="Overlaid Lines", total, 0)),
                                                   s.attempt=sum(ifelse(design=="Scatterplot", evals, 0)),
-                                                  s.correct=sum(ifelse(design=="Scatterplot", total, 0)))
-prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
-               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
-               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
-               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
+                                                  s.detect=sum(ifelse(design=="Scatterplot", total, 0)))
+alpha <- 1-(1-0.95^(1/48))/2 
+##8 correlations * 6 design = 48 boxplots, so this the students t quantile for the joint 95% CI
+prop <- mutate(prop, l.power = l.detect/l.attempt, 
+               s.power = s.detect/s.attempt,
+               df = s.attempt + l.attempt - 1, 
+               difference=s.power-l.power, 
+               t = qt(alpha, df),
+               se = sqrt(s.power*(1-s.power)/s.attempt+l.power*(1-l.power)/l.attempt),
+               upper = difference + t*se,
+               lower = difference - t*se)
 ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
@@
 @
- \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a 95\% confidence interval.
-The red line indicates the point where both plot designs are equally powerful while values above this line indicate that the scatter plot is more powerful than the overlaid line graph.
-The largest differences appear in negative correlation, where the scatter plot is more powerful.}
+ \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a joint 95\% confidence interval.
+The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid lines power drops substantially.}
  \label{powerdiff}
 \end{figure}
-\fi
 
@@
 <<random, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-randomeffects <- summarise(group_by(prediction, subj_id), re=max(diff))
+randomeffects <- summarise(group_by(prediction, subj_id), re=unique(diff))
 ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
@@
 \centering
-<<random2, dependson='setup', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 ability <- merge(full.sub[,c("subj_id", "pic_id", "score", "weights")], bylineup.sub[c("pic_id", "percent")], by="pic_id")
@@
 
-This research showed that the scatterplot is better than the overlaid lines for displaying two time series when the purpose is to examine the association between the two series.
+This paper showed that the scatterplot is better than the overlaid lines for displaying two time series when the purpose is to examine the association between the two series.
 This is especially true if correlation between the series is negative.

### 2016-01-13 paper/Extra Material.R

@@ -24,15 +24,13 @@ ggplot(data=subjplot, aes(x=byeval)) + geom_histogram(binwidth=0.05) + ylim(0,30
 #mean 0.40
 
 #Choice plots#
-full.subs$choice_reason <-sapply(strsplit(as.character(unlist(full.subs$choice_reason)), "[^0-9]+"), "[[", 1)
-ggplot(data=full.subs) + geom_bar(aes(x=choice_reason))
-choice <- full.subs %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
+full.sub$choice_reason <-sapply(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+"), "[[", 1)
+ggplot(data=full.sub) + geom_bar(aes(x=choice_reason))
+choice <- full.sub %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
 choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
-c1 <- ggplot(data=subset(choice, correct==0), aes(x=choice_reason)) + geom_bar() + 
-  facet_grid(plot~rsgn) + labs(title="Did not detect", x=element_blank(), y=element_blank())  
-c2 <- ggplot(data=subset(choice, correct==1), aes(x=choice_reason)) + geom_bar() + 
-  facet_grid(plot~rsgn) + labs(title="Detected", x=element_blank, y=element_blank())  
-grid.arrange(c1, c2, ncol=2, top = "Choice by plot design, true plot correlation and detection")
+choice$correct <- factor(choice$correct, labels=c("Did not detect", "Detected"))
+ggplot(data=choice, aes(x=choice_reason, fill=correct)) + geom_bar() + facet_grid(design~rsgn) +
+  theme(legend.position="bottom", axis.title=element_blank()) + scale_fill_discrete(name="True Plot")
 
 #Multiple selections#
 m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
@@ -63,3 +61,9 @@ ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point()
 attempts <- summarise(group_by(full.sub, subj_id, pic_id), attempts = max(attempts))
 ggplot(data=attempts) + geom_bar(aes(x=as.factor(attempts))) + labs(x="Number of selections", y=NULL)
[…]

### 2016-01-13 paper/additional material.Rnw
Additional Material

--- 2016-01-13 paper/paper.Rnw
@@
 \begin{document}
-%\SweaveOpts{concordance=TRUE}
 
@@
 
-\title{\bf Comparing the Power of Plot Designs to Reveal Correlation}
-
-\if0\blind
-{
-\author{Nathaniel Tomasetti and Dianne Cook\\
-Department of Econometrics and Business Statistics, Monash University\\
-}
-  \maketitle
-} \fi
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Comparing the Power of Plot Designs to Reveal Correlation}
-\end{center}
-  \medskip
-} \fi
-
-\bigskip
-
-\begin{abstract}
-Visual inference in EDA is prone to type 1 errors from the over-interpretation of randomness \citep{Daniel, Diaconis}.
-Two competing plot designs, the scatter plot and overlaid line graph are both popular in the analysis of time series data.
-Lineups \citep{Majumder:2013, Wickham} allow the visual inference power of a graphic display to be evaluated, and were used to compare the plot designs.
-We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the decision for 2089 lineup evaluations.
-The results show that the scatter plot is both the faster and more powerful plot design, despite its inability to display the time dimension.
-\end{abstract}
-
-\noindent
-{\it Keywords:}  Lineups, visual inference, power comparison, scatter plot, line graph, data visualisation, visual analytics, time series, temporal data
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
-\newpage
+\title{Additional Material}
 
@@
 
-<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
+<<setup, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+library(maps)
 library(ggplot2)
-library(mnormt)
+library(dplyr)
 library(tidyr)
-library(knitr)
-library(nullabor)
-library(dplyr)
-library(mFilter)
-library(grid)
-library(gridExtra)
-library(stringr)
-library(maps)
-#library(readr)
-#library(lubridate)
-# Convert these to read_csv at some point
-correl <- read.csv("correlations.csv", stringsAsFactors = FALSE) ##Correlations of every plot in every lineup
 full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
@@
 
-##Summarise data by lineup
-
-#sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
-#sasadata <- mutate(sasdata, rabs = as.numeric(rabs),
-#                   subj_id = as.numeric(subj_id)
-#                   design = ifelse(design=="Scatterplot", 1, 0))
-#write.csv(subsetted, "sas/sasdata.csv", row.names = FALSE)
-##The code that is exported to fit the model into SAS, see sas/model fit.sas
-
-# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
-#                                      smoothed=c(0,1), design=c(0,1),
-#                                      subj_id=full.sub$subj_id))
-#pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)
-#write.csv(pred, "sas/predgrid.csv", row.names=FALSE) # these lines are done, but its a large file so commented out for knitr
-#Code exported to SAS to predict new values
-
 colnames(prediction) <- c("rabs", "t", "alpha", "rsgn", "smoothed", "design", "subj_id", "weights", "detected", "log_time", "RandomEf", "FixedEf", "residual")
@@
                      diff = round(RandomEf - FixedEf,5))
-##Data from SAS
-
-gen_true_data <- function(n, r, smoothed = FALSE) {
-  d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2))) 
-  d$X1 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X1))
-  d$X2 <- as.vector(arima.sim(list(ar=n/100), n, innov=d$X2))
-  if (smoothed) {
-    d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
-    d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
-    d <- as.data.frame(d)
-  }
-  d$X1 <- scale(d$X1)
-  d$X2 <- scale(d$X2)
-  d$t <- 1:n
-  return(d)
-}
-##Generates a pair of data, length n, correlation r
-
-gen_null <- function(n, m=20, smoothed = FALSE){
-  nd <- NULL
-  for(i in 1:(m-1)) {
-    d <- data.frame(X1 = as.vector(arima.sim(list(ar=n/100), n)),
-                    X2 = as.vector(arima.sim(list(ar=n/100), n)))
-    if (smoothed) {
-      d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
-      d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
-      d <- as.data.frame(d)
-    }
-    d$X1 <- scale(d$X1)
-    d$X2 <- scale(d$X2)
-    nd <- rbind(nd, d)
-  } 
-  nd$.n <- rep(1:(m-1), each = n) 
-  nd$t <- 1:n
-  return(nd)
-}
-##Generates m-1 pairs of uncorrelated data, length n
 @
 
-\section{Introduction}
-
-In order to work with data, it first must be understood. Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first.
-This is especially so today, for vast databases that have been assembled in the big data era, that now need the data scientist to unravel the meaning of the numbers. 
-Without preset hypotheses to test, the power of statistical inference is impotent, and without hypotheses the data analyst can stumble blindly trying to build up models of structure in the data. 
-To understand data requires good visualisation.
-This idea was formed as early as the 18th century, when William Playfair institutionalized the then revolutionary idea of graphing government and economic data.
-Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information \citep{Sachs}.
-Since then, advances in computing power have allowed statistical graphics to flourish, spearheaded by \citet{Tukey} in 1965 into the new domain of exploratory data analysis (EDA) \citep{Friendly}.
-EDA can be thought of as a well understood \citep{Cleveland, Vanderplas} set of tools and techniques required to visualise information, to physically see what the data contains.
-In particular, it incorporates a free roaming approach, where the analyst is able to explore structure to find whatever relationships and structure exists within.
-Critically, the analyst does not have to have any pre-conceived ideas or hypotheses -- they are not specifically looking for any one particular thing.
-EDA emphasises letting the data inform us and can lead to the discovery of otherwise unexpected relationships, many of which may seem to become completely obvious after discovery.
-With the knowledge provided by EDA, ideas are generated about what relationships between variables may potentially exist.
-This then enables the analyst to use these new hypotheses upon which to apply classical inference and rigorously conduct tests with new data.
-EDA is also related to the field of model diagnostics (MD), where a model can be continuously refined through the visualisation of its fit, its residuals, and the interactions with its variables.
-Both EDA and MD follow the same framework: Visualise the data, look for patterns that suggest an underlying relationship, and if one is found implement it into the model then continue the exploration of the data.
-It has long been thought that EDA and statistical inference were worlds apart, but recent work in \citet{Buja} and \citet{Majumder:2013} bridges this chasm.
-Framing a data plot as a test statistic, which when compared to plots of null data, places EDA into the statistical inference recipe.
-The null hypothesis underlying a particular plot, is generically that there is no pattern, and particular types of plots implicitly regulate what “no  pattern” means.
-For example, a scatter plot of two variables is used to explore for some sort of association, so the implicit hypothesis is that there is no association between the two variables.
-The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association. 
-The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true.
-If an impartial observer asked to pick the plot that is different from the rest, picks the data plot this suggests there exists a pattern that is not the result of chance, and the null hypothesis is rejected.
-For a scatter plot, this departure from the null, might be a single outlier, or few outliers, a non-linear pattern, or clusters, which the human eye detects as more different in this data plot than in any of the nulls.
-This is the reason why visualisation remains important today, human eyes can detect patterns which would not be detected mathematically.
-But eyes need calibration, which the lineup protocol provides.
-On its own, with a single plot, because of random sampling in collecting data, it is easy for the analyst to imagine a pattern when no real structure exists in the population.
-Visual skills of an observer in EDA is prone to Type I error, where the null hypothesis is rejected when it is actually true, caused by the inherently random formation of patterns when visualised which are attributed to structure rather than chance.
-\citet{Daniel} warns against this, by providing 40 pages of plots from the null distribution, he encourages data analysts to understand the patterns that can be created from data without inherent structure.
-(This is what \citet{Buja} call the Rorschach protocol.) By being aware of what can appear in this type of data, the analyst should be more wary of claiming any visual feature they find is a true relationship.
-But this is not enough, even seasoned data analysts can be misled.
-\citet{Diaconis} introduced the notion of 'magical thinking' which argues that people commonly suffer from the over-interpretation of randomness, particularly if it matches a pre-conception.
-If it suits their particular bias, an analyst may make false discoveries of structure and false rejections of the null hypothesis.
-Whilst using both can complement each other, the treatment of Type I error has led EDA and inferential statistics to be considered as very disparate pursuits.
-The mathematical rigor that governs classical inference relies on a framework that acknowledges and controls for the rate of Type I error.
-Section 2 describes the visual inference protocols which put EDA more firmly into the rigorous framework of statistical inference.
+\section{Subject Details}
 
@@
 \centering
-  \includegraphics[width=0.5\textwidth]{cnyline}
-  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
-(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
-  \label{cny}
-\end{figure}
-
-In this paper we utilised the methods in visual inference to study plot design to investigate the relationship between two temporal variables.
-There is a substantial literature suggesting that scatter plots are the appropriate display to read association \citep{Cleveland, Harrison, Li}, however, when the two variables are temporal it is common to display them as time series, drawn on the same plot.
-Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and there is a strong argument for its presence in financial data \citep{Sher, Turtle}.
-It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data.
-
-\begin{figure}[htbp]
-  \centering
-    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
-    \includegraphics[width=0.45\textwidth]{nzdline}}
-  \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
-Which one can you read the correlation from more accurately?
-Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
-  \label{nzd}
-\end{figure}
-
-To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by \citet{Wertheimer}.
-This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists.
-\citet{Lee} work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot.
-However, keep in mind magical thinking \citep{Diaconis},   the desire to find any pattern may lead to a false discovery of positive correlation.
-The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier.
-If positive correlation does not exist, the observer may be fooled into thinking it does anyway.
-Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot.
-Figure~\ref{cny} highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
-However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
-Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm.
-\citet{Robbins} is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
-When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small.
-However, this is incorrect, as looking at the small horizontal distance is comparing the lines at different dates.
-It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993.
-\citet{Vanderplas} examines this effect in more detail.
-It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope.
-As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice.
-However, in practice many data analysts are split between the two major alternatives for temporal data.
-The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{murders}
-  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
-The difference between the two represents murders outside of the city.
-What size would you judge the difference to be in 1994?}
-  \label{murder}
-\end{figure}
-
-There has been other research on the perception of temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
-Javed et al.
-examined optimal ways to visualise local features, such as, which variable had the highest value at a given point in time, and global features, such as, comparing the size of the overall slope of many different variables.
-The finding was that a different graphical layout is optimal for each type of task.
-However, they did not test for the perception of correlation, which can be treated as both a local and a global feature within the graph.
-The slope of each variable must be compared at a point in time, and they must have a persistent relationship across at least the majority of the series.
-However there is a quandary, the visual features that improve local tasks are unsuited to global tasks and vice versa; so we must now find a form that is well suited to assess both types of tasks simultaneously.
-This paper addresses this deficiency.
-Section 2 describes the lineup method utilised to rigorously compare plot designs.
-Section 3 explains the experimental design.
-Section 4 contains the results of the experiment and Section 5 discusses the implications of the findings.
-
-\section{The Lineup Protocol}
-
-With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing?
-In order to use visualisation effectively, we will get the most benefit if the type of graphic display chosen is best suited to the task at hand, be it EDA, MD, or even the presentation of results.
-We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within.
-This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
-They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
-The lineup protocol is inspired in part by the police lineup \citep{Wickham}.
-We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people).
-If the real data plot is selected by an uninvolved observer as being most different from the other plots there is evidence that the structure of that data has led to a significant difference in the plot (That the criminal is sufficiently different from the innocents).
-This constitutes a rejection of the null hypothesis.
-If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error.
-$p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup.
-It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hofmann:2014}.
-If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution.
-The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
-However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test.
-For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
-To control our error rate, it is initially obvious that we can simply change $m$, with an increase in null lineups having an inverse reduction in $\alpha$.
-We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:
-
-\begin{equation}
-\label{pvalue}
-  \centering
-  p-value = P(X \geq x | H_{0}) = 1 - B_{n, 1/m}(x-1)
-\end{equation}
-
-If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
-We used $m= 20$, giving us a significance level (and Type 1 error rate) of $\alpha = 0.05$.
-The plot of 'true' data is placed randomly amongst 19 null plots to form the lineup.
-
-If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability $p_{j} = s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
-In the event that a lineup has observers making different numbers of selections, we have a binomial distribution with different success probabilities, and a Poisson Binomial distribution must be used to calculate the p-value.
-The probability mass function of a Poisson Binomial distribution can be written as:
-
-\begin{equation} 
-\label{pvalue2}
-\centering
-P(X=x) = \sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})
-\end{equation}
-
-Where $T_{x}$ is all possible sets of $x$ trials out of $K$ resulting in a success, $A \in T_{x}$ is a specific set of $x$ trials resulting in a success and $A^{c}$ is the complement to $A$, the set of $K-x$ trials resulting in a failure.
-
-\begin{figure}[htbp]
-  \centering
-<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-set.seed(1234)
-a <- gen_true_data(24, 0.95)
-td <- gather(a, variable, value, c(X1, X2))
-b <- gen_null(24)
-nd_l <- gather(b, variable, value, c(X1, X2))
-pos <- 7
-td <- data.frame(.n=rep(pos, nrow(td)), td)
-lg <- nd_l$.n < pos
-nd_l1 <- data.frame(.n=nd_l$.n[lg], t=nd_l$t[lg], variable=nd_l$variable[lg], 
-                    value=nd_l$value[lg])
-nd_l2 <- data.frame(.n=nd_l$.n[!lg], t=nd_l$t[!lg], variable=nd_l$variable[!lg], 
-                    value=nd_l$value[!lg])
-nd_l2$.n <- nd_l2$.n + 1
-d <- rbind(nd_l1, td, nd_l2)
-
-ggplot(data=d, aes(x=t, y=value, colour=variable)) + geom_line() + facet_wrap(~.n, ncol=4, scales="free_y")  +
-    theme_bw() + theme(axis.title.x = element_blank(),
-                       axis.title.y = element_blank(),
-                       axis.text.x = element_blank(),
-                       axis.text.y = element_blank(),
-                       axis.ticks = element_blank(),
-                       legend.position="none")
+<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+location <- read.csv("location.csv")
+map.dat <- map_data("world")
+ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
 @
-  \caption{A lineup ($m = 20$) of line graphs. Which plot shows the series with the strongest correlation?}
-  \label{lineup:example}
-\end{figure}
-
-Essentially, the more time our subjects pick the correct plot out of the lineup, the more confident we are that the real plot was visually significantly different to the nulls; and that the type of graphic involved has the power to display association.
-The lineup effectively allows us to conduct a hypothesis test on visual features, with the competing hypotheses:
-
-\begin{itemize}
-  \item $H_{0}$ : There is no association between the two series (evidence for $H_{0}$: the associated data is indistinguishable from the nulls).
-  \item $H_{1}$: There is an association between the two series (evidence against $H_{0}$: the associated data can be distinguished from the nulls).
-\end{itemize}
-
-As the power of a statistical test is defined as the probability of rejecting $H_{0}$ when it is false, the power of a lineup test is viewed as the probability of detecting the true plot.
-We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ correctly detected the true plot out of the lineup.
-We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as \citep{Hofmann:2012}:
-
-\begin{equation} \label{propci}
-  \centering
-  \hat\pi_{1} - \hat\pi_{2}\pm t_{1-\alpha,2K-1}\sqrt{\hat\pi_{1} (1-\hat\pi_{1}) /K_{1} + \hat\pi_{2} (1-\hat\pi_{2}) /K_{2} }
-\end{equation}
-
-Some individuals may have a better natural ability at detecting the correct correlation pattern in the graphics, but by having each participant viewing multiple lineups and each lineup being viewed by multiple different participants, this can be controlled via a random effect variable in the model.
-We recruited participants through Amazon's Mechanical Turk, where the lineup protocol has been applied to a number of problems in prior papers \citep{Heer, Majumder:2013, Wickham}.
-Amazon’s Mechanical Turk \citep{MTurk} is a labour crowd-sourcing platform developed to give easy access to workers with basic tasks paid in line with the United States minimum wage.
-MTurk can be used to recruit subjects to read a variety of graphics and report on particular visual tasks.
-The time taken to decide, confidence in the decision and the reason for that decision can also be recorded.
-\citet{Heer} use MTurk to replicate previous findings in graphical perception from \citet{Cleveland} and find that it is a valid method of data collection, once controls to eliminate anyone 'gaming' the system by randomly selecting answers to minimise time spent working are implemented.
-Further details of our use of Amazons’ Mechanical Turk is included in Section 3.
-The protocol has successfully been used to measure statistical power as a means to determine plot type superiority in \citet{Hofmann:2012}, and this work follows that approach.
-
-\section{Experimental Design}
-
-To allow for a greater control of distributions of the data in the lineups, we use simulated data.
-The data must contain autocorrelation for there to be meaningful information in the time dimension.
-To fulfil this requirement, each plot has a standardised pair of AR(1) models defined by:
-
-$$Y_{1,t} = \beta \times Y_{1,t-1} + e_{1,t}$$
-
-$$Y_{2,t} = \beta \times Y_{2,t-1} + e_{2,t}$$
-
-With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ being generated with from a bivariate normal distribution:
-
-$$e_{t} \sim \mathcal{N} \left( \mu = \left| \begin{array}{c}0 \\ 0 \end{array} \right| , \hspace{4mm} \Sigma = \left| \begin{array}{cc}1 & \rho \\ \rho & 1\end{array} \right| \right)$$
-
-The variation in the covariance parameter allows for control in the correlation in the pairs of data.
-The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$.
-True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$.
-Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$.
-Further to this, relationships weaker than this point are often of little interest to analysts.
-Hence we found it unnecessary to test a more full range of correlations.
-Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$.
-The null pairs often had correlation generated spuriously through the simulation process.
-This was particularly problematic for generating small sample data, but setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
-For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications.
-To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component.
-This filter creates a trend by penalising both errors in smoothness and in fit.
-The trend of the series, $\tau_{t}$ is defined by the equation:
-
-\begin{equation}
-\centering
-\label{hpfilter}
-\min_{\tau}(\sum_{t=1}^{T}(y_{t}-\tau_{t})^2 + \lambda \sum_{t=2}^{T-1}[(\tau_{t+1}-\tau_{t})-(\tau_{t}-\tau_{t-1})]^2)
-\end{equation}
-
-Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation.
-However, for $t = \{12, 24\}$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = \{48, 96\}$.
-Each generated lineup was produced as both a scatter plot and an overlaid line graph.
-
-\begin{figure}[htbp]
-  \centering
-<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-set.seed(123)
-uns <- gen_true_data(96, 0.8)
-uns_l <- gather(uns, variable, value, -t)
-p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
-    theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
-    theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-
-set.seed(123)
-sm <- gen_true_data(96, 0.8, smoothed = TRUE)
-sm_l <- gather(sm, variable, value, -t)
-p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
-    theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-
-grid.arrange(p1a, p2a, p3a, p4a, ncol=2)
-@
-  \caption{The same AR(1) data both unsmoothed and smoothed for both plot designs. t = 96, correlation = 0.86}
-  \label{smoothing}
-\end{figure}
-
-The lineups were generated with three basic factors:
-
-\begin{itemize}
-  \item Factor 1: Plot design, levels = scatter plot, overlaid time series
-  \item Factor 2: Sample size (t), levels = \{12, 24, 48, 96\}
-  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$
-\end{itemize}
-
-Additionally, there was a fourth factor for $t = \{48, 96\}$:
-
-\begin{itemize}
-  \item Factor 4: Smoothed, levels = yes, no
-\end{itemize}
-
-Each combination of factors was replicated in three different lineups, giving a total of 288 lineups (3 replications x 2 plots x 8 correlations x 6 sample size and smoothness combinations). 
-The position of the actual data plot in the lineup was randomized.
-The order that the factors are presented to a subject were also randomized, and each subject saw 5 scatter plots and 5 overlaid lines at a range of all absolute correlation levels.
-Each subject did not see the same data more than once.
-Subjects were asked to pick the plot in the lineup that has the strongest association, with basic examples of negatively and positively correlated plots.
-They also answered with reasons for choosing the plot, and the confidence that they have that this really is the plot showing the strongest correlation.
-Each subject evaluated 10 lineups, plus a further two trials to eliminate people attempted to game the system.
-If a trial was not answered correctly, the subject could not evaluate the remaining lineups.
-If it was answered correctly, the trial data was discarded and the subject moved onto the rest of the lineups.
-Earlier work by \citet{Majumder:2014} has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
-We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
-In total there were 2088 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations leaving us with 1684 lineup evaluations for the analysis.
-
-Subjects were allowed to make multiple selections from each lineup evaluated, additional selections on lineup $i$ by subject $j$ are penalised according to a score defined by:
-\begin{equation}
-\label{score}
-{score}_{ij} =\frac{m-s_{ij}}{m-1} \mathbbm{I}_{ij}
-\end{equation}
-
-where $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise, $s_{ij}$ is the number of selections made and $m = 20$ is the number of plots in the lineup.
-
-However, for the model fit to power, multiple selections are seperated and given a weighting equal to to inverse of the number of selections made in that lineup evaluation.
-This increases our data from 1684 to 1849 evaluations.
-
-\begin{figure}[htbp]
-  \centering
-<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-set.seed(101)
-a <- gen_true_data(24, 0.95)
-b <- gen_true_data(24, -0.95)
-a_l <- gather(a, variable, value, -t)
-b_l <- gather(b, variable, value, -t)
-p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
-@
-  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
-  \label{designs}
-\end{figure}
-
-The web site that the Turk workers accessed to complete the task can be seen at http://104.236.245.153:8080/mahbub/turk18/index.html.
-You can also try out the tasks, but your results will not be recorded.
-(The experiment cost \$210 to run.)
-
-\section{Results}
-
-\subsection{Detection, Time and Confidence}
-
-\begin{figure}[htbp]
- \centering
-<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
-ggplot(data=bylineup, aes(x=r, y=percent, colour=design)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
-  scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
-  theme(legend.position="bottom")
-@
-  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size.
-The dots are data points, the lines represent a naive model average fit.
-Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at revealing negative correlation.}
-    \label{results}
-\end{figure}
-
-Figure~\ref{results} shows the raw results of the survey and Figure~\ref{powerdiff} shows the difference in power with a joint $95\%$ confidence interval calculated according to  formula~\ref{propci}.
-Values above zero indicate that the scatter plot had more power to detect the true plot than the overlaid lines.
-Performance between the two plot designs appears to be similar for positive correlation, with the scatter plot power gradually improving relative to the overlaid lines as sample size increases.
-The scatter plot appears to be the superior choice as it retains its power with negative correlation, where the overlaid lines performance suffered.
-This is not surprising, as positive and negative correlation appear as symmetric graphs in the scatter plot but the two characteristics appear very differently in the overlaid line graph (Figure~\ref{designs}).
-The right side shows the performance benefits of smoothing overlaid line time series data.
-Much of the 'noise' in the time series is removed and studying the overall trend shows a greatly improved ability to successfully detect the true data plot when it is negatively correlated.
-The scatter plot still appears to be the superior plot, but much of the gap between the two is reduced for positive correlation.
-
-\begin{figure}[htbp]
-  \centering
-<<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-prop <- select(bylineup, smoothed, n, evals, total, r, design)
-prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", evals, 0)),
-                                                  l.detect=sum(ifelse(design=="Overlaid Lines", total, 0)),
-                                                  s.attempt=sum(ifelse(design=="Scatterplot", evals, 0)),
-                                                  s.detect=sum(ifelse(design=="Scatterplot", total, 0)))
-alpha <- 1-(1-0.95^(1/48))/2 
-##8 correlations * 6 design = 48 boxplots, so this the students t quantile for the joint 95% CI
-prop <- mutate(prop, l.power = l.detect/l.attempt, 
-               s.power = s.detect/s.attempt,
-               df = s.attempt + l.attempt - 1, 
-               difference=s.power-l.power, 
-               t = qt(alpha, df),
-               se = sqrt(s.power*(1-s.power)/s.attempt+l.power*(1-l.power)/l.attempt),
-               upper = difference + t*se,
-               lower = difference - t*se)
-ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
-@
- \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a joint 95\% confidence interval.
-The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid lines power drops substantially.}
- \label{powerdiff}
+\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
+\label{map}
 \end{figure}
@@
 \centering
-<<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
+<<users, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+users <- read.csv("turk18_users.csv")
+users$gender <- ifelse(users$gender==1, "Male", "Female")
+ggplot(data=users, aes(x=factor(age), fill=factor(academic_study))) + geom_bar() + facet_wrap(~gender) +
+  scale_fill_manual(values=c("#550000", "#883388", "#1133FF", "#FF9933", "#555555"))
 @
- \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data.
-The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
- \label{samplesize}
-\end{figure}
-
-Figure~\ref{samplesize} demonstrates the effects of changing sample size for unsmoothed data.
-The scatter plot has increased power to detect correlation almost universally for any increase.
-Sample size influences the overlaid lines differently, they appears to benefit for positively correlated true plots, but it does not improve the weak performance in negative correlation.
-Figure~\ref{time} and figure~\ref{confidence} give a breakdown of the interactions between the response variables: plot detection, time and confidence.
-Figure~\ref{time} is the histogram of the time taken to record a response, split by plot design and sign of correlation and true plot detection.
-The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed.
-We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a choice.
-Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to pick the true plot and 32.6 seconds if a null was selected.
-The scatter plots performed much better, with the true plot picked 52.9\% of the time, with an average of 22.7 seconds required to make a decision.
-If the true plot was detected it took only 18.6 seconds, and 29.3 seconds for a null plot selection.
-
-\begin{figure}[htbp]
-  \centering
-<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-time <- select(full, design, log_time, correct, rsgn)
-time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
-tm <- summarise(group_by(time, correct, design, rsgn), m=mean(log_time))
-t1 <- ggplot(data=subset(time, correct==0), aes(x=log_time)) + geom_histogram(binwidth=0.5) + 
-  geom_vline(data=subset(tm, correct==0), aes(xintercept = m), color="red") +
-  facet_grid(design~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
-t2 <- ggplot(data=subset(time, correct==1), aes(x=log_time)) + geom_histogram(binwidth=0.5) + 
-  geom_vline(data=subset(tm, correct==1), aes(xintercept = m), color="red") +
-  facet_grid(design~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
-grid.arrange(t1, t2, ncol=2)
-@
-  \caption{Histograms of log time taken to make a decision.
-r+ and r- indicate the sign of true plot correlation.
-The vertical red line is the average time of the particular group.
-On average scatter plots are faster and the true plot is detected more often, with r-.}
-    \label{time} 
-\end{figure}
-
-Subjects were asked to give a number from 1 to 5 to rate how confident they were in there choice (1 - most confident, 5 - least confident, average = 3.34).
-There were no statistically significant differences between plot designs, however subjects were slightly less confident on true plot detections (by $0.25 \pm  0.16, p=0.002$) and every one percent extra time required to make a decision increased confidence (by $0.19 \pm 0.07, p<0.001$).
-One possible explanation is that subjects often focused on a null plot and picked it confidently.
-This would suggest that the null plots often produced visually interesting features subjects assumed were caused by the association between plots.
-Subjects spending longer on a plot immediately draws two hypotheses:
-
-\begin{enumerate}
-  \item The lineup was more difficult and it took longer to determine which plot was best, reducing confidence.
-  \item The subject picked a plot relatively easily but spent extra time going over its details and making comparisons to alternative plots, increasing confidence.
-\end{enumerate}
-
-The result suggests that the latter hypothesis is more likely to be correct.
-We also found a negative relationship between plot detection and time spent on a lineup, supporting the major relationship found: quick decisions are less confident but more likely to successfully detect the true plot.
-
-\begin{figure}[htbp]
-  \centering
-<<confidence, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-conf <- select(full, design, conf_level, correct, rsgn)
-conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
-cm <- summarise(group_by(conf, correct, design, rsgn), m=mean(conf_level))
-c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf_level)) + geom_histogram(binwidth=1) + 
-  geom_vline(data=subset(cm, correct==0), aes(xintercept = m), color="red") +
-  facet_grid(design~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
-c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf_level)) + geom_histogram(binwidth=1) + 
-  geom_vline(data=subset(cm, correct==1), aes(xintercept = m), color="red") +
-  facet_grid(design~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
-grid.arrange(c1, c2, ncol=2)
-@
-  \caption{Histograms of reported confidence levels.
-1 - Low, 5 - High.
-Detection had a minor detrimental impact on confidence, whilst time taken increased it.
-There was no statistically significant difference on confidence level between plot designs.}
-  \label{confidence} 
-\end{figure}
-
-%Next we analysed the individual ability of each of our subjects.
-Figure~\ref{confidence2} plots the accuracy of each subject on the y axis versus the average confidence on the x axis.
-While we expected that subjects with greater visual skills would report lower values of confidence, there does not appear to be any relationship between the two.
-The accuracy is normally distributed (Jarque-Bera p value = 0.418), with a mean of 42.5\% and a standard deviation of 17.8\%.
-The strongest subject successfully identified the true plot 91\% of the time, whilst the weakest was never able to identify the true plot.
-As all subjects were shown a balanced mix of true plot correlations, the extreme difference in individual visual ability is striking.
-Visual ability will be controlled via a subject specific random effect in the models in the next section.
-
-Figure~\ref{confidence2} compares the reported confidence for each selection against the true plot correlation.
-The results are mixed with no consistent increase or decrease in confidence for changes in correlation across different plot designs and sample size.
-Strong correlation appears to decrease confidence more often than it is increased.
-
-\begin{figure}[htbp]
-  \centering
-<<confidence2, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-conf2 <- select(full, conf_level, r, smoothed, n, design) #c(7, 15
-qplot(r, 6-conf_level, data=conf2, alpha=I(0.1)) + 
-  geom_smooth(method="loess") + 
-  facet_grid(n~design+smoothed) +
-  labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
-@
-  \caption{Reported confidence against true plot correlation.
-Darker points indicate a confidence level selected more regularly.
-Strong correlation is more often associated with a lower reported confidence.}
- \label{confidence2}
-\end{figure}
-
-\subsection{Selection Analysis}
-
-There are two major reasons why a subject may fail to detect the true plot out of a lineup.
-Either the plot design was unable to display correlation for that particular combination of factors (and the analyst would commit a Type II error by not rejecting the null hypothesis), or a null plot that had a greater correlation than the true plot was selected instead (and the analyst would commit a Type I error by rejecting the null hypothesis).
-We next can analyse these kinds of errors with the pin plots found in Figure~\ref{pins}.
-The frequency of selection of each plot in each lineup, either true or null, is represented by a pin, placed at the value of the correlation between the two variables shown in that plot.
-The correlation of the selected plot is on the x-axis and the frequency of selection is plotted on the y axis.
-The correlation of the simulated (true) data ($\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$ in the plot, the sample size ($t \in \{12, 24, 48, 96\}$), and the plot design (unsmoothed/smoothed, scatterplot/lineplot) is used to facet the plots.
-The blue pins indicate the true plots and the red pins indicate the null plots.
-If the plot design has the power to display correlation, we would expect the highest correlation plot to be selected, leading to a Type I error where a plot from the null distribution spuriously had high correlation.
-We find that this is generally not the case.
-The selections appear to have a wide range of correlations, indicating that if a subject selected a null plot it was typically not due to excess spurious correlation.
-Instead the failure to select the highest correlated plot indicates that a Type II error was committed rather than Type I as the plot designs did not have the power to differentiate between medium and low correlation plots.
-However, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected.
-For the overlaid lines this mostly occurred only for plots with strong positive correlation, which is consistent with its general lack of power to reveal negative relationships in data.
-
-\begin{figure}[htbp]
-  \centering
-<<pins, dependson='setup', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-count.setup <- select(full, pic_id, response_0, actual_location, design, smoothed, n)
-count.setup$response_0 <- as.character(count.setup$response_0)
-counts <- data.frame(table(count.setup$pic_id, count.setup$response_0))
-counts$Var1 <- as.numeric(as.character(counts$Var1))
-counts$Var2 <- as.numeric(as.character(counts$Var2))
-colnames(counts) <- c("lineup", "plot", "count")
-pin <- merge(counts, bylineup, by.x="lineup", by.y="pic_id")
-pin <- select(pin, lineup, plot, count, smoothed, actual_location, n, design, r, evals)
-pin <- mutate(pin, r = factor(r),
-              true = ifelse(plot==actual_location, "True plot", "Null plot"), 
-              percent = count/evals,
-              base = 0,
-              correlation = diag(as.matrix(correl[pin$lineup, (pin$plot+2)])))
-pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + 
-         labs(title ="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) +
-         labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + 
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
-@
-  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
-On the x-axis is the correlation of the plot (there typically is spurious correlation in the null plots), and on the y-axis is the proportion of times a plot was chosen.
-Each plot contains data from the three lineups with the requisite factor combination.
-Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins.
-Strongly correlated plot seem to be selected regardless of if the relationship is real or spurious.}
-  \label{pins}
-\end{figure}
-
-Occasionally there was an anomaly where a single null plot was selected with a high frequency despite having a weak correlation.
-Occasionally the null distribution cangenerate visually interesting patterns that draw the analyst’s attention.
-We asked subjects for reasons behind their selections and found that choosing on the basis that “the points in the plot were close together”, or had “a unique pattern” more often led to an unsuccessful detection.
-Conversely, decisions based on “matching the peaks and troughs” of the overlaid lines or “seeing a line formed in the dots” of the scatter plot were more likely to be successful in identification of the true plot.
-However, these true detections are, on average, made less confidently than unsuccessful evaluations, which indicates that decisions made on the basis of closeness or patterns are more convincing of an association than the more successful reasons.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{anom1}
-  \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
-The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33).
-The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037.
-Subjects most often cited the clustering effect shown as the reason for their selection, an example of the over-interpretation of randomness.}
-    \label{anom1}
+\caption{Age and education by gender. Most subjects are in the younger brackets and there are more males than females}
+\label{ageed}
 \end{figure}
@@
 \centering
-  \includegraphics[width=0.5\textwidth]{anom2}
-  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
-The true plot in position 8 was not detected in five evaluations.
-However, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
-When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.}
-  \label{anom2}
-\end{figure}
+<<choices, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+full.sub$choice_reason <-sapply(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+"), "[[", 1)
+ggplot(data=full.sub) + geom_bar(aes(x=choice_reason))
+choice <- full.sub %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
+choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
+choice$correct <- factor(choice$correct, labels=c("Did not detect", "Detected"))
+ggplot(data=choice, aes(x=choice_reason, fill=correct)) + geom_bar() + facet_grid(design~rsgn) +
+  theme(legend.position="bottom", axis.title=element_blank()) + scale_fill_discrete(name="True Plot")
+@
 
-Figures~\ref{anom1} and \ref{anom2} show lineups where a null plot with a strong pattern led to its selection over the data plot.
-Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
-In Figure~\ref{anom1}, the true data plot is in position 4, but 9 out of 11 subjects chose plot 6.
-Plot 6 is unusual, most of the points are scattered around the vertical median, but there are two outlying points, one with a high y-value and one with a low y-value.
-Other concurrent work has suggested that outliers are pre-attentive patterns and this plot selection by the observers is consistent with that.
-Figure~\ref{anom2} above is another example of where subjects chose a null plot that had smaller correlation than the data plot.
-Four out of five evaluations selected the null plot in position 13,  instead of the data plot in position 8.
-Plot 13 has a correlation of 0.225, smaller (in absolute value) than the -0.707 of plot 8.
-This selection is in-line with Wertheimer’s gestalt law of common fate.
-It appears to be strongly correlated in the beginning of the series but breaks apart towards the end.
-The selection implies that subjects selecting this plot focused on the joint path of the two series early in the plot and ignored the contradictory evidence of no relationship as they diverged later in the plot.
-It also confounded by the seeming preference to find positive correlation over negative correlation, and we suspect that if one of the series in plot 8 had the values flipped, then the strong positive correlation would have been noticed.
+\begin{figure}[htbp]
+\centering 
+<<avscore, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=LSE>>=  
+subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
+subjplot <- mutate(subjplot, avscore= total/eval)
+ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.05)
+@
 
-\subsection{Modelling Graphical Power}
+\section{Multiple Selection}
 
-A mixed effects logit model is used separately for scatter plot and overlaid line evaluations. The tested model is:
+\begin{figure}[htbp]
+\centering
+<<multiselect, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
+m.freq <- data.frame(table(m.select$pic_id))
+m.lineup <- merge(m.freq, Turk, by.x="Var1", by.y="pic_id")
+ggplot(data=m.lineup) + geom_bar(aes(x=r, fill=plot), binwidth=0.2) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
+ggplot(data=subset(m.lineup, smoothed==1)) + geom_bar(aes(x=r, fill=plot), binwidth=0.2) + facet_wrap(~n)
+@
 
-\begin{eqnarray} 
-\label{modelpower}
-Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
-\end{eqnarray}
-
-Where $\psi$ is the logit link function, $\psi(x) = \frac{e^{x}}{1+e^{x}}$. Additionally:
-
-\begin{itemize}
-\item $x_{1,i}$ is the absolute value of the true plot correlation, $\rho$, with $i \in \{0.3, 0.5, 0.7, 0.9\}$,
-\item $x_{2,j}$ is a dummy variable for the sign of the true plot correlation, with $j = 0$ for negative correlation, 
-\item $x_{3,k}$ is a dummy variable for the smoothing procedure, with $k = 0$ for unsmoothed data, 
-\item $x_{4,t}$ is the sample size, with $t \in \{12, 24, 48, 96\}$ ,
-\item $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type,
-\item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 157$.
-\end{itemize}
-
-It is assumed that $\epsilon_{ijkt} \sim \mathcal{N} (0, \Sigma_{ijkt})$ and $\gamma \sim \mathcal{N} (0, \delta)$.
-
-If a subject made multiple selections from the same lineup, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
-
-\begin{table}[htp]
-\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design.
-Sample size for scatter plot evaluations: 900, overlaid lines: 949, total: 1849.
-Default plot design is overlaid lines.
-The model is fitted using the GLIMMIX procedure in SAS.}
-\label{tab1}
-\centering
-<<powerfit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
-powercoef <- powercoef[,-1]
-rownames(powercoef) <- c("Intercept", "abs(r)", "sign(r)", "abs(r)*sign(r)", "smoothed", "t", "design", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*smoothed", "design*t")
-colnames(powercoef) <- c("Estimate", "Standard Error", "df", "t-value", "p-value")
-kable(powercoef, format="latex", digits=2)
+\begin{table}
+\caption{Number of selection per lineup evaluation. The subject with the most selections is reported }
+\label{tabattempts}
+<<tabattempts, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+attempts <- summarise(group_by(full.sub, subj_id, pic_id), attempts = max(attempts))
+attempts.table <- data.frame(table(attempts$attempts))
+attempts.table$most <- c(0,0,0,5,3,1,1)
+colnames(attempts.table) <- c("Selections", "Whole Sample", "Subject 123")
+kable(attempts.table, format="latex")
 @
@@
 
-Table \ref{tab1} summarises the model estimates.
-It can be confusing the interpret the model estimates, and at first glance it appears the negative plot coefficient implies that overlaid lines are substantially more powerful than the scatter plot, but this is misleading and interpreting the full model involves combining the estimates for the different interactions with $\rho$.
-$\beta_{2}$ and $\beta_{8}$ model the power difference between low and high correlations, while $\beta_{3}, \beta_{4}, \beta_{9}$ and $\beta_{10}$ model any increase in power appearing only in positive correlation.
-Taking these into account, the scatter plot shows superior performance relative to the overlaid lines at higher values of abs($\rho$).
-It needs to be restated that we only tested correlations at  $0.3, 0.5, 0.7$ and $0.9$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
-
-%The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (Figure~\ref{smoothing}.
-Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
-
-Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
-The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
-\citet{Majumder:2014} show that this random visual skill is not impacted by demographics such as age and gender; whilst education increases the power of a graphic display by only approximately one percent, not a practically significant amount.
-We did not include these characteristics of the individual in the modelling due to this lack of significance.
-Figure~\ref{powermodel} especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
-
 \begin{figure}[htbp]
 \centering
-<<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-predplot <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
-predplot <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), 
-                   FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
-##Logit link function transform
+<<multiattempts, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+ggplot(data=attempts) + geom_bar(aes(x=as.factor(attempts))) + labs(x="Number of selections", y=NULL)
+@
+\caption{Histogram of table~\ref{attempts}}
+\label{attempts}
+\end{figure}
 
-qplot(r, RandomEf, data=predplot, color=design, 
-      group=interaction(subj_id, design), geom="line", alpha=I(0.1)) + 
-  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
-  xlab("Correlation") + ylab("Estimated Power") +
-  geom_line(aes(y=FixedEf), size=2, alpha=1) + theme_bw() + facet_wrap(~t, ncol = 2) + theme(aspect.ratio=1.2)
-@
-  \caption{Predicted power fit from the mixed effects logit power model.
-The thicker lines shows the fixed effects prediction while the thinner lines show the range of random effects, the observer's visual ability.
-The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
-  \label{powermodel}
-\end{figure}
+\section{Model Validation}
 
@@
 
-\iffalse
-This unrestricted model used in to test the symmetry restriction for the scatter plot design was written earlier as:
-
-$$Power_{UR,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \beta_{5} x_{1,i} \* x_{2,j} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
-
-Testing if the scatter plot displays positive and negative correlation equally is equivalent to testing the hypothesis:
-
-\begin{itemize}
-  \item $H_{0}: \beta_{2} = \beta_{5} = 0$
-  \item $H_{1}:$ Either equality does not hold.
-\end{itemize}
-
-The restricted model is therefore:
-
-$$Power_{R,ijkt} = \psi\left\{ \beta_{0} + \beta_{1} x_{1,i} + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \gamma z_{s} + \epsilon_{ijkt} \right\}$$
-
-The unrestricted model has a log-likelihood value of -346.9, where the restricted model has a log-likelihood of -355.8.
-The test statistic is therefore, with a null distributed as a $\chi^{2}_{2}$ variable.
-The null hypothesis is rejected (p = 0.02).
-We find that the scatter plot has asymmetric performance over negative and positive values of correlation.
-This may be a result of unwittingly introducing bias into the methodology, the first scatter plot shown to subjects in the instruction set always had strong positive correlation.
-A similar setup rejects the symmetric performance of the overlaid line graphs with a log likelihood ratio test statistic of 105.30 (p-value < 0.001).
-Such an extreme result is due to a systematic difference in the overlaid line graph and cannot be attributed simply to bias.
-\fi
-
-\subsection{Modelling time taken}
-
-The time taken to evaluate a lineup is recorded and log transformed to reduce the impact of several large ($> 500$ seconds) outliers.
-
-\begin{eqnarray} 
-\label{timemodel}
-ln(time_{ijklmt}) &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \beta_{13} x_{6,m} + \beta_{14} x_{5_l} \* x_{6,m} \gamma z_{s} + \epsilon_{ijklmt} \nonumber 
-\end{eqnarray}
-
-The structure of the model ~\ref{timemodel} is similar to that in Equation~\ref{modelpower} except the link function is replace with the identity function, $\psi(x) = x$, and a new variable, $x_{6,m}$, is introduced with $m = \{0,1\}$ indicating if the subject detected the true plot, where $m = 0$ is a failure to detect the true plot.
-
-\begin{table}[htp]
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
-\label{tab2}
-\centering
-<<timefit, dependson='setup', results='asis', echo=FALSE, message=FALSE, error=FALSE, warning=FALSE>>=
-time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] 
-##separate into main effects then interaction effects in same order
-rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "design", "design*Smoothed", "design*t", "design*abs(r)", "design*sign(r)", "design*abs(r)*sign(r)", "design*detected")
-time.fit$Effect <- NULL
-colnames(time.fit) <- c("Estimate", "standard error", "df", "t-value", "p-value")
-kable(time.fit[1:14,], format="latex", digits=2)
-@
-\end{table}
-
-Table~\ref{tab2} shows a similar story for time as Table~\ref{tab1} did for power, that the overlaid lines have better performance close to zero correlation, but the scatter plot is much faster at high correlations.
-Postivie correlation reduces time for both plot designs, but only the scatter plot sees improvements in negative correlation.
-Detecing the true plot was also faster than a null.
-Neither smoothness nor sample size has an impact on the time for either plot design, despite having statistically significant impacts on power.
-%An overlaid line graph having positive correlation increases the time required, eliminating the benefits from the almost significant plot design term.
-This result seems counter-intuitive as the weak performance of the overlaid lines in negative correlation implies that negative would require more time than positive, but the interactions between plot design, correlation and sign result in there being almost no decrease in time for strongly negatively correlated line plots.
-
-\section{Conclusion}
-
-This paper showed that the scatterplot is better than the overlaid lines for displaying two time series when the purpose is to examine the association between the two series.
-This is especially true if correlation between the series is negative.
-This comparison was made possible by the use of the lineup protocol for comparing plot designs.
-The power of two plot designs for reading correlation between variables was modelled using a logistic regression incorporating subject’s individual visual ability as a random effect.
-We found that the scatterplot, in most situations is both more powerful and quicker to process than the overlaid line graph.
-The overlaid line graph appeared to have stronger performance at extremely low correlations, but this work did not test absolute values of correlation below 0.3.
-Other non-linear associations may exist in data that are not measured by correlation, and it may be of interest to compare plot power with a broader range of true plot features.
-Investigation of the individual selections of the lineup evaluations finds some results in-line with Wertheimer’s Gestalt Law of common fate, as demonstrated in Figure 15.
-The law suggested that people would find relationships between two lines that briefly move together, regardless of whether the relationship actually existed.
-Whilst this is shown, the overlaid line graph did not have a corresponding increase in confidence compared to the scatter plot; so research into similar confidence-boosting visual anomalies that can occur in scatter plots would be necessary to further investigate the law.
-Diaconis's Magical Thinking was supported by the finding on the relationships between time, confidence and correctness.
-A more difficult true plot may be picked rarely and require a longer to make a decision; resulting in decreased subject confidence.
-However, we found that time increases confidence but reduces correctness.
-Highly confident, but incorrect decisions, suggests subjects may have been distracted by strong, spurious patterns in the null plots.
-If a data analyst is required to use overlaid line graphs in their work, substantial improvements to power and hence readability can be made by smoothing time series data if the sample size is long enough for these techniques to become feasible.
-
-\bibliographystyle{asa}
-\bibliography{references}
-
-\section{Appendix}
-
-\begin{table}[htp]
-\caption{The true plot detections / attempts for each lineup in the data collection. Results ignore multiple selection weighting. True plot correlation is on the x axis and t is on the y axis.}
-\label{tab3}
-\centering
-<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-turk.tbl <- select(bylineup, design, smoothed, r, n, correct, evals)
-turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$evals, sep="/")
-turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
-scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Scatterplot")$result, ncol=8))
-scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Scatterplot")$result, ncol=8))
-line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Overlaid Lines")$result, ncol=8))
-line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Overlaid Lines")$result, ncol=8))
-colnames(scatter.uns) <- c("Unsmoothed Scatterplot", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(scatter.sm) <- c("Smoothed Scatterplot","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.uns) <- c("Unsmoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.sm) <- c("Smoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-kable(scatter.uns, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables2, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(scatter.sm, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables3, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(line.uns, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables4, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(line.sm, format="latex")
-@
-\end{table}
-
-\begin{figure}[htbp]
-\centering
-<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-location <- read.csv("location.csv")
-map.dat <- map_data("world")
-ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
-@
-\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
-\label{map}
-\end{figure}
-
 \end{document}

### 2016-01-13 paper/paper.Rnw
Additional Material

@@
 library(stringr)
-library(maps)
 #library(readr)
@@
 
-\begin{figure}
-  \centering
-<<residuals, dependson='setup', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-glmm.resid <- subset(prediction, detected==0 | detected==1)
-ggplot(data=glmm.resid) + geom_histogram(aes(y=..density.., x=residual), binwidth=2) + labs(x=NULL, y=NULL)
-@
-  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
-    \label{residuals}
-\end{figure}
-
-Analysing the level one residuals, $\hat{\epsilon_{i}}$ and level two residuals, $\hat{\gamma_{i}}$ allows us to test the model fit as described \citet{Loy}.
-There are some negative extreme outliers but the distributional assumption appears to hold.
-
-\begin{figure}[htbp]
-  \centering
-<<random, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-randomeffects <- summarise(group_by(prediction, subj_id), re=unique(diff))
-ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
-@
-    \caption{Histograms of estimated subject random effects, $\hat{\gamma}$. They appear to be normally distributed with mean zero.}
-  \label{random}
-\end{figure}
-
-In Figure~\ref{random}, the level two residuals, $\hat{\gamma}$, generated appear to be normally distributed validating the assumptions of the modelling process.
-
-The scores of each subject given by equation~\ref{score} can be summed to give a total score per subject $j$:
-
-$${score}_{j} = \sum_{i}\{{score}_{ij} | i \in L_{j}\}$$
-
-where $L_{j}$ is the set of lineups evaluated by subject $j$.
-
-The lineup $i$ with $K$ evaluations has an expected score:
-
-$${E[score]}_{i} = \frac{1}{K}\sum_{j}{score}_{ij}$$
-
-summing these per subject gives an expected total score per subject:
-
-$${E[score]}_{j} = \sum_{i}\{{E[score]}_{i} | i \in L_{j}\}$$
-
-Finally, we measure visual ability by:
-
-$${Visual Ability}_{j} = {score}_{j}-{E[score]}_{j}$$.
-
-Comparing these to the estimated random effects $\hat\gamma$ in Figure~\ref{random2} shows that there is a strong correlation between visual ability and the random effects, further validating the model.
-
-\begin{figure}[htbp]
-\centering
-<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
-ability <- merge(full.sub[,c("subj_id", "pic_id", "score", "weights")], bylineup.sub[c("pic_id", "percent")], by="pic_id")
-ability <- summarise(group_by(ability, subj_id), actual=sum(score), estimated=sum(percent*weights))
-ability$diff <- ability$actual - ability$estimated
-ability.re <- merge(ability, randomeffects, by="subj_id")
-ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs(x="Visual Ability", y="Random Effects")
-@
-\caption{Visual ability plotted against random effects.
-The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
-\label{random2}
-\end{figure}
-
 \iffalse
@@
 
-\begin{figure}[htbp]
-\centering
-<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-location <- read.csv("location.csv")
-map.dat <- map_data("world")
-ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
-@
-\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
-\label{map}
-\end{figure}
-
 \end{document}

### 2016-01-13 paper/Extra Material.R

@@ -10,60 +10,7 @@ extreme <- merge(extreme, full.sub[c(9,11)], by=0)
 #investigate pic_id = 61 name = s459.... => true plot in 18 looks alright, but most people picked plot 17 with cor = 0.24. 17 looks very good
 #if you ignore the outliers. Two chose reason 4, one chose reason 1. Last person selected 7 & 20, neither plot is especially interesting.
 
-#Subject evaluations#
-subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights*attempts), attempt = sum(attempts), detect = sum(correct))
-subjplot <- mutate(subjplot, byeval = detect/eval, byattempt = detect/attempt)
-#one subject has no successful evaluations: Saw ten lineups and made ten selections. Harder lineups: low correl scatter, negative lines.
-#did see 0.7 scatter and two 0.9 lines which should be reasonable chance to get correct (pic_id = 181, 140, 44)
-#was only person to get 0.7 scatter wrong (sd4V..., true = 12, chose 1, reason 1), but 0.7 is not high enough to get really large residuals.
-#11/13 detected line 140 (levq...., true = 4, chose 17, reason 2, this one is really easy) and 4/7 deteced in line 44. (lYbEZ, picked 19, true = 3, reason=2)
-#subjects with one successful detection tended to have negative lines and low correlation scatters
-ggplot(data=subjplot, aes(x=byattempt)) + geom_histogram(binwidth=0.05) + ylim(0,30) #proportion: detection to attempts
-#mean 0.37
-ggplot(data=subjplot, aes(x=byeval)) + geom_histogram(binwidth=0.05) + ylim(0,30) # proportion: detection to lineup evals.
-#mean 0.40
 
-#Choice plots#
-full.sub$choice_reason <-sapply(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+"), "[[", 1)
-ggplot(data=full.sub) + geom_bar(aes(x=choice_reason))
-choice <- full.sub %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
-choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
-choice$correct <- factor(choice$correct, labels=c("Did not detect", "Detected"))
-ggplot(data=choice, aes(x=choice_reason, fill=correct)) + geom_bar() + facet_grid(design~rsgn) +
[…]

### 2016-01-13 paper/supplementary.Rnw
supplementary material

--- 2016-01-13 paper/additional material.Rnw
@@
 
-\title{Additional Material}
+\if0\blind
+{
+\author{Nathaniel Tomasetti and Dianne Cook\\
+Department of Econometrics and Business Statistics, Monash University\\
+}
+  \maketitle
+} \fi
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation}
+\end{center}
+  \medskip
+} \fi
 
@@
 
-\section{Subject Details}
+\section{Subject Demographics}
 
@@
 
+\section{Descriptive Statistics for Individuals}
+
 \begin{figure}[htbp]
 \centering
-<<choices, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<choices, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 full.sub$choice_reason <-sapply(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+"), "[[", 1)
@@
 \centering 
-<<avscore, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=LSE>>=  
+<<avscore, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>= 
 subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
@@
 
-\section{Multiple Selection}
-
 \begin{figure}[htbp]
 \centering
-<<multiselect, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<multiselect, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
 m.freq <- data.frame(table(m.select$pic_id))
-m.lineup <- merge(m.freq, Turk, by.x="Var1", by.y="pic_id")
+m.lineup <- merge(m.freq, bylineup, by.x="Var1", by.y="pic_id")
 ggplot(data=m.lineup) + geom_bar(aes(x=r, fill=plot), binwidth=0.2) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
@@
 \begin{table}
-\caption{Number of selection per lineup evaluation. The subject with the most selections is reported }
+\caption{Number of selections per lineup evaluation. The subject with the most selections is reported.}
 \label{tabattempts}
-<<tabattempts, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<tabattempts, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 attempts <- summarise(group_by(full.sub, subj_id, pic_id), attempts = max(attempts))
@@
 \centering
-<<multiattempts, dependson='setup' =, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<multiattempts, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
 ggplot(data=attempts) + geom_bar(aes(x=as.factor(attempts))) + labs(x="Number of selections", y=NULL)

### 2016-01-13 paper/paper.Rnw
Redone fullsub

@@
 Lineups \citep{Majumder:2013, Wickham} allow the visual inference power of a graphic display to be evaluated, and were used to compare the plot designs.
-We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the decision for 2089 lineup evaluations.
+We collected data on the detection rate of correlated pairs of AR(1) simulations, the time required and the confidence of the selection for 1600 lineup evaluations.
 The results show that the scatter plot is both the faster and more powerful plot design, despite its inability to display the time dimension.
@@
 
-full$subj_id <- factor(full$ip_address, labels=1:207)
-y <- data.frame(table(full$subj_id))
-keep <- y$Var1[y$Freq>9]
-##Recording who evaluated at least ten lineups. Done prior to splitting multiple selections
-
 full <- full %>% mutate(design = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"),
@@
                          ##count number of selections in response_0 column
-                         weights = 1/attempts) %>%
-  transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0) %>%
+                         weights = 1/attempts)
+
+full$subj_id <- factor(full$nick_name, labels=1:212)
+y <- data.frame(table(full$subj_id))
+keep <- y$Var1[y$Freq>9]
+full.sub <- filter(full, subj_id %in% keep) 
+full.sub <- Reduce(rbind, by(full.sub, full.sub$subj_id, head, n=10))
+##Remove subjects with less than 10 evaluations. If more than 10, keep only first 10.
+
+full <- full %>%  transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0) %>%
   ##split multiple selections in response_0 into different rows
@@
 
-full.sub <- filter(full, subj_id %in% keep) 
-##keeping only subjects with >0 evaluations in actual modelling to better estimate random effects
+full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0) %>%
+  mutate(correct = ifelse(response_0 == actual_location, 1, 0), 
+         score = ifelse(correct==1, (20-attempts)/19, 0)) %>%
+  select(subj_id, correct, score, r, rsgn, rabs, smoothed, n, design, log_time, response_0,
+         actual_location, pic_id, choice_reason, attempts, weights, alpha, conf_level)
 
@@
 The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true.
-If an impartial observer asked to pick the plot that is different from the rest, picks the data plot this suggests there exists a pattern that is not the result of chance, and the null hypothesis is rejected.
+If an impartial observer asked to select the plot that is most different from the rest, does select the data plot this suggests there exists a pattern that is not the result of chance, and the null hypothesis is rejected.
 For a scatter plot, this departure from the null, might be a single outlier, or few outliers, a non-linear pattern, or clusters, which the human eye detects as more different in this data plot than in any of the nulls.
@@
 The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
-However, there is the possibility that the true plot was picked by chance, that we have committed a Type 1 error in the lineup test.
+However, there is the possibility that the true plot was selected by chance, that we have committed a Type 1 error in the lineup test.
 For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
@@
 We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was picked at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:
+A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was selected at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:
 
@@
 
-If all $K$ observers pick the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
+If all $K$ observers select the same plot out of a lineup of twenty, we result in a p-value as small as $1/m^K$ .
 We used $m= 20$, giving us a significance level (and Type 1 error rate) of $\alpha = 0.05$.
@@
 
-Essentially, the more time our subjects pick the correct plot out of the lineup, the more confident we are that the real plot was visually significantly different to the nulls; and that the type of graphic involved has the power to display association.
+Essentially, the more time our subjects select the true plot out of the lineup, the more confident we are that that plot was visually significantly different to the nulls; and that the type of graphic involved has the power to display association.
 The lineup effectively allows us to conduct a hypothesis test on visual features, with the competing hypotheses:
@@
 As the power of a statistical test is defined as the probability of rejecting $H_{0}$ when it is false, the power of a lineup test is viewed as the probability of detecting the true plot.
-We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ correctly detected the true plot out of the lineup.
+We approximate the power as $\hat\pi = x/K$, where $x$ observers out of $K$ detected the true plot out of the lineup.
 We can them estimate the power difference of competing lineups, $\hat\pi_{1} - \hat\pi_{2}$. An $\alpha \times 100\%$ confidence interval is calculated as \citep{Hofmann:2012}:
@@
 
-Some individuals may have a better natural ability at detecting the correct correlation pattern in the graphics, but by having each participant viewing multiple lineups and each lineup being viewed by multiple different participants, this can be controlled via a random effect variable in the model.
+Some individuals may have a better natural ability at detecting a correlation pattern in the graphics, but by having each participant viewing multiple lineups and each lineup being viewed by multiple different participants, this can be controlled via a random effect variable in the model.
 We recruited participants through Amazon's Mechanical Turk, where the lineup protocol has been applied to a number of problems in prior papers \citep{Heer, Majumder:2013, Wickham}.
@@
 True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$.
-Pilot testing found that the rate of picking the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly picking the true plot, for pairs with correlation below $\rho = 0.3$.
+Pilot testing found that the rate of selecting the true data plot fell to approximately $5\%$, the $1/m$ rate of randomly selecting the true plot, for pairs with correlation below $\rho = 0.3$.
 Further to this, relationships weaker than this point are often of little interest to analysts.
@@
 Each subject did not see the same data more than once.
-Subjects were asked to pick the plot in the lineup that has the strongest association, with basic examples of negatively and positively correlated plots.
+Subjects were asked to select the plot in the lineup that has the strongest association, with basic examples of negatively and positively correlated plots.
 They also answered with reasons for choosing the plot, and the confidence that they have that this really is the plot showing the strongest correlation.
-Each subject evaluated 10 lineups, plus a further two trials to eliminate people attempted to game the system.
-If a trial was not answered correctly, the subject could not evaluate the remaining lineups.
-If it was answered correctly, the trial data was discarded and the subject moved onto the rest of the lineups.
+Each subject evaluated 10 lineups, plus a further two simple trials to eliminate people attempted to game the system.
+If the true plot was not detected in the trial, the subject could not evaluate the remaining lineups.
+If it was detected, the trial data was discarded and the subject moved onto the rest of the lineups.
 Earlier work by \citet{Majumder:2014} has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
 We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
-In total there were 2088 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations leaving us with 1684 lineup evaluations for the analysis.
+In total there were 2091 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations.
+Some subjects chose to participate in the experiment more than once, so we removed all data from repeated trials by these subjects.
+We were left with 1600 evaluations by 160 subjects.
 
-Subjects were allowed to make multiple selections from each lineup evaluated, additional selections on lineup $i$ by subject $j$ are penalised according to a score defined by:
+Subjects were allowed to make multiple plot selections from each lineup evaluated, additional selections on lineup $i$ by subject $j$ are penalised according to a score defined by:
+
 \begin{equation}
@@
 However, for the model fit to power, multiple selections are seperated and given a weighting equal to to inverse of the number of selections made in that lineup evaluation.
-This increases our data from 1684 to 1849 evaluations.
+This increases our data from 1600 to 1762 evaluations.
 
@@
 <<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
+ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Detected", x="Correlation") + theme(legend.position="bottom")
 @
@@
 The time data was highly positively skewed with a maximum of 1760 seconds so the variable was log-transformed.
-We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a choice.
-Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to pick the true plot and 32.6 seconds if a null was selected.
-The scatter plots performed much better, with the true plot picked 52.9\% of the time, with an average of 22.7 seconds required to make a decision.
+We found that the subject could identify the true plot out of the overlaid line graph 33.6\% of the time and it took on average $e^{3.442}$ = 31.2 seconds to make a selection.
+Successful detections were slightly faster than unsuccessful decision, taking on average 28.7 seconds to select the true plot and 32.6 seconds if a null was selected.
+The scatter plots performed much better, with the true plot selected 52.9\% of the time, with an average of 22.7 seconds required to make a decision.
 If the true plot was detected it took only 18.6 seconds, and 29.3 seconds for a null plot selection.
@@
 
-Subjects were asked to give a number from 1 to 5 to rate how confident they were in there choice (1 - most confident, 5 - least confident, average = 3.34).
+Subjects were asked to give a number from 1 to 5 to rate how confident they were in their selection (1 - most confident, 5 - least confident, average = 3.34).
 There were no statistically significant differences between plot designs, however subjects were slightly less confident on true plot detections (by $0.25 \pm  0.16, p=0.002$) and every one percent extra time required to make a decision increased confidence (by $0.19 \pm 0.07, p<0.001$).
-One possible explanation is that subjects often focused on a null plot and picked it confidently.
+One possible explanation is that subjects often focused on a null plot and selected it confidently.
 This would suggest that the null plots often produced visually interesting features subjects assumed were caused by the association between plots.
@@
   \item The lineup was more difficult and it took longer to determine which plot was best, reducing confidence.
-  \item The subject picked a plot relatively easily but spent extra time going over its details and making comparisons to alternative plots, increasing confidence.
+  \item The subject selected a plot relatively easily but spent extra time going over its details and making comparisons to alternative plots, increasing confidence.
 \end{enumerate}
@@
 \item $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type,
-\item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 157$.
+\item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 160$.
 \end{itemize}
@@
 \caption{Weighted Mixed Effects Logit Model for the predicted power of plot design.
-Sample size for scatter plot evaluations: 900, overlaid lines: 949, total: 1849.
+Sample size for scatter plot evaluations: 859, overlaid lines: 903, total: 1762.
 Default plot design is overlaid lines.
@@
 
-\subsection{Modelling time taken}
+\subsection{Modelling Time Taken}
 
@@
 Whilst this is shown, the overlaid line graph did not have a corresponding increase in confidence compared to the scatter plot; so research into similar confidence-boosting visual anomalies that can occur in scatter plots would be necessary to further investigate the law.
-Diaconis's Magical Thinking was supported by the finding on the relationships between time, confidence and correctness.
-A more difficult true plot may be picked rarely and require a longer to make a decision; resulting in decreased subject confidence.
-However, we found that time increases confidence but reduces correctness.
-Highly confident, but incorrect decisions, suggests subjects may have been distracted by strong, spurious patterns in the null plots.
+Diaconis's Magical Thinking was supported by the finding on the relationships between time, confidence and true plot detection.
+A more difficult true plot may be selected rarely and require a longer to make a decision; resulting in decreased subject confidence.
+However, we found that time increases confidence but reduces the detection rate.
+Highly confident null selections suggests that subjects may have been distracted by strong, spurious patterns in the null plots.
 If a data analyst is required to use overlaid line graphs in their work, substantial improvements to power and hence readability can be made by smoothing time series data if the sample size is long enough for these techniques to become feasible.
@@
 
-\section{Appendix}
-
-\begin{table}[htp]
-\caption{The true plot detections / attempts for each lineup in the data collection. Results ignore multiple selection weighting. True plot correlation is on the x axis and t is on the y axis.}
-\label{tab3}
-\centering
-<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-turk.tbl <- select(bylineup, design, smoothed, r, n, correct, evals)
-turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$evals, sep="/")
-turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
-scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Scatterplot")$result, ncol=8))
-scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Scatterplot")$result, ncol=8))
-line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Overlaid Lines")$result, ncol=8))
-line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Overlaid Lines")$result, ncol=8))
-colnames(scatter.uns) <- c("Unsmoothed Scatterplot", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(scatter.sm) <- c("Smoothed Scatterplot","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.uns) <- c("Unsmoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-colnames(line.sm) <- c("Smoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
-kable(scatter.uns, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables2, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(scatter.sm, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables3, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(line.uns, format="latex")
-@
-\end{table}
-\begin{table}
-\centering
-<<answertables4, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-kable(line.sm, format="latex")
-@
-\end{table}
-
 \end{document}

### 2016-01-13 paper/supplementary.Rnw
Redone fullsub

@@
 
+\title{\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation
+
 \if0\blind
@@
 
-\section{Descriptive Statistics for Individuals}
+\section{Descriptive Statistics of Evaluations}
+
+\begin{table}[htp]
+\caption{The true plot detections / evaluations for each lineup in the data collection.
+Lineups with the same factor combination are seperated by row.
+Results ignore multiple selection and any removed evaluations.
+True plot correlation is on the x axis and t is on the y axis.}
+\label{tabeval}
+\centering
+<<answertables, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+turk.tbl <- select(bylineup, design, smoothed, r, n, correct, evals)
+turk.tbl$result <- paste(round(turk.tbl$correct, 0), turk.tbl$evals, sep="/")
+turk.tbl <- turk.tbl[order(turk.tbl$r, turk.tbl$n),]
+scatter.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Scatterplot")$result, ncol=8))
+scatter.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Scatterplot")$result, ncol=8))
+line.uns <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==0 & design=="Overlaid Lines")$result, ncol=8))
+line.sm <- data.frame(c(12, "", "", 24, "", "", 48, "", "", 96, "", ""), matrix(subset(turk.tbl, smoothed==1 & design=="Overlaid Lines")$result, ncol=8))
+colnames(scatter.uns) <- c("Unsmoothed Scatterplot", "-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(scatter.sm) <- c("Smoothed Scatterplot","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.uns) <- c("Unsmoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+colnames(line.sm) <- c("Smoothed Overlaid Lines","-0.9", "-0.7", "-0.5", "-0.3", "0.3", "0.5", "0.7", "0.9")
+kable(scatter.uns, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables2, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(scatter.sm, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables3, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(line.uns, format="latex")
+@
+\end{table}
+\begin{table}
+\centering
+<<answertables4, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+kable(line.sm, format="latex")
+@
+\end{table}
 
@@
 <<choices, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-full.sub$choice_reason <-sapply(strsplit(as.character(unlist(full.sub$choice_reason)), "[^0-9]+"), "[[", 1)
+full$choice_reason <-sapply(strsplit(as.character(unlist(full$choice_reason)), "[^0-9]+"), "[[", 1)
 ggplot(data=full.sub) + geom_bar(aes(x=choice_reason))
-choice <- full.sub %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
+choice <- full %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
 choice$rsgn <- factor(choice$rsgn, labels=c("r -","r +"))
@@
 @
+\caption{Reasons given for plot selection}
+\label{reasons}
+\end{figure}
 
@@
 @
+\caption{Average score of subjects included in modelling}
+\label{avscore}
+\end{figure}
 
@@
 m.lineup <- merge(m.freq, bylineup, by.x="Var1", by.y="pic_id")
-ggplot(data=m.lineup) + geom_bar(aes(x=r, fill=plot), binwidth=0.2) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
-ggplot(data=subset(m.lineup, smoothed==1)) + geom_bar(aes(x=r, fill=plot), binwidth=0.2) + facet_wrap(~n)
+ggplot(data=m.lineup) + geom_bar(aes(x=factor(r), fill=design), binwidth=0.2) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
 @
+\caption{Characteristics of lineups with at least one subject making multiple selections}
+\label{lineupmulti}
+\end{figure}
 
-\begin{table}
+\begin{table}[htp]
 \caption{Number of selections per lineup evaluation. The subject with the most selections is reported.}
@@
 @
-\caption{Histogram of table~\ref{attempts}}
-\label{attempts}
+\caption{Histogram of table~\ref{tabattempts}}
+\label{attemptshist}
 \end{figure}
@@
 
-\begin{figure}
+\begin{figure}[htbp]
   \centering

### 2016-01-13 paper/paper.Rnw
Few more fixes

@@
 #sasadata <- mutate(sasdata, rabs = as.numeric(rabs),
-#                   subj_id = as.numeric(subj_id)
+#                   subj_id = as.numeric(subj_id),
 #                   design = ifelse(design=="Scatterplot", 1, 0))
-#write.csv(subsetted, "sas/sasdata.csv", row.names = FALSE)
-##The code that is exported to fit the model into SAS, see sas/model fit.sas
+#write.csv(sasdata, "sas/sasdata.csv", row.names = FALSE)
+#The code that is exported to fit the model into SAS, see sas/model fit.sas
 
-# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
+#pred <- data.frame(expand.grid(rabs=seq(0, 0.9, 0.01), t=c(12, 24, 48, 96), alpha=log(.05/0.95), rsgn=c(0,1),
 #                                      smoothed=c(0,1), design=c(0,1),
-#                                      subj_id=full.sub$subj_id))
+#                                      subj_id=1:160))
 #pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)

### 2016-01-13 paper/supplementary.Rnw
Few more fixes

@@
 
-\title{\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation
+\title{\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation}
 
@@
 \centering
-<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=3>>=
 location <- read.csv("location.csv")
@@
 full$choice_reason <-sapply(strsplit(as.character(unlist(full$choice_reason)), "[^0-9]+"), "[[", 1)
-ggplot(data=full.sub) + geom_bar(aes(x=choice_reason))
 choice <- full %>% transform(choice_reason= strsplit(as.character(choice_reason), "")) %>% unnest(choice_reason)
@@
 
-\begin{table}[htp]
+\begin{table}[htbp]
 \caption{Number of selections per lineup evaluation. The subject with the most selections is reported.}
-\label{tabattempts}
+\label{table}
 <<tabattempts, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
@@
 @
-\caption{Histogram of table~\ref{tabattempts}}
+\caption{Histogram of table~\ref{table}}
 \label{attemptshist}
@@
   \centering
-<<random, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+<<random, dependson='setup', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 randomeffects <- summarise(group_by(prediction, subj_id), re=unique(diff))

### 2016-01-13 paper/paper.Rnw
SAS refit

--- 2016-01-13 paper/paper.Rnw
@@
 full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
-prediction <- read.csv("sas/predfit.csv") ##Large file, very slow, fitted data from SAS
+prediction <- read.csv("sas/predfit.csv") ##fitted data from SAS
 time.fit <- read.csv("sas/time.csv") ##coefficients for time model from SAS
@@
 #sasdata <- select(full.sub, rabs, n, alpha, rsgn, smoothed, design, subj_id, weights, correct, log_time)
-#sasadata <- mutate(sasdata, rabs = as.numeric(rabs),
-#                   subj_id = as.numeric(subj_id)
+#sasdata <- mutate(sasdata, rabs = as.numeric(rabs),
+#                   subj_id = as.numeric(subj_id),
 #                   design = ifelse(design=="Scatterplot", 1, 0))
-#write.csv(subsetted, "sas/sasdata.csv", row.names = FALSE)
-##The code that is exported to fit the model into SAS, see sas/model fit.sas
+#write.csv(sasdata, "sas/sasdata.csv", row.names = FALSE)
+#The code that is exported to fit the model into SAS, see sas/model fit.sas
 
-# data.frame(expand.grid(rabs=rabs, t=t, alpha=log(.05/0.95), rsgn=c(0,1),
+#pred <- data.frame(expand.grid(rabs=seq(0, 0.9, 0.01), t=c(12, 24, 48, 96), alpha=log(.05/0.95), rsgn=c(0,1),
 #                                    smoothed=c(0,1), design=c(0,1),
-#                                      subj_id=full.sub$subj_id))
+#                                      subj_id = as.numeric(as.character(unique(full.sub$subj_id)))))
 #pred <- subset(pred, smoothed==0 | smoothed == 1 & t ==48 | smoothed ==1 & t ==96)
-#write.csv(pred, "sas/predgrid.csv", row.names=FALSE) # these lines are done, but its a large file so commented out for knitr
+#write.csv(pred, "sas/predgrid.csv", row.names=FALSE)
 #Code exported to SAS to predict new values
@@
 <<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-predplot <- select(prediction, r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf)
-predplot <- mutate(prediction, RandomEf = exp(RandomEf)/(1+exp(RandomEf)), 
+predplot <- prediction %>% filter(is.na(detected)) %>%
+            select(r, t, rsgn, smoothed, design, subj_id, RandomEf, FixedEf) %>%
+            mutate(RandomEf = exp(RandomEf)/(1+exp(RandomEf)), 
                    FixedEf = exp(FixedEf)/(1+exp(FixedEf)))
@@
  & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \beta_{13} x_{6,m} + \beta_{14} x_{5_l} \* x_{6,m} \gamma z_{s} + \epsilon_{ijklmt} \nonumber 
+ & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \nonumber 
 \end{eqnarray}
 
-The structure of the model ~\ref{timemodel} is similar to that in Equation~\ref{modelpower} except the link function is replace with the identity function, $\psi(x) = x$, and a new variable, $x_{6,m}$, is introduced with $m = \{0,1\}$ indicating if the subject detected the true plot, where $m = 0$ is a failure to detect the true plot.
+The structure of the model ~\ref{timemodel} is similar to that in Equation~\ref{modelpower} except the link function is replace with the identity function, $\psi(x) = x$.
 

### 2016-01-13 paper/supplementary.Rnw
SAS refit

@@
 
-Analysing the level one residuals, $\hat{\epsilon_{i}}$ and level two residuals, $\hat{\gamma_{i}}$ allows us to test the model fit as described \citet{Loy}.
+Analysing the level one residuals, $\hat{\epsilon_{i}}$, and level two residuals, $\hat{\gamma_{i}}$, allows us to test the model fit as described \citet{Loy}.
 There are some negative extreme outliers but the distributional assumption appears to hold.

### 2016-01-14 paper/paper.Rnw
Conclusion

@@
 
-The structure of the model ~\ref{timemodel} is similar to that in Equation~\ref{modelpower} except the link function is replace with the identity function, $\psi(x) = x$.
+Other than the removal of the link functon, the structure of the model identical to that in Equation~\ref{modelpower}.
 
@@
 
+Power
+-In low sample sizes, both designs have similarly weak performance - 
+-Relationships between t=12 is dubious at best, requires $\rho$ > 0.7 to have more than 50\% power
+-In higher sample sizes, the scatterplot improves dramatically while the line plot remains unchanged
+-The line plot has much weaker performance in negative correlation than positive
+-If possible, smooth the lines to improve performance
+-If negative association is expected, flip one line to make it positive
+
+Time
+-The scatterplot is faster than the lines at high correlation
+-Smoothing and sample size have no real effect
+
+Confidence
+-People could be distracted by spurious patterns in the null plots
+-These were very confident selections
+-The data analyst needs to understand what kinds of patterns can be generated randomly
+
+Overall
+-Despite hiding trends etc., the simplistic scatter plot seems to be best suited to these tasks.
+
+\iffalse
 This paper showed that the scatterplot is better than the overlaid lines for displaying two time series when the purpose is to examine the association between the two series.
@@
 If a data analyst is required to use overlaid line graphs in their work, substantial improvements to power and hence readability can be made by smoothing time series data if the sample size is long enough for these techniques to become feasible.
+\fi
 

### 2016-01-20 paper/paper.Rnw
working on intro

@@
 
-In order to work with data, it first must be understood. Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first.
-This is especially so today, for vast databases that have been assembled in the big data era, that now need the data scientist to unravel the meaning of the numbers. 
-Without preset hypotheses to test, the power of statistical inference is impotent, and without hypotheses the data analyst can stumble blindly trying to build up models of structure in the data. 
-To understand data requires good visualisation.
+%In order to work with data, it first must be understood. Statistical inference requires hypotheses to be established prior to data collection, but often data is collected first.
+%This is especially so today, for vast databases that have been assembled in the big data era, that now need the data scientist to unravel the meaning of the numbers. 
+%Without preset hypotheses to test, the power of statistical inference is impotent, and without hypotheses the data analyst can stumble blindly trying to build up models of structure in the data. 
+%To understand data requires good visualisation.
 This idea was formed as early as the 18th century, when William Playfair institutionalized the then revolutionary idea of graphing government and economic data.
-Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information \citep{Sachs}.
+%Far easier than reading tables of numbers, these ideas were powerful, and by providing the basic building blocks for plotting statistical data, his graphic designs became a conduit for the communication of otherwise complex information \citep{Sachs}.
 Since then, advances in computing power have allowed statistical graphics to flourish, spearheaded by \citet{Tukey} in 1965 into the new domain of exploratory data analysis (EDA) \citep{Friendly}.
-EDA can be thought of as a well understood \citep{Cleveland, Vanderplas} set of tools and techniques required to visualise information, to physically see what the data contains.
+%EDA can be thought of as a well understood \citep{Cleveland, Vanderplas} set of tools and techniques required to visualise information, to physically see what the data contains.
 In particular, it incorporates a free roaming approach, where the analyst is able to explore structure to find whatever relationships and structure exists within.
-Critically, the analyst does not have to have any pre-conceived ideas or hypotheses -- they are not specifically looking for any one particular thing.
+%Critically, the analyst does not have to have any pre-conceived ideas or hypotheses -- they are not specifically looking for any one particular thing.
 EDA emphasises letting the data inform us and can lead to the discovery of otherwise unexpected relationships, many of which may seem to become completely obvious after discovery.
-With the knowledge provided by EDA, ideas are generated about what relationships between variables may potentially exist.
+%With the knowledge provided by EDA, ideas are generated about what relationships between variables may potentially exist.
 This then enables the analyst to use these new hypotheses upon which to apply classical inference and rigorously conduct tests with new data.
-EDA is also related to the field of model diagnostics (MD), where a model can be continuously refined through the visualisation of its fit, its residuals, and the interactions with its variables.
+%EDA is also related to the field of model diagnostics (MD), where a model can be continuously refined through the visualisation of its fit, its residuals, and the interactions with its variables.
 Both EDA and MD follow the same framework: Visualise the data, look for patterns that suggest an underlying relationship, and if one is found implement it into the model then continue the exploration of the data.
-It has long been thought that EDA and statistical inference were worlds apart, but recent work in \citet{Buja} and \citet{Majumder:2013} bridges this chasm.
+%It has long been thought that EDA and statistical inference were worlds apart, but recent work in \citet{Buja} and \citet{Majumder:2013} bridges this chasm.
 Framing a data plot as a test statistic, which when compared to plots of null data, places EDA into the statistical inference recipe.
-The null hypothesis underlying a particular plot, is generically that there is no pattern, and particular types of plots implicitly regulate what “no  pattern” means.
+%The null hypothesis underlying a particular plot, is generically that there is no pattern, and particular types of plots implicitly regulate what “no  pattern” means.
 For example, a scatter plot of two variables is used to explore for some sort of association, so the implicit hypothesis is that there is no association between the two variables.
-The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association. 
-The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true.
+%The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association. 
+%The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true.
 If an impartial observer asked to select the plot that is most different from the rest, does select the data plot this suggests there exists a pattern that is not the result of chance, and the null hypothesis is rejected.
-For a scatter plot, this departure from the null, might be a single outlier, or few outliers, a non-linear pattern, or clusters, which the human eye detects as more different in this data plot than in any of the nulls.
+%For a scatter plot, this departure from the null, might be a single outlier, or few outliers, a non-linear pattern, or clusters, which the human eye detects as more different in this data plot than in any of the nulls.
 This is the reason why visualisation remains important today, human eyes can detect patterns which would not be detected mathematically.
 But eyes need calibration, which the lineup protocol provides.
-On its own, with a single plot, because of random sampling in collecting data, it is easy for the analyst to imagine a pattern when no real structure exists in the population.
+%On its own, with a single plot, because of random sampling in collecting data, it is easy for the analyst to imagine a pattern when no real structure exists in the population.
 Visual skills of an observer in EDA is prone to Type I error, where the null hypothesis is rejected when it is actually true, caused by the inherently random formation of patterns when visualised which are attributed to structure rather than chance.
@@
   \includegraphics[width=0.5\textwidth]{cnyline}
-  \caption{Exchange rates for Australian dollars and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid time series.
+  \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.
 (Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
@@
     \includegraphics[width=0.45\textwidth]{nzdline}}
-  \caption{Exchange rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
-Which one can you read the correlation from more accurately?
+  \caption{Cross-rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
+From which plot is correlation more accurately read?
 Point-wise correlation between the two series, ignoring autodependence, is 0.64.}

### 2016-01-20 paper/paper.Rnw
first two paragraphs of intro

@@
   \begin{center}
-    {\LARGE\bf Comparing the Power of Plot Designs to Reveal Correlation}
+    {\LARGE\bf What is the Best Way to Display Two Time Series to Read Association?}
 \end{center}
@@
 \noindent
-{\it Keywords:}  Lineups, visual inference, power comparison, scatter plot, line graph, data visualisation, visual analytics, time series, temporal data
+{\it Keywords:}  data visualisation, statistical graphics, visual inference, power comparison, scatter plot, line graph, visual analytics, time series, temporal data, finance, econometrics
 \vfill
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{cnyline}
-  \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.
-(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
-  \label{cny}
-\end{figure}
-
-In this paper we utilised the methods in visual inference to study plot design to investigate the relationship between two temporal variables.
-There is a substantial literature suggesting that scatter plots are the appropriate display to read association \citep{Cleveland, Harrison, Li}, however, when the two variables are temporal it is common to display them as time series, drawn on the same plot.
-Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and there is a strong argument for its presence in financial data \citep{Sher, Turtle}.
-It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data.
+Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and in financial data \citep{Sher, Turtle}.
+It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data (Figure \ref{nzd}).
+In the field of statistics, the recommendation for exploring association visually is to use a scatterplot.
+There is substantial literature supporting this \citep{Cleveland, Harrison, Li}.
+However, it is still common that when the two variables are temporal they are displayed as overlaid time series (line graphs).
+In this paper we have utilized new methods for rigorously comparing plot design, using visual inference, to investigate which is the best display if reading association is important.
 
@@
 
-To decide on an appropriate display requires awareness of cognitive principles in psychology such as the Gestalt law of common fate described by \citet{Wertheimer}.
-This law describes a tendency in human perception to 'see' a positive relationship between two objects that briefly move together over time, leading people to believe that they stay linked and share a 'common fate' (Figure~\ref{cny}), even if no relationship exists.
-\citet{Lee} work finds that humans are able to perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that a graphic than can utilise a dimension for time may be able to use the information provided by temporal dependence and potentially outperform the graph types that do not, such as the scatter plot.
-However, keep in mind magical thinking \citep{Diaconis},   the desire to find any pattern may lead to a false discovery of positive correlation.
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{cnyline}
+  \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.
+(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
+  \label{cny}
+\end{figure}
+
+To determine on an appropriate display requires awareness of cognitive principles such as the Gestalt law of common fate \citep{Wertheimer}.
+People have a tendency to `see' a positive association between two objects that only briefly move together over time, leading them to believe, perhaps erroneously, that they are linked and share a `common fate' (Figure~\ref{cny}).
+To the contrary, \citet{Lee} suggests that people may better perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that overlaid line graphs, that utilize a time axis, may outperform the graph types that do not, such as the scatter plot.
+The magical thinking ideas described in \citet{Diaconis} would counter \citet{Lee}'s findings.
+Magical thinking means that humans have a strong desire to find patterns, even imaginary, or mirages, which may lead to a false discovery of positive correlation from a small period of synchrony.
+
 The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier.

### 2016-01-20 paper/paper.Rnw
four paragraphs of intro

@@
 To determine on an appropriate display requires awareness of cognitive principles such as the Gestalt law of common fate \citep{Wertheimer}.
-People have a tendency to `see' a positive association between two objects that only briefly move together over time, leading them to believe, perhaps erroneously, that they are linked and share a `common fate' (Figure~\ref{cny}).
+People have a tendency to `see' a positive association between two objects that only briefly move together over time, leading them to believe, perhaps erroneously, that they are linked and share a `common fate'.
 To the contrary, \citet{Lee} suggests that people may better perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that overlaid line graphs, that utilize a time axis, may outperform the graph types that do not, such as the scatter plot.
@@
 Magical thinking means that humans have a strong desire to find patterns, even imaginary, or mirages, which may lead to a false discovery of positive correlation from a small period of synchrony.
-
-The law then has two effects, if positive correlation exists in temporal data, the observer may be drawn to it easier.
-If positive correlation does not exist, the observer may be fooled into thinking it does anyway.
-Either effect would lead to a more confident rejection of the null than if the same data was displayed as a scatter plot.
-Figure~\ref{cny} highlights the dangerous interaction between 'magical thinking' and the Gestalt Law, the observer is drawn to the period from April 8 to April 22, and may falsely believe that the two currencies may continue to move together.
+Figure~\ref{cny} highlights the interaction between `magical thinking' and the Gestalt Law, the observer may be drawn to the period from April 8 to April 22, and may falsely believe that the two currencies are strongly tied.
 However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
-Incorrectly relying on two currencies, or many other sets of economic variables, being dependent on each other can be extremely dangerous for a firm.
-\citet{Robbins} is similarly critical of the time-based line graph, where it is argued that information from overlaid line graphs can easily be misleading (Figure~\ref{murder}).
-When both lines dive after 1993, it is natural to compare the horizontal distance and say that the difference between the lines in 1994 is small.
-However, this is incorrect, as looking at the small horizontal distance is comparing the lines at different dates.
-It is not obvious that the difference in 1994 is the second greatest difference across the entire range, only slightly smaller than in 1993.
-\citet{Vanderplas} examines this effect in more detail.
-It is dubbed the 'Sine Illusion', where the human eye is poor at interpreting lines with such a sharp slope.
-As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice.
+
+The law then has two effects.
+If positive correlation exists in temporal data, the observer may be drawn to it faster from an overlaid line graph display.
+If positive correlation does not exist, or only exists ephemerally, the observer may be fooled by the overlaid line graphs into thinking there is strong association.
+This latter effect is a problem.
+
+In practice we have lots of choices in how to present information in exploratory analyses or to communicate information.
+If linear correlation is the only aspect of the data of interest, we might only calculate the correlation between two series.
+However, if correlation is not the appropriate statistic, because the association is more complex, or if there are outliers, we can typically only discover this by making pictures.
+Learning how people read plots of data is an important component of data analysis.
+If information is to be communicated in graphical form it is good to know what we can confidently expect the audience to infer.
+
+There have been a few studies related to plotting temporal data.
+\citet{Robbins} writes a blog and consults on plotting financial data.
+She is critical of the overlaid line graph, because it is misleading if the purpose is to read the difference between the two series.
+Her arguments date back a long time to \citet{playfair,tufte,friendly}.
+It is well-established that we cannot read differences between two series accurately, that the difference should be plotted separately if that is the information of interest.
+
+\citet{Vanderplas} examines a related problem, dubbed the `sine illusion', where the human eye is poor at interpreting lines with sharp slopes.
+For time series, this means that people do not effectively read variance of seasonal series.
+
+Correlation study by the ... guys
+
+%As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice.
 However, in practice many data analysts are split between the two major alternatives for temporal data.

### 2016-01-20 paper/paper.Rnw
nearly done with intro

@@
 
-The law then has two effects.
+The law then, has two effects.
 If positive correlation exists in temporal data, the observer may be drawn to it faster from an overlaid line graph display.
@@
 It is well-established that we cannot read differences between two series accurately, that the difference should be plotted separately if that is the information of interest.
-
-\citet{Vanderplas} examines a related problem, dubbed the `sine illusion', where the human eye is poor at interpreting lines with sharp slopes.
-For time series, this means that people do not effectively read variance of seasonal series.
-
-Correlation study by the ... guys
+\citet{Vanderplas} examines a related problem, dubbed the ``sine illusion'', where the human eye is is fooled into erroroneously thinking that variation is larger at the peaks and troughs of seasonal time series.
+\citet{Vanderplas} developed a mathematical correction to the illusion.
+
+Reading correlation between two sets of non-temporal context numbers has been examined in many papers.
+In \citet{harrison2014ranking} different graphical displays of two sets of numbers are ranked of their effectiveness by Weber's Law \citep{webers} which provides a relationship between perception and the ``effect size'' in the data.
+For example, two series with correlation 0.9 should be proportionately easier to distinguish from noise than two series with correlation 0.3.
+The visual display should reflect this difference is ease.
+Early papers \citep{Cleveland1982,lane1985} showed that people read association from a scatterplot differently than they would linearly interpret a correlation coefficient.
 
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{murders}
-  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
+%\begin{figure}[htbp]
+%  \centering
+%  \includegraphics[width=0.5\textwidth]{murders}
+%  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
 The difference between the two represents murders outside of the city.
 What size would you judge the difference to be in 1994?}
-  \label{murder}
-\end{figure}
+%  \label{murder}
+%\end{figure}
 
@@
 Section 4 contains the results of the experiment and Section 5 discusses the implications of the findings.
+
+Visual Inference background next
 

### 2016-01-20 paper/paper.Rnw
done with intro

@@
 
-\title{\bf Comparing the Power of Plot Designs to Reveal Correlation}
+\title{\bf What is the Best Way to Display Two Time Series to Read Association?}
 
@@
 She is critical of the overlaid line graph, because it is misleading if the purpose is to read the difference between the two series.
-Her arguments date back a long time to \citet{playfair,tufte,friendly}.
+Her arguments date back a long time to Pfari, as reported in \citet{Tufte:1986} and \citet{friendly}.
 It is well-established that we cannot read differences between two series accurately, that the difference should be plotted separately if that is the information of interest.
@@
 \citet{Vanderplas} developed a mathematical correction to the illusion.
+\citet{Javed} examined the perception of many different sorts of features in multiple time series, and found that there was no optimal plot design for reading all of the different features tested.
+They did not examine perception of association between two series, though.
 
@@
 
-There has been other research on the perception of temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
+%There has been other research on the perception of temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
 Javed et al.
@@
 However there is a quandary, the visual features that improve local tasks are unsuited to global tasks and vice versa; so we must now find a form that is well suited to assess both types of tasks simultaneously.
-This paper addresses this deficiency.
-Section 2 describes the lineup method utilised to rigorously compare plot designs.
+
+This paper rigorously examines the perception of association between two time series in two different displays, and with an additional inquiry into whether smoothing of series affects the perception, using the lineup protocol.
+Section 2 describes the lineup protocol, and the background to its use for comparing the power of plot designs for different purposes.
 Section 3 explains the experimental design.
@@
 
-Visual Inference background next
-
-\section{The Lineup Protocol}
+\section{The lineup protocol for comparing plot designs}
 
@@
 
-\section{Experimental Design}
+\section{Experimental design}
 
@@
 
-\subsection{Detection, Time and Confidence}
+\subsection{Detection, time and confidence}
 
@@
 
-\subsection{Selection Analysis}
+\subsection{Selection analysis}
 
@@
 
-\subsection{Modelling Graphical Power}
+\subsection{Modelling graphical power}
 
@@
 
-\subsection{Modelling Time Taken}
+\subsection{Modelling time taken}
 

### 2016-01-20 paper/paper.Rnw
first two paragraphs of section 2

@@
 Both EDA and MD follow the same framework: Visualise the data, look for patterns that suggest an underlying relationship, and if one is found implement it into the model then continue the exploration of the data.
-%It has long been thought that EDA and statistical inference were worlds apart, but recent work in \citet{Buja} and \citet{Majumder:2013} bridges this chasm.
+% Section 2 describes the visual inference protocols which put EDA more firmly into the rigorous framework of statistical inference.
+
+Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and in financial data \citep{Sher, Turtle}.
+It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data (Figure \ref{nzd}).
+In the field of statistics, the recommendation for exploring association visually is to use a scatterplot.
+There is substantial literature supporting this \citep{Cleveland, Harrison, Li}.
+However, it is still common that when the two variables are temporal they are displayed as overlaid time series (line graphs).
+In this paper we have utilized new methods for rigorously comparing plot design, using visual inference, to investigate which is the best display if reading association is important.
+
+\begin{figure}[htbp]
+  \centering
+    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
+    \includegraphics[width=0.45\textwidth]{nzdline}}
+  \caption{Cross-rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
+From which plot is correlation more accurately read?
+Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
+  \label{nzd}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{cnyline}
+  \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.
+(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
+  \label{cny}
+\end{figure}
+
+To determine on an appropriate display requires awareness of cognitive principles such as the Gestalt law of common fate \citep{Wertheimer}.
+People have a tendency to `see' a positive association between two objects that only briefly move together over time, leading them to believe, perhaps erroneously, that they are linked and share a `common fate'.
+To the contrary, \citet{Lee} suggests that people may better perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that overlaid line graphs, that utilize a time axis, may outperform the graph types that do not, such as the scatter plot.
+The magical thinking ideas described in \citet{Diaconis} would counter \citet{Lee}'s findings.
+Magical thinking means that humans have a strong desire to find patterns, even imaginary, or mirages, which may lead to a false discovery of positive correlation from a small period of synchrony.
+Figure~\ref{cny} highlights the interaction between `magical thinking' and the Gestalt Law, the observer may be drawn to the period from April 8 to April 22, and may falsely believe that the two currencies are strongly tied.
+However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
+
+The law then, has two effects.
+If positive correlation exists in temporal data, the observer may be drawn to it faster from an overlaid line graph display.
+If positive correlation does not exist, or only exists ephemerally, the observer may be fooled by the overlaid line graphs into thinking there is strong association.
+This latter effect is a problem.
+
+In practice we have lots of choices in how to present information in exploratory analyses or to communicate information.
+If linear correlation is the only aspect of the data of interest, we might only calculate the correlation between two series.
+However, if correlation is not the appropriate statistic, because the association is more complex, or if there are outliers, we can typically only discover this by making pictures.
+Learning how people read plots of data is an important component of data analysis.
+If information is to be communicated in graphical form it is good to know what we can confidently expect the audience to infer.
+
+There have been a few studies related to plotting temporal data.
+\citet{Robbins} writes a blog and consults on plotting financial data.
+She is critical of the overlaid line graph, because it is misleading if the purpose is to read the difference between the two series.
+Her arguments date back a long time to Pfari, as reported in \citet{Tufte:1986} and \citet{friendly}.
+It is well-established that we cannot read differences between two series accurately, that the difference should be plotted separately if that is the information of interest.
+\citet{Vanderplas} examines a related problem, dubbed the ``sine illusion'', where the human eye is is fooled into erroroneously thinking that variation is larger at the peaks and troughs of seasonal time series.
+\citet{Vanderplas} developed a mathematical correction to the illusion.
+\citet{Javed} examined the perception of many different sorts of features in multiple time series, and found that there was no optimal plot design for reading all of the different features tested.
+They did not examine perception of association between two series, though.
+
+Reading correlation between two sets of non-temporal context numbers has been examined in many papers.
+In \citet{harrison2014ranking} different graphical displays of two sets of numbers are ranked of their effectiveness by Weber's Law \citep{webers} which provides a relationship between perception and the ``effect size'' in the data.
+For example, two series with correlation 0.9 should be proportionately easier to distinguish from noise than two series with correlation 0.3.
+The visual display should reflect this difference is ease.
+Early papers \citep{Cleveland1982,lane1985} showed that people read association from a scatterplot differently than they would linearly interpret a correlation coefficient.
+
+%As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice.
+However, in practice many data analysts are split between the two major alternatives for temporal data.
+The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.
+
+%\begin{figure}[htbp]
+%  \centering
+%  \includegraphics[width=0.5\textwidth]{murders}
+%  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
+The difference between the two represents murders outside of the city.
+What size would you judge the difference to be in 1994?}
+%  \label{murder}
+%\end{figure}
+
+%There has been other research on the perception of temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
+Javed et al.
+examined optimal ways to visualise local features, such as, which variable had the highest value at a given point in time, and global features, such as, comparing the size of the overall slope of many different variables.
+The finding was that a different graphical layout is optimal for each type of task.
+However, they did not test for the perception of correlation, which can be treated as both a local and a global feature within the graph.
+The slope of each variable must be compared at a point in time, and they must have a persistent relationship across at least the majority of the series.
+However there is a quandary, the visual features that improve local tasks are unsuited to global tasks and vice versa; so we must now find a form that is well suited to assess both types of tasks simultaneously.
+
+This paper rigorously examines the perception of association between two time series in two different displays, and with an additional inquiry into whether smoothing of series affects the perception, using the lineup protocol.
+Section 2 describes the lineup protocol, and the background to its use for comparing the power of plot designs for different purposes.
+Section 3 explains the experimental design.
+Section 4 contains the results of the experiment and Section 5 discusses the implications of the findings.
+
+\section{The lineup protocol for comparing plot designs}
+
+It has long been thought that EDA and statistical inference were worlds apart, but this is no longer true.
+\citet{Buja} and \citet{Majumder:2013} bridges the chasm.
 Framing a data plot as a test statistic, which when compared to plots of null data, places EDA into the statistical inference recipe.
-%The null hypothesis underlying a particular plot, is generically that there is no pattern, and particular types of plots implicitly regulate what “no  pattern” means.
-For example, a scatter plot of two variables is used to explore for some sort of association, so the implicit hypothesis is that there is no association between the two variables.
-%The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association. 
+Underlying any data plot, there is an implicit null hypothesis that there is ``no pattern''.
+Particular types of plots may regulate what no  pattern means.
+For example, a scatter plot is used to explore for association between the two variables, so the implicit null hypothesis is that there is no association between the two variables.
+The alternative hypothesis is that there is some association, although it is not required to specify precisely the type of association.
+The departure from the null might be a single outlier, or few outliers, a non-linear trend, or clusters.
+To test the null hypothesis the data plot is placed in a lineup with plots of null data, data generated from a scenario of no pattern.
+An impartial viewer is asked to select the plot that is the most different from the others.
+If they select the data plot as the most different, this counts as a rejection of the null hypothesis, and follow-up questions may determine the type of departure: clusters, outlier, nonlinearity in the case of a scatterplot.
+With multiple independent observers a $p$-value can be calculated to measure the statistical significance of the departure from the null.
+This process is called the ``lineup protocol''.
+
+The lineup protocol equips exploratory visual methods with a rigorous procedure that enables statistical inference.
+This is important because visualization remains essential for data analysis because human eyes can detect patterns which would not be detected mathematically.
+The lineup protocol provides the calibration of the human eyes to prevent us from erroneously jumping to conclusions about magic patterns.
+Earlier than the protocol proscription, we can find preliminary ideas along these lines, for example \citet{Daniel} recommends looking at plots of data from a null distribution for exploring two factor design experimental data.
+His paper provides 40 pages of plots from the null!
+He encourages data analysts to understand the patterns that might be seen in samples data drawn from populations without inherent structure.
+(\citet{Buja} calls this the Rorschach protocol.)
+
+%which the human eye detects as more different in this data plot than in any of the nulls.
 %The plot of the data is placed in a lineup of plots of null data, data generated assuming that the null hypothesis is true.
@@
 The mathematical rigor that governs classical inference relies on a framework that acknowledges and controls for the rate of Type I error.
-Section 2 describes the visual inference protocols which put EDA more firmly into the rigorous framework of statistical inference.
-
-Temporal dependence is present in many macroeconomic variables \citep{Nelson, Perron} and in financial data \citep{Sher, Turtle}.
-It is a common, but untested, belief that the inclusion of time in the graphics will increase the detail of information displayed and allow relationships to be examined more comprehensively, so many analysts utilise the overlaid line graph to examine their data (Figure \ref{nzd}).
-In the field of statistics, the recommendation for exploring association visually is to use a scatterplot.
-There is substantial literature supporting this \citep{Cleveland, Harrison, Li}.
-However, it is still common that when the two variables are temporal they are displayed as overlaid time series (line graphs).
-In this paper we have utilized new methods for rigorously comparing plot design, using visual inference, to investigate which is the best display if reading association is important.
-
-\begin{figure}[htbp]
-  \centering
-    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
-    \includegraphics[width=0.45\textwidth]{nzdline}}
-  \caption{Cross-rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
-From which plot is correlation more accurately read?
-Point-wise correlation between the two series, ignoring autodependence, is 0.64.}
-  \label{nzd}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{cnyline}
-  \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.
-(Both sets of values were standardized to bring them to the same scale for plotting.) What would you report the correlation between the two series to be?}
-  \label{cny}
-\end{figure}
-
-To determine on an appropriate display requires awareness of cognitive principles such as the Gestalt law of common fate \citep{Wertheimer}.
-People have a tendency to `see' a positive association between two objects that only briefly move together over time, leading them to believe, perhaps erroneously, that they are linked and share a `common fate'.
-To the contrary, \citet{Lee} suggests that people may better perceive a relationship between objects that move with temporal synchrony in accordance with the Gestalt Law, implying that overlaid line graphs, that utilize a time axis, may outperform the graph types that do not, such as the scatter plot.
-The magical thinking ideas described in \citet{Diaconis} would counter \citet{Lee}'s findings.
-Magical thinking means that humans have a strong desire to find patterns, even imaginary, or mirages, which may lead to a false discovery of positive correlation from a small period of synchrony.
-Figure~\ref{cny} highlights the interaction between `magical thinking' and the Gestalt Law, the observer may be drawn to the period from April 8 to April 22, and may falsely believe that the two currencies are strongly tied.
-However, over the entire series the linear correlation between the two currencies is -0.12, indicating there is no actual relationship.
-
-The law then, has two effects.
-If positive correlation exists in temporal data, the observer may be drawn to it faster from an overlaid line graph display.
-If positive correlation does not exist, or only exists ephemerally, the observer may be fooled by the overlaid line graphs into thinking there is strong association.
-This latter effect is a problem.
-
-In practice we have lots of choices in how to present information in exploratory analyses or to communicate information.
-If linear correlation is the only aspect of the data of interest, we might only calculate the correlation between two series.
-However, if correlation is not the appropriate statistic, because the association is more complex, or if there are outliers, we can typically only discover this by making pictures.
-Learning how people read plots of data is an important component of data analysis.
-If information is to be communicated in graphical form it is good to know what we can confidently expect the audience to infer.
-
-There have been a few studies related to plotting temporal data.
-\citet{Robbins} writes a blog and consults on plotting financial data.
-She is critical of the overlaid line graph, because it is misleading if the purpose is to read the difference between the two series.
-Her arguments date back a long time to Pfari, as reported in \citet{Tufte:1986} and \citet{friendly}.
-It is well-established that we cannot read differences between two series accurately, that the difference should be plotted separately if that is the information of interest.
-\citet{Vanderplas} examines a related problem, dubbed the ``sine illusion'', where the human eye is is fooled into erroroneously thinking that variation is larger at the peaks and troughs of seasonal time series.
-\citet{Vanderplas} developed a mathematical correction to the illusion.
-\citet{Javed} examined the perception of many different sorts of features in multiple time series, and found that there was no optimal plot design for reading all of the different features tested.
-They did not examine perception of association between two series, though.
-
-Reading correlation between two sets of non-temporal context numbers has been examined in many papers.
-In \citet{harrison2014ranking} different graphical displays of two sets of numbers are ranked of their effectiveness by Weber's Law \citep{webers} which provides a relationship between perception and the ``effect size'' in the data.
-For example, two series with correlation 0.9 should be proportionately easier to distinguish from noise than two series with correlation 0.3.
-The visual display should reflect this difference is ease.
-Early papers \citep{Cleveland1982,lane1985} showed that people read association from a scatterplot differently than they would linearly interpret a correlation coefficient.
-
-%As these problems are unique to the line graph, they argue that the scatter plot, which does not have any ability to display time, is the superior choice.
-However, in practice many data analysts are split between the two major alternatives for temporal data.
-The overlaid line graph is justified by its ability to present more information to the analyst; but many argue that it is this extra information that is misleading, and revert to the scatter plot for its strong non-temporal performance.
-
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics[width=0.5\textwidth]{murders}
-%  \caption{\citep{Robbins} Murders in New York State (Orange) and New York City (Blue).
-The difference between the two represents murders outside of the city.
-What size would you judge the difference to be in 1994?}
-%  \label{murder}
-%\end{figure}
-
-%There has been other research on the perception of temporal displays, such as in \citet{Javed}, but they do not examine perception of association.
-Javed et al.
-examined optimal ways to visualise local features, such as, which variable had the highest value at a given point in time, and global features, such as, comparing the size of the overall slope of many different variables.
-The finding was that a different graphical layout is optimal for each type of task.
-However, they did not test for the perception of correlation, which can be treated as both a local and a global feature within the graph.
-The slope of each variable must be compared at a point in time, and they must have a persistent relationship across at least the majority of the series.
-However there is a quandary, the visual features that improve local tasks are unsuited to global tasks and vice versa; so we must now find a form that is well suited to assess both types of tasks simultaneously.
-
-This paper rigorously examines the perception of association between two time series in two different displays, and with an additional inquiry into whether smoothing of series affects the perception, using the lineup protocol.
-Section 2 describes the lineup protocol, and the background to its use for comparing the power of plot designs for different purposes.
-Section 3 explains the experimental design.
-Section 4 contains the results of the experiment and Section 5 discusses the implications of the findings.
-
-\section{The lineup protocol for comparing plot designs}
 

### 2016-01-20 paper/paper.Rnw
lots more to do in section 2, but stopping for now

@@
 The lineup protocol provides the calibration of the human eyes to prevent us from erroneously jumping to conclusions about magic patterns.
-Earlier than the protocol proscription, we can find preliminary ideas along these lines, for example \citet{Daniel} recommends looking at plots of data from a null distribution for exploring two factor design experimental data.
+Prior to the protocol description, we can find preliminary ideas along these lines, for example \citet{Daniel} recommends looking at plots of data from a null distribution for exploring two factor design experimental data.
 His paper provides 40 pages of plots from the null!
 He encourages data analysts to understand the patterns that might be seen in samples data drawn from populations without inherent structure.
-(\citet{Buja} calls this the Rorschach protocol.)
+(\citet{Buja} has named this the Rorschach protocol.)
 
@@
 
-With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing?
+%With the general lack of inference present in exploratory data analysis, how do we attempt to control error in hypothesis testing?
 In order to use visualisation effectively, we will get the most benefit if the type of graphic display chosen is best suited to the task at hand, be it EDA, MD, or even the presentation of results.
-We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within.
+%We then need to find the statistical power of graphics, the ability of a graphical display to convey information on the structure of the data within.
 This can be found by using the lineup protocol developed in 2009 by \citet{Buja}, which is easily implemented with the nullabor R package.
-They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
-The lineup protocol is inspired in part by the police lineup \citep{Wickham}.
+%They created a statistically rigorous framework with properties explored primarily in \citet{Majumder:2014} and further in \citet{Hofmann:2014} that allows us to conduct these hypothesis tests with visual inference, thus to some degree allowing the conjoining of EDA with classical inferential statistics.
+%The lineup protocol is inspired in part by the police lineup \citep{Wickham}.
 We place the plot of the 'true' data generated with some underlying structure (The criminal) in amongst plots of data that were generated from a null distribution (The innocent people).
 If the real data plot is selected by an uninvolved observer as being most different from the other plots there is evidence that the structure of that data has led to a significant difference in the plot (That the criminal is sufficiently different from the innocents).
-This constitutes a rejection of the null hypothesis.
+%This constitutes a rejection of the null hypothesis.
 If one of the null plots is chosen instead then either the plot design did not have the sufficient power to display the true relationship (A Type 2 error in EDA), or the null plot exhibited a strong relationship that was generated randomly and an analyst that saw this plot and decided that the data had some structure would’ve committed a Type 1 error.
-$p_{i}$, the probability that plot $i$ is chosen in the lineup depends not only on the plot design $d$, but also the signal strength, $q_{i}$, of that plot and of every other competing plot in the lineup.
-It can be defined as some unknown function $f_{i,d} (q_{1},\dots,q_{20})$ \citep{Hofmann:2014}.
-If the true plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the true plot has a greater signal strength than those generated under the null distribution.
-The plot design that has a human observer selecting the true plot the most often will thus minimise both Type 1 error and Type 2 error in EDA hypothesis testing.
-However, there is the possibility that the true plot was selected by chance, that we have committed a Type 1 error in the lineup test.
-For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type 1 error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
-To control our error rate, it is initially obvious that we can simply change $m$, with an increase in null lineups having an inverse reduction in $\alpha$.
-We recruit human observers to judge the lineups, but this can quickly lead to a large cognitive burden to sort through more and more null plots.
-A much more powerful option is to have multiple different viewers of each lineup, with $K$ observers; the probability that at a particular plot was selected at least $x$ times under the null hypothesis is binomially distributed \citep{Buja, Hofmann:2014} with:
+
+% First the p-value calculation for a lineup, multiple observers
+
+The lineup protocol specifies that the data plot be placed in a random location in a field of $(m-1)$ null plots.
+Let $p_{i}$ be the probability that plot $i$ is chosen from the lineup.
+It depends on the plot design, $d$, and the signal strength, $q_{i}$, of the structure displayed in plot relative to the other plots in the lineup.
+This can be defined as a function $f_{i,d} (q_{1},\dots,q_{20})$\citep{Hofmann:2014}.
+If the data plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the data plot has a higher signal strength than the null plots.
+The plot design that has a human observer selecting the data plot the most often will thus minimise both Type I error and Type II error in the hypothesis testing.
+
+However, there is the possibility that the true plot was selected by chance, that we have committed a Type I error in the lineup test.
+For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type I error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
+To control our error rate, we could increase change $m$, this could produce a large cognitive burden on one observer, having to sort through many plots.
+An alternative approach is to recruit more observers, $K$, to evaluate a lineups, then the probability that plot $i$ is selected by $x$ observers under the null hypothesis follows a binomial distribution \citep{Majumder:2013}:
 
@@
 
+% Multiple picks
+
 If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability $p_{j} = s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
@@
 Where $T_{x}$ is all possible sets of $x$ trials out of $K$ resulting in a success, $A \in T_{x}$ is a specific set of $x$ trials resulting in a success and $A^{c}$ is the complement to $A$, the set of $K-x$ trials resulting in a failure.
+
+% Then power calculation for comparing designs
 

### 2016-01-24 paper/paper.Rnw
third paragraph of section 2

@@
 To test the null hypothesis the data plot is placed in a lineup with plots of null data, data generated from a scenario of no pattern.
+For the scatterplot this can be achieved by permuting the values on one of the variables to break any association between the two variables.
 An impartial viewer is asked to select the plot that is the most different from the others.
-If they select the data plot as the most different, this counts as a rejection of the null hypothesis, and follow-up questions may determine the type of departure: clusters, outlier, nonlinearity in the case of a scatterplot.
+If they select the data plot as the most different, this counts as a rejection of the null hypothesis, and follow-up questions on why they made their choice may determine the type of departure from the null.
 With multiple independent observers a $p$-value can be calculated to measure the statistical significance of the departure from the null.
@@
 He encourages data analysts to understand the patterns that might be seen in samples data drawn from populations without inherent structure.
-(\citet{Buja} has named this the Rorschach protocol.)
+(\citep{Buja} has named this the Rorschach protocol.)
 
@@
 
-The lineup protocol specifies that the data plot be placed in a random location in a field of $(m-1)$ null plots.
+The lineup protocol specifies that the data plot be placed in a randomly selected location in a field of $(m-1)$ null plots.
+
 Let $p_{i}$ be the probability that plot $i$ is chosen from the lineup.
-It depends on the plot design, $d$, and the signal strength, $q_{i}$, of the structure displayed in plot relative to the other plots in the lineup.
+The choice will depend on the plot design, $d$, and the signal strength, $q_{i}$, of the structure displayed in plot relative to the other plots in the lineup.
 This can be defined as a function $f_{i,d} (q_{1},\dots,q_{20})$\citep{Hofmann:2014}.
 If the data plot is detected, it indicates that the plot design could convey the desired information about the underlying structure and that the data plot has a higher signal strength than the null plots.
-The plot design that has a human observer selecting the data plot the most often will thus minimise both Type I error and Type II error in the hypothesis testing.
-
-However, there is the possibility that the true plot was selected by chance, that we have committed a Type I error in the lineup test.
-For a lineup of $m$ plots, the probability of selecting any plot when they are all generated by the null distribution is $1/m$, setting the Type I error rate, $\alpha$, of a lineup hypothesis test to equal $1/m$.
+
+If the data plot has no pattern, that is, the null hypothesis is true, there is a chance of making a Type I error.
+For a lineup of $m$ plots, the probability of selecting any plot is $1/m$, setting the Type I error rate, $\alpha$, of the hypothesis test to equal $1/m$.
 To control our error rate, we could increase change $m$, this could produce a large cognitive burden on one observer, having to sort through many plots.
@@
 \end{equation}
+
+The plot design that has a human observer selecting the data plot the most often will thus minimise both Type I error and Type II error in the hypothesis testing.
 

### 2016-01-25 paper/paper.Rnw
Subfigures

@@
 \usepackage[font=small,skip=5pt]{caption}
-\usepackage{subcaption}
+\usepackage[aboveskip=2pt]{subcaption}
 \usepackage{graphicx}
@@
   \centering
-<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+  \begin{subfigure}[t]{0.24\textwidth}
+<<smoothing, dependson='setup', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(123)
@@
 uns_l <- gather(uns, variable, value, -t)
-p1a <- ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() +labs(title="Unsmoothed Overlaid Lines") +
-    theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p3a <- ggplot(data=uns, aes(x=X1, y=X2)) + geom_point()+ labs(title="Unsmoothed Scatterplot)") +
-    theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-
 set.seed(123)
@@
 sm_l <- gather(sm, variable, value, -t)
-p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
-    theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-
-grid.arrange(p1a, p2a, p3a, p4a, ncol=2)
-@
-  \caption{The same AR(1) data both unsmoothed and smoothed for both plot designs. t = 96, correlation = 0.86}
+ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{smootha}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<smoothing2, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=uns, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{smoothb}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<smoothing3, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{smoothc}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<smoothing4, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{smoothd}
+\end{subfigure}
+  \caption{The same AR(1) data both unsmoothed (\subref{smootha} and \subref{smoothb}) and smoothed (\subref{smoothc} and \subref{smoothd}) for both plot designs. t = 96, correlation = 0.86}
   \label{smoothing}
@@
   \centering
-<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+  \begin{subfigure}[t]{0.24\textwidth}
+<<designs, dependson='setup', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(101)
@@
 b_l <- gather(b, variable, value, -t)
-p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
-@
-  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
+ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{designa}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<designs2, dependson='designs', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{designb}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<designs3, dependson='designs', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{designc}
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+<<designs4, dependson='designs', fig.width=4, fig.height=4, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\subcaption{}\label{designd}
+\end{subfigure}
+\caption{The same two data sets in both plot designs. t=24, correlation (\subref{designa} and \subref{designc}) = 0.946, correlation (\subref{designb} and \subref{designd})  = -0.943}
 \label{designs}
@@
 ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
+        geom_hline(aes(yintercept=0, colour="red")) + labs(x="Correlation", y="Difference between plot designs") + facet_grid(n~smoothed)
 @
@@
 <<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Detected", x="Correlation") + theme(legend.position="bottom")
+ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="True plot detected rate", x="Correlation") + theme(legend.position="bottom")
 @
@@
   \centering
-<<pins, dependson='setup', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+  \begin{subfigure}[htbp]{1\textwidth}
+<<pins, dependson='setup', out.width='1\\linewidth',  out.height='0.2\\textheight', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 count.setup <- select(full, pic_id, response_0, actual_location, design, smoothed, n)
@@
               correlation = diag(as.matrix(correl[pin$lineup, (pin$plot+2)])))
-pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
+ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
          geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
          geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + 
-         labs(title ="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
+         labs(y="Percent Chosen") + theme(legend.position="none")
+@
+\subcaption{}\label{pin1}
+\end{subfigure}
+\begin{subfigure}[htbp]{1\textwidth}
+<<pin2, dependson='pins', out.width='1\\linewidth', out.height='0.2\\textheight', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
          geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
          geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
+         labs(y="Percent Chosen") + theme(legend.position="none")
+@
+\subcaption{}\label{pin2}
+\end{subfigure}
+\begin{subfigure}[htbp]{1\textwidth}
+<<pin3, dependson='pin', out.width='1\\linewidth', out.height='0.2\\textheight', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
          geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
          geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) +
-         labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + 
+         labs(y="Percent Chosen") + theme(legend.position="none")
+@
+\subcaption{}\label{pin3}
+\end{subfigure}
+\begin{subfigure}[htbp]{1\textwidth}
+<<pin4, dependson='smoothing', out.width='1\\linewidth', out.height='0.2\\textheight', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + 
          geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
          geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
-@
-  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
+         labs(y="Percent Chosen") + theme(legend.position="none")
+@
+\subcaption{}\label{pin4}
+\end{subfigure}
+  \caption{Pin plots of the subject selections: (\subref{pin1}) unsmoothed scatterplots, (\subref{pin2}) smoothed scatterplots, (\subref{pin3}) unsmoothed overlaid lines, (\subref{pin4}) smoothed overlaid lines.
 On the x-axis is the correlation of the plot (there typically is spurious correlation in the null plots), and on the y-axis is the proportion of times a plot was chosen.

### 2016-01-25 paper/supplementary.Rnw
Conclusion
Subfigures

@@
 
-\title{\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation}
+\title{\bf Supplementary Material: What is the Best Way to Display Two Time Series to Read Association?}
 
@@
   \begin{center}
-    {\LARGE\bf Supplementary Material: Comparing the Power of Plot Designs to Reveal Correlation}
+    {\LARGE\bf Supplementary Material: What is the Best Way to Display Two Time Series to Read Association?}
 \end{center}
@@
 
-The scores of each subject given by equation~\ref{score} can be summed to give a total score per subject $j$:
+The scores of each subject given by equation 5 of the main paper can be summed to give a total score per subject $j$:
 

### 2016-01-25 paper/paper.Rnw
fixing lots of small things, captions, figure styles

--- 2016-01-24 paper/paper.Rnw
@@
   \centering
-<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+<<lineup.example, dependson='setup', fig.width=6, fig.height=6, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(1234)
@@
 @
-  \caption{A lineup ($m = 20$) of line graphs. Which plot shows the series with the strongest correlation?}
+  \caption{Lineup of overlaid line graphs of size $m = 20$. In which plot are the two series most strongly associated?}
   \label{lineup:example}
@@
   \centering
-<<smoothing, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+  \begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{smootha}
+<<smoothing, dependson='setup', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(123)
@@
 sm_l <- gather(sm, variable, value, -t)
-p2a <- ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Smoothed Overlaid Lines") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4a <- ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + labs(title="Smoothed Scatterplot") +
-    theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-
-grid.arrange(p1a, p2a, p3a, p4a, ncol=2)
-@
-  \caption{The same AR(1) data both unsmoothed and smoothed for both plot designs. t = 96, correlation = 0.86}
+ggplot(data=uns_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{smoothb}
+<<smoothing2, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=uns, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{smoothc}
+<<smoothing3, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=sm_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{smoothd}
+<<smoothing4, dependson='smoothing', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=sm, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+  \caption{Plots illustrating the experimental design.
+The same AR(1) data both unsmoothed (\subref{smootha} and \subref{smoothb}) and smoothed (\subref{smoothc} and \subref{smoothd}) for both plot designs, with t = 96 and correlation = 0.86}
   \label{smoothing}
@@
   \centering
-<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+  \begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{designa}
+<<designs, dependson='setup', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 set.seed(101)
@@
 b_l <- gather(b, variable, value, -t)
-p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
-      theme(axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
-      theme(aspect.ratio=1, axis.title.x = element_blank(),
-          axis.title.y = element_blank(),
-          axis.text.x = element_blank(),
-          axis.text.y = element_blank(),
-          axis.ticks = element_blank(),
-          legend.position="none")
-grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
-@
-  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
+ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{designb}
+<<designs2, dependson='designs', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{designc}
+<<designs3, dependson='designs', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\begin{subfigure}[t]{0.24\textwidth}
+\subcaption{}\label{designd}
+<<designs4, dependson='designs', fig.width=4, fig.height=4, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + theme(aspect.ratio=1, axis.title.x = element_blank(), axis.title.y = element_blank(),
+          axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks = element_blank(), legend.position="none")
+@
+\end{subfigure}
+\caption{Plots illustrating the experimental design: the two different plot types of the same data with $t=24$, $r= 0.946$ (\subref{designa} and \subref{designc}), $r= -0.943$ (\subref{designb} and \subref{designd}).}
 \label{designs}
@@
 ggplot(data=bylineup, aes(x=r, y=percent, colour=design)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
-  scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
+  scale_colour_discrete(name = "Plot Design") + labs(y="Detection rate", x="Correlation") +
   theme(legend.position="bottom")
@@
 ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
-        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
+        geom_hline(aes(yintercept=0, colour="red")) + labs(x="Correlation", y="Power difference") + facet_grid(n~smoothed)
 @
@@
 <<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
-ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Detected", x="Correlation") + theme(legend.position="bottom")
+ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Detection rate", x="Correlation") + theme(legend.position="bottom")
 @
@@
   \centering
-<<confidence2, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+<<confidence2, dependson='setup', fig.width=6, fig.height=6, out.width='0.9\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 conf2 <- select(full, conf_level, r, smoothed, n, design) #c(7, 15
@@
   \centering
-<<pins, dependson='setup', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+%  \begin{subfigure}[htbp]{1\textwidth}
+%\subcaption{}\label{pin1}
+<<pins, dependson='setup', fig.height=4, fig.width=7, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
 count.setup <- select(full, pic_id, response_0, actual_location, design, smoothed, n)
@@
               correlation = diag(as.matrix(correl[pin$lineup, (pin$plot+2)])))
-pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
+pin$design <- factor(pin$design, levels=c("Scatterplot", "Overlaid Lines"), labels=c("scat","lines"))
+pin$smoothed <- factor(pin$smoothed, levels=c("0", "1"), labels=c("N","Y"))
+#ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + 
+#         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+#         geom_point(aes(x=correlation, y=percent, colour=true))  + facet_grid(n~r) +
+#         scale_x_continuous(breaks=c(-1,0,1)) + scale_y_continuous(breaks=c(0,0.5,1)) +
+#         labs(y="Proportion",x="") + theme(legend.position="none")
+@
+%\end{subfigure}
+%\begin{subfigure}[htbp]{1\textwidth}
+%\subcaption{Scatterplots}\label{pin2}
+<<pin2, dependson='pins', fig.height=8, fig.width=7, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+ggplot(data=subset(pin)) + 
          geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + 
-         labs(title ="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
-pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) +
-         labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + 
-         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
-         geom_point(aes(x=correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + 
-         labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
-grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
-@
-  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size.
+         geom_point(aes(x=correlation, y=percent, colour=true)) + facet_grid(design+n+smoothed~r) + 
+         scale_x_continuous(breaks=c(-1,-0.5,0,0.5,1), labels=c("1","-","0","+","1")) + 
+         scale_y_continuous(breaks=c(0,0.5,1)) +
+         labs(y="Proportion",x="Correlation between series") + theme(legend.position="none")
+#ggplot(data=subset(pin, design == "Scatterplot")) + 
+#         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+#         geom_point(aes(x=correlation, y=percent, colour=true)) + facet_grid(n+smoothed~r) + 
+#         scale_x_continuous(breaks=c(-1,0,1)) + scale_y_continuous(breaks=c(0,0.5,1)) +
+#         labs(y="Proportion",x="Correlation") + theme(legend.position="none")
+#ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + 
+#         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+#         geom_point(aes(x=correlation, y=percent, colour=true)) + facet_grid(n~r) + 
+#         scale_x_continuous(breaks=c(-1,0,1)) + scale_y_continuous(breaks=c(0,0.5,1)) +
+#         labs(y="Proportion",x="") + theme(legend.position="none")
+@
+%\end{subfigure}
+%\begin{subfigure}[htbp]{1\textwidth}
+%\subcaption{}\label{pin3}
+<<pin3, dependson='pin', fig.height=5, fig.width=7, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+#ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) +
+#         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+#         geom_point(aes(x=correlation, y=percent, colour=true)) + facet_grid(n~r) +         
+#         scale_x_continuous(breaks=c(-1,0,1)) + scale_y_continuous(breaks=c(0,0.5,1)) +
+#         labs(y="Proportion",x="") + theme(legend.position="none")
+@
+%\end{subfigure}
+%\begin{subfigure}[htbp]{1\textwidth}
+%\subcaption{Overlaid line graphs}\label{pin4}
+<<pin4, dependson='smoothing', fig.height=5, fig.width=7, out.width='0.9\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
+#ggplot(data=subset(pin, design == "Overlaid Lines")) + 
+#         geom_errorbar(aes(x=correlation, ymin=base, ymax=percent, colour=true)) +
+#         geom_point(aes(x=correlation, y=percent, colour=true)) + facet_grid(n+smoothed~r) + 
+#         scale_x_continuous(breaks=c(-1,0,1)) + scale_y_continuous(breaks=c(0,0.5,1))+
+#         labs(y="Proportion",x="Correlation") + theme(legend.position="none")
+@
+%\end{subfigure}
+  \caption{Pin plots showing subject selections for each plot in all lineups for all factor levels in the experiment.
+Proportion of times plot selected plotted by correlation between series in each plot.
+Blue indicates data plot.
+Longer pins indicate more people selected that plot in the lineup.
+Results for the three replicate lineups for each treatment are  plotted together.
+Generally, subjects selected the plot from the lineup with the strongest association.
+With small sample size this is less obvious.
+For overlaid lines, subjects tended to more often pick the plot showing the strongest positive association, particularly if the correlation between series in the data plot was low.
+}
+    %: (\subref{pin1}) unsmoothed scatterplots, (\subref{pin2}) smoothed scatterplots, (\subref{pin3}) unsmoothed overlaid lines, (\subref{pin4}) smoothed overlaid lines.
 On the x-axis is the correlation of the plot (there typically is spurious correlation in the null plots), and on the y-axis is the proportion of times a plot was chosen.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{anom1}
+  \includegraphics[width=0.9\textwidth]{anom1}
   \caption{A lineup of unsmoothed scatter plots with a sample size of 24.
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{anom2}
+  \includegraphics[width=0.9\textwidth]{anom2}
   \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
@@
  \centering
-<<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
+<<powermodel, dependson='setup', fig.height=8, fig.width=8, out.width='0.9\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
 predplot <- prediction %>% filter(is.na(detected)) %>%

### 2016-01-27 paper/supplementary.Rnw
supplementary

@@
 
-\section{Subject Demographics}
-
-\begin{figure}[htbp]
-\centering
-<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=3>>=
-location <- read.csv("location.csv")
-map.dat <- map_data("world")
-ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
-  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
-@
-\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
-\label{map}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-<<users, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
-users <- read.csv("turk18_users.csv")
-users$gender <- ifelse(users$gender==1, "Male", "Female")
-ggplot(data=users, aes(x=factor(age), fill=factor(academic_study))) + geom_bar() + facet_wrap(~gender) +
-  scale_fill_manual(values=c("#550000", "#883388", "#1133FF", "#FF9933", "#555555"))
-@
-\caption{Age and education by gender. Most subjects are in the younger brackets and there are more males than females}
-\label{ageed}
-\end{figure}
-
-\section{Descriptive Statistics of Evaluations}
+\section{Lineup Results}
 
@@
 
+\section{Subject Demographics}
+
 \begin{figure}[htbp]
+\centering
+<<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=3>>=
+location <- read.csv("location.csv")
+map.dat <- map_data("world")
+ggplot() + geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) + theme_bw() +
+  geom_point(data=location, aes(x=lon, y=lat, colour= "blue")) +scale_colour_manual(values=c("#FF5500", "#00FFFF")) + theme(aspect.ratio = 0.67, axis.text=element_blank(), axis.title=element_blank(), legend.position="none")
+@
+\caption{The location of the recruited subjects. The majority reside in North America but some are spread around other countries in Central America, Europe and Asia.}
+\label{map}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+<<users, dependson='setup', echo=FALSE, results='asis', fig.height=3, message=FALSE, error=FALSE, warning=FALSE>>=
+users <- read.csv("turk18_users.csv")
+users$academic_study <- factor(users$academic_study)
+levels(users$academic_study) <- c("At Most High School", "Some Undergraduate", "Undergraduate Degree", "Some Postgraduate", "Postgraduate Degree")
+users$age <- factor(users$age)
+levels(users$age) <- c("18-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61+")
+ggplot(data=users, aes(x=age, fill=academic_study)) + geom_bar() + facet_wrap(~gender) +
+  scale_fill_manual(name = element_blank(), values=c("#550000", "#883388", "#1133FF", "#FF9933", "#555555")) + labs(x="Age", y=element_blank()) + theme(axis.text.x = element_text(angle = 270))
+@
+\caption{Age and education by gender. Most subjects are in the younger brackets and there are more males than females}
+\label{ageed}
+\end{figure}
+
+This section describes the demographics of the subjects in the experiment recruited through Amazon's Mechanical Turk.
+Age, gender, education and location all did not have a statistically significant effect on the selections made.
+
+\section{Descriptive Statistics of Evaluations}
+
+\begin{figure}[h]
 \centering
@@
 choice$correct <- factor(choice$correct, labels=c("Did not detect", "Detected"))
+choice$choice_reason <- factor(choice$choice_reason)
+levels(choice$choice_reason) <- c("Close Together", "Matching Peaks", "Pattern Formed", "Formed a line", "Other")
 ggplot(data=choice, aes(x=choice_reason, fill=correct)) + geom_bar() + facet_grid(design~rsgn) +
@@
 @
-\caption{Reasons given for plot selection}
+\caption{Reasons given for plot selection. After controlling for lineup factors, any explanatory power of the reason is removed.}
 \label{reasons}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering 
-<<avscore, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>= 
-subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
-subjplot <- mutate(subjplot, avscore= total/eval)
-ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.05)
-@
-\caption{Average score of subjects included in modelling}
-\label{avscore}
 \end{figure}
@@
 m.lineup <- merge(m.freq, bylineup, by.x="Var1", by.y="pic_id")
-ggplot(data=m.lineup) + geom_bar(aes(x=factor(r), fill=design), binwidth=0.2) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
+ggplot(data=m.lineup) + geom_bar(aes(x=factor(r), fill=design)) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
 @
-\caption{Characteristics of lineups with at least one subject making multiple selections}
+\caption{Characteristics of lineups with at least one subject making multiple selections.
+Overlaid lines lineups have multiple selections more often, and true plot correlation does not appear to have an impact except at large sample sizes.}
 \label{lineupmulti}
@@
 
-\begin{table}[htbp]
-\caption{Number of selections per lineup evaluation. The subject with the most selections is reported.}
+\begin{table}[h]
+\caption{Number of selections for each lineup evaluation. The subject with the most selections is additionally reported.}
 \label{table}
@@
 
+\iffalse
 \begin{figure}[htbp]
@@
 \label{attemptshist}
+\end{figure}
+\fi
+
+Subjects are allowed to make multiple selections on each lineup evaluation, with a score for each evaluation defined by:
+
+\begin{equation}
+\label{score}
+{score}_{ij} =\frac{m-s_{ij}}{m-1} \mathbbm{I}_{ij}
+\end{equation}
+
+Where $m = 20$, the number of plots in a linup, $s_{ij}$ is the number of selections made by subject $i$ on lineup $j$ and $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise.
+The most selections made on any lineup evaluation is 8.
+
+\begin{figure}[htbp]
+\centering 
+<<avscore, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>= 
+subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
+subjplot <- mutate(subjplot, avscore= total/eval)
+ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.1)
+@
+\caption{The average score of our subjects, according to formula \ref{score}. There is a wide range of visual ability to be controlled with random effects in the modelling}
+\label{avscore}
 \end{figure}
@@
 Analysing the level one residuals, $\hat{\epsilon_{i}}$, and level two residuals, $\hat{\gamma_{i}}$, allows us to test the model fit as described \citet{Loy}.
-There are some negative extreme outliers but the distributional assumption appears to hold.
+There are some negative extreme outliers but the distributional assumptions appear to hold.
 
@@
 
-The scores of each subject given by equation 5 of the main paper can be summed to give a total score per subject $j$:
+The scores of each subject given by equation~\ref{score} be summed to give a total score per subject $j$:
 

### 2016-01-27 paper/supplementary.Rnw
supplementary

@@
 @
-\caption{Age and education by gender. Most subjects are in the younger brackets and there are more males than females}
+\caption{Age and education by gender. Younger age brackets and males are over-represented in the sample, but these factors caused no significant impact on results.}
 \label{ageed}
@@
 
-This section describes the demographics of the subjects in the experiment recruited through Amazon's Mechanical Turk.
-Age, gender, education and location all did not have a statistically significant effect on the selections made.
+Despite there being a wide range of subject demographics in the experiment recruited through Amazon's Mechanical Turk, age, gender, education and location all did not have a statistically significant effect on power of either plot design.
 
-\section{Descriptive Statistics of Evaluations}
+Subjects are allowed to make multiple selections on each lineup evaluation, with a score for each evaluation defined by:
+
+\begin{equation}
+\label{score}
+{score}_{ij} =\frac{m-s_{ij}}{m-1} \mathbbm{I}_{ij}
+\end{equation}
+
+Where $m = 20$, the number of plots in a linup, $s_{ij}$ is the number of selections made by subject $i$ on lineup $j$ and $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise.
+The most selections made on any lineup evaluation is 8.
+
+\begin{figure}[htbp]
+\centering 
+<<avscore, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=4>>= 
+subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
+subjplot <- mutate(subjplot, avscore= total/eval)
+ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.1)
+@
+\caption{The average score of our subjects, according to formula \ref{score}. There is a wide range of visual ability to be controlled with random effects in the modelling}
+\label{avscore}
+\end{figure}
+
+\section{Evaluation Descriptive Statistics}
 
@@
 \centering
-<<choices, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<choices, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=4>>=
 full$choice_reason <-sapply(strsplit(as.character(unlist(full$choice_reason)), "[^0-9]+"), "[[", 1)
@@
 @
-\caption{Reasons given for plot selection. After controlling for lineup factors, any explanatory power of the reason is removed.}
+\caption{Reasons given for plot selection. After controlling for lineup factors, there is no explanatory power.}
 \label{reasons}
@@
 \centering
-<<multiselect, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<multiselect, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=4>>=
 m.select <- summarise(group_by(subset(full.sub, attempts>1), subj_id, pic_id), attempts=max(attempts))
@@
 \caption{Characteristics of lineups with at least one subject making multiple selections.
-Overlaid lines lineups have multiple selections more often, and true plot correlation does not appear to have an impact except at large sample sizes.}
+Overlaid line lineups have multiple selections more often, and true plot correlation does not appear to have an impact except at large sample sizes.}
 \label{lineupmulti}
@@
 \begin{table}[h]
-\caption{Number of selections for each lineup evaluation. The subject with the most selections is additionally reported.}
+\centering
+\caption{Number of times $n$ selections from a lineup evaluation were made.
+The subject who made the most selections is additionally reported.
+There is no relationship between a subject's demographics and a likelihood to make multiple selections.}
 \label{table}
@@
 \centering
-<<multiattempts, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>=
+<<multiattempts, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=4>>=
 ggplot(data=attempts) + geom_bar(aes(x=as.factor(attempts))) + labs(x="Number of selections", y=NULL)
@@
 \fi
-
-Subjects are allowed to make multiple selections on each lineup evaluation, with a score for each evaluation defined by:
-
-\begin{equation}
-\label{score}
-{score}_{ij} =\frac{m-s_{ij}}{m-1} \mathbbm{I}_{ij}
-\end{equation}
-
-Where $m = 20$, the number of plots in a linup, $s_{ij}$ is the number of selections made by subject $i$ on lineup $j$ and $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise.
-The most selections made on any lineup evaluation is 8.
-
-\begin{figure}[htbp]
-\centering 
-<<avscore, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE>>= 
-subjplot <- summarise(group_by(full.sub, subj_id), eval=sum(weights), total = sum(score))
-subjplot <- mutate(subjplot, avscore= total/eval)
-ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.1)
-@
-\caption{The average score of our subjects, according to formula \ref{score}. There is a wide range of visual ability to be controlled with random effects in the modelling}
-\label{avscore}
-\end{figure}
 

### 2016-01-31 paper/paper.Rnw
Currency rates

@@
   \centering
-    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
-    \includegraphics[width=0.45\textwidth]{nzdline}}
+    \centerline{\begin{subfigure}[width=0.45\textwidth]
+    <<nzd, error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+    rates <- select(read.csv("rates.csv"), date, AUD, CNY, NZD)[36:69,]
+    rates <- mutate(rates, AUD2=(AUD-mean(AUD))/sd(AUD), CNY2=(CNY-mean(CNY))/sd(CNY)) #scaling
+    ggplot(data=rates, aes(x=AUD, y=NZD)) + geom_point()
+    @
+    \end{subfigure}
+    }
+    \begin{subfigure}[width=0.45\textwidth]
+    <<nzd2, dependson='nzd', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+    rates_l <- gather(rates, currency, value, -date)
+    ggplot(data=subset(rates_l, currency != "CNY" & currency != "CNY2" & currency != "AUD2"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line()
+    @
+    \end{subfigure}
   \caption{Cross-rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{cnyline}
+\begin{figure}[width=0.5\textwidth][htbp]
+  \centering
+  <<cny, dependson='nzd2', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+   ggplot(data=subset(rates_l, currency != "NZD" & currency != "CNY" & currency != "AUD"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line() 
+  @
   \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.

### 2016-01-31 paper/paper.Rnw
Currency rates

--- 2016-01-25 paper/paper.Rnw
@@
   \centering
-    \centerline{\includegraphics[width=0.45\textwidth]{nzdscatter}
-    \includegraphics[width=0.45\textwidth]{nzdline}}
+    \begin{subfigure}[htbp]{0.45\textwidth}
+    <<nzd, error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+    library(dplyr)
+    library(tidyr)
+    library(ggplot2)
+    rates <- select(read.csv("rates.csv"), date, AUD, CNY, NZD)[36:69,]
+    rates <- mutate(rates, AUD2=(AUD-mean(AUD))/sd(AUD), CNY2=(CNY-mean(CNY))/sd(CNY)) #scaling
+    ggplot(data=rates, aes(x=AUD, y=NZD)) + geom_point()
+    @
+    \end{subfigure}
+    \begin{subfigure}[htbp]{0.45\textwidth}
+    <<nzd2, dependson='nzd', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+    rates_l <- gather(rates, currency, value, -date)
+    ggplot(data=subset(rates_l, currency != "CNY" & currency != "CNY2" & currency != "AUD2"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line()
+    @
+    \end{subfigure}
   \caption{Cross-rates for Australian and NZ dollars against US dollar, from Mar 30-May 2, 2015, shown using a scatterplot (left) and overlaid line graph (right).
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{cnyline}
+  <<cny, dependson='nzd2', out.width='0.5\\textwidth', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
+   ggplot(data=subset(rates_l, currency != "NZD" & currency != "CNY" & currency != "AUD"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line() 
+  @
   \caption{Cross-rates for Australian dollar and Chinese yuan against the United States dollar from Mar 30-May 2, 2015, shown as overlaid line graph.

### 2016-02-01 paper/paper.Rnw
Restructure

@@
 # Convert these to read_csv at some point
-correl <- read.csv("correlations.csv", stringsAsFactors = FALSE) ##Correlations of every plot in every lineup
-full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
-prediction <- read.csv("sas/predfit.csv") ##fitted data from SAS
-time.fit <- read.csv("sas/time.csv") ##coefficients for time model from SAS
-powercoef <- read.csv("sas/PowerEstimates.csv") ##coefficients for power model from SAS
+correl <- read.csv("data/correlations.csv", stringsAsFactors = FALSE) ##Correlations of every plot in every lineup
+full <- read.csv("data/full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
+prediction <- read.csv("data/predfit.csv") ##fitted data from SAS
+time.fit <- read.csv("data/time.csv") ##coefficients for time model from SAS
+powercoef <- read.csv("data/PowerEstimates.csv") ##coefficients for power model from SAS
 
@@
     library(ggplot2)
-    rates <- select(read.csv("rates.csv"), date, AUD, CNY, NZD)[36:69,]
+    rates <- select(read.csv("data/rates.csv"), date, AUD, CNY, NZD)[36:69,]
     rates <- mutate(rates, AUD2=(AUD-mean(AUD))/sd(AUD), CNY2=(CNY-mean(CNY))/sd(CNY)) #scaling

### 2016-02-01 paper/supplementary.Rnw
Currency rates
Restructure

@@
 library(tidyr)
-full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
-prediction <- read.csv("sas/predfit.csv") ##Large file, very slow, fitted data from SAS
-time.fit <- read.csv("sas/time.csv") ##coefficients for time model from SAS
-powercoef <- read.csv("sas/PowerEstimates.csv") ##coefficients for power model from SAS
+full <- read.csv("data/full.csv", row.names=1, stringsAsFactors = FALSE) ##MTurk evaluations
+prediction <- read.csv("data/predfit.csv") ##Large file, very slow, fitted data from SAS
+time.fit <- read.csv("data/time.csv") ##coefficients for time model from SAS
+powercoef <- read.csv("data/PowerEstimates.csv") ##coefficients for power model from SAS
 
@@
 <<map, dependson='setup', echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE, fig.height=3>>=
-location <- read.csv("location.csv")
+location <- read.csv("data/location.csv")
 map.dat <- map_data("world")
@@
 <<users, dependson='setup', echo=FALSE, results='asis', fig.height=3, message=FALSE, error=FALSE, warning=FALSE>>=
-users <- read.csv("turk18_users.csv")
+users <- read.csv("data/turk18_users.csv")
 users$academic_study <- factor(users$academic_study)

### 2016-02-01 code/gen.lineups.R

@@ -1,0 +1,360 @@
+library(ggplot2)
+library(mnormt)
+library(reshape2)
+library(gridExtra)
+library(mFilter)
+library(nullabor)
+library(dplyr)
+library(gridSVG)
+library(grid)
+
+js <- "http://www.hofroe.net/examples/lineup/fhaction.js"
+
+gen_true_data <- function(n, r, smoothed = FALSE, a) {
+  d <- data.frame(rmnorm(n, c(0, 0), matrix(c(1, r, r, 1), 2, 2)))
+  d$X1 <- as.vector(arima.sim(list(ar=a), n, innov=d$X1))
+  d$X2 <- as.vector(arima.sim(list(ar=a), n, innov=d$X2))
+  if (smoothed) {
+    d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
+    d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
+    d <- as.data.frame(d)
+  }
+  d$X1 <- scale(d$X1)
+  d$X2 <- scale(d$X2)
+  d$t <- 1:n
+  return(d)
+}
+
+gen_null <- function(n, m=20, smoothed = FALSE, a){
+  nd <- NULL
+  for(i in 1:(m-1)) { 
+    d <- data.frame(X1=as.vector(arima.sim(list(ar=a), n)),
+                    X2=as.vector(arima.sim(list(ar=a), n)))
+    if (smoothed) {
+      d$X1 <- as.vector(hpfilter(d$X1, freq=1, type="lambda", drift = FALSE)[[2]])
+      d$X2 <- as.vector(hpfilter(d$X2, freq=1, type="lambda", drift = FALSE)[[2]])
+      d <- as.data.frame(d)
+    }
+    d$X1 <- scale(d$X1)
+    d$X2 <- scale(d$X2)
+    nd <- rbind(nd, d)
+  } 
+  nd$.n <- rep(1:(m-1), each = n) 
+  nd$t <- 1:n
+  return(nd)
+}
+
+nc <- function(null) {
+  b <- c(rep(0, 19))
+  for (i in 1:19) {
+    b[i] <- cor(subset(null, .n==i)[,1], subset(null, .n==i)[,2])
+  }
+  return(b)
+}
+
[…]

### 2016-02-01 paper/Extra Material.R

@@ -1,16 +1,0 @@
-#Residuals#
-
-extreme <- subset(glmm.resid, residual <= -10 | residual >= 10)
-extreme <- merge(extreme, full.sub[c(9,11)], by=0)
-
-##sources of extreme residuals:
-#multiple selections - one was correct, others are predicted to be correct but aren't
-#only person to get an easy one wrong
-#only person to get a hard one right
-#investigate pic_id = 61 name = s459.... => true plot in 18 looks alright, but most people picked plot 17 with cor = 0.24. 17 looks very good
-#if you ignore the outliers. Two chose reason 4, one chose reason 1. Last person selected 7 & 20, neither plot is especially interesting.
-
-
-
-
-

### 2016-02-01 paper/paper.Rnw
citing

@@
 He encourages data analysts to understand the patterns that might be seen in samples data drawn from populations without inherent structure.
-(\citep{Buja} has named this the Rorschach protocol.)
+\citet{Buja} has named this the Rorschach protocol.)
 
@@
 
+\section{Acknowledgements}
+
+The authors gratefully acknowledge funding from the National Science Foundation Grant DMS \#1007697.
+
+All code required to reproduce this paper can be found at https://github.com/NTomasetti/correlation.
+
 \bibliographystyle{asa}

### 2016-02-01 paper/paper.Rnw
Citations

@@
 
+The models used for this paper were generated using SAS software, Version 9 of the SAS System for Unix with the MIXED and GLIMMIX procedures.
+Copyright \©2015 SAS Institute Inc.
+SAS and all other SAS Institute Inc.
+product or service names are registered trademarks or trademarks of SAS Institute Inc., Cary, NC, USA.
+
+Further analysis was carried out using \citet{R} version 3.1.2 and \cited{RStudio} version 0.99.484 using the packages: \citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
+
 All code required to reproduce this paper can be found at https://github.com/NTomasetti/correlation.

### 2016-02-01 code/gen.lineups.R

@@ -1,6 +1,6 @@
 library(ggplot2)
 library(mnormt)
-library(reshape2)
+library(tidyr)
 library(gridExtra)
 library(mFilter)
 library(nullabor)
@@ -86,7 +86,7 @@ lineupscatter <- function(name, td, nd, pos, save=TRUE, w=8, h=8) {
 }
 
 lineupline <- function(name, td, nd, pos, save=TRUE, w=8, h=8) {
-  nd_l <- as.data.frame(melt(nd, id=c(".n", "t")))
+  nd_l <- as.data.frame(gather(nd, variable, value, -.n, -t))
   td <- data.frame(.n=rep(pos, nrow(td)), td)
   # Make data structure ourself
   if (pos == 1) {
@@ -130,7 +130,7 @@ Lineups <- function(n, m=20, c = 3, smoothed = FALSE, a=n/100, cor.matrix=FALSE)
   id <- 1
   for(z in 1:1000){
     real <- gen_true_data(n, r = (-0.9 + 1.8*z/1000), smoothed, a)
-    real_long <- melt(real, id="t")
+    real_long <- gather(real, value, variable, -t)
     cor <- cor(real$X1, real$X2)
     pos <- sample(1:20, 1)
     if (cor > 0.885 & cor < 0.915 & count[1] < c) { 
@@ -328,7 +328,7 @@ Lineups.trial <- function(n, dep=TRUE, m=20, c = 10, smoothed = FALSE) {
   id <- 1
   for(z in 1:1000){
     real <- gen_true_data(n, r = (-0.9 + 1.8*z/1000), dep, smoothed)
-    real_long <- melt(real, id="t")
+    real_long <- gather(real, variable, value, -t)
     cor <- cor(real$X1, real$X2)
     pos1 <- sample(1:20, 1)
     pos2 <- sample(1:20, 1)

### 2016-02-01 paper/paper.Rnw
Citations

--- 2016-02-01 paper/paper.Rnw
@@
 
+The models used for this paper were generated using SAS software, Version 9 of the SAS System for Unix with the MIXED and GLIMMIX procedures.
+Copyright \copyright\ 2015 SAS Institute Inc.
+SAS and all other SAS Institute Inc.
+product or service names are registered trademarks or trademarks of SAS Institute Inc., Cary, NC, USA.
+
+Further analysis was carried out using \citet{R} version 3.1.2 and \citet{RStudio} version 0.99.484 using the packages: \citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
+
 All code required to reproduce this paper can be found at https://github.com/NTomasetti/correlation.

### 2016-02-01 paper/paper.Rnw
Citations

@@
 
-Further analysis was carried out using \citet{R} version 3.1.2 and \citet{RStudio} version 0.99.484 using the packages: \citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
+Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: \citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
 

### 2016-02-01 paper/paper.Rnw
Citations

@@
 
-Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: \citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
+Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: dplyr, ggplot2, grid, gridExtra, gridSVG, knitr, maps, mFilter, mnormt, nullabor, string and tidyr.
+%\citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
 

### 2016-02-01 paper/paper.Rnw
Citations

@@
 Reading correlation between two sets of non-temporal context numbers has been examined in many papers.
-In \citet{harrison2014ranking} different graphical displays of two sets of numbers are ranked of their effectiveness by Weber's Law \citep{webers} which provides a relationship between perception and the ``effect size'' in the data.
+In \citet{Harrison} different graphical displays of two sets of numbers are ranked of their effectiveness by Weber's Law \citep{webers} which provides a relationship between perception and the ``effect size'' in the data.
 For example, two series with correlation 0.9 should be proportionately easier to distinguish from noise than two series with correlation 0.3.

### 2016-02-03 paper/paper.Rnw
Legend

@@
     rates_l <- gather(rates, currency, value, -date)
-    ggplot(data=subset(rates_l, currency != "CNY" & currency != "CNY2" & currency != "AUD2"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line()
+    ggplot(data=subset(rates_l, currency != "CNY" & currency != "CNY2" & currency != "AUD2"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line() + theme(legend.position="bottom")
     @
@@
   <<cny, dependson='nzd2', out.width='0.5\\textwidth', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
-   ggplot(data=subset(rates_l, currency != "NZD" & currency != "CNY" & currency != "AUD"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line() 
+   ggplot(data=subset(rates_l, currency != "NZD" & currency != "CNY" & currency != "AUD"), aes(x=as.Date(date), y=value, colour=currency)) + geom_line() + theme(legend.position="bottom")
   @

### 2016-02-04 paper/paper.Rnw
equation
grammar

@@
 There is substantial literature supporting this \citep{Cleveland, Harrison, Li}.
-However, it is still common that when the two variables are temporal they are displayed as overlaid time series (line graphs).
+However, it is still common that when the two variables are temporal they are displayed as an overlaid time series (line graphs).
 In this paper we have utilized new methods for rigorously comparing plot design, using visual inference, to investigate which is the best display if reading association is important.
@@
 \label{modelpower}
-Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
+Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{1,i} \* x_{2,j} + \beta_{4} x_{3,k} + \nonumber \\
+ & & \beta_{5} x_{4,t} + x_{5,l}(\beta_{6} + \beta_{7} x_{1,i}  + \beta_{8} x_{2,j}  + \nonumber \\
+ & & \beta_{9} x_{1,i} \* x_{2,j} + \beta_{10} x_{3,k} + \beta_{11} x_{4,t})  + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
 \end{eqnarray}
@@
 \label{timemodel}
-ln(time_{ijklmt}) &=& \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
- & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}\*(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber\\
- & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \nonumber 
+ln(time_{ijklt}) &=& \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{1,i} \* x_{2,j} + \beta_{4} x_{3,k} + \nonumber \\
+ & & \beta_{5} x_{4,t} + x_{5,l}(\beta_{6} + \beta_{7} x_{1,i}  + \beta_{8} x_{2,j}  + \nonumber \\
+ & & \beta_{9} x_{1,i} \* x_{2,j} + \beta_{10} x_{3,k} + \beta_{11} x_{4,t})  + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber 
 \end{eqnarray}
 
-Other than the removal of the link functon, the structure of the model identical to that in Equation~\ref{modelpower}.
+Other than the removal of the link function, the structure of the model identical to that in Equation~\ref{modelpower}.
 

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 If the plot design has the power to display correlation, we would expect the highest correlation plot to be selected, leading to a Type I error where a plot from the null distribution spuriously had high correlation.
-We find that this is generally not the case.
-The selections appear to have a wide range of correlations, indicating that if a subject selected a null plot it was typically not due to excess spurious correlation.
-Instead the failure to select the highest correlated plot indicates that a Type II error was committed rather than Type I as the plot designs did not have the power to differentiate between medium and low correlation plots.
-However, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected.
+We find that this is generally the case at higher sample sizes, but at $t = 12$, the selected plos have a wider range of correlations, indicating a general inability of the plot designs to differentiate betwen high and low correlaiton, a type II error.
+With the catter plot, for the scatter plot, whenever there was a plot with a very strong correlation, usually in the form of the true plot, it was overwhelmingly selected.
 For the overlaid lines this mostly occurred only for plots with strong positive correlation, which is consistent with its general lack of power to reveal negative relationships in data.
@@
 %\end{subfigure}
-  \caption{Pin plots showing subject selections for each plot in all lineups for all factor levels in the experiment.
-Proportion of times plot selected plotted by correlation between series in each plot.
-Blue indicates data plot.
-Longer pins indicate more people selected that plot in the lineup.
-Results for the three replicate lineups for each treatment are  plotted together.
-Generally, subjects selected the plot from the lineup with the strongest association.
-With small sample size this is less obvious.
-For overlaid lines, subjects tended to more often pick the plot showing the strongest positive association, particularly if the correlation between series in the data plot was low.
-}
+  \caption{Pin plots showing subject selections for each plot in all lineups for all factor levels in the experiment, with duplicates of the same factor combination together.
+Proportion of times a plot was selected is on the y-axis and correlation between the pair of time series on the x-axis.
+Blue indicates a true data plot and red indicates a null plot.
+Generally, subjects selected the plot from the lineup with the strongest association in larger sample sizes.
+For overlaid lines, subjects tended to more often pick the plot showing the strongest positive association, even if the relationship in the true plot was negative.}
     %: (\subref{pin1}) unsmoothed scatterplots, (\subref{pin2}) smoothed scatterplots, (\subref{pin3}) unsmoothed overlaid lines, (\subref{pin4}) smoothed overlaid lines.
@@
   \includegraphics[width=0.9\textwidth]{anom2}
-  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707.
-The true plot in position 8 was not detected in five evaluations.
-However, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
-When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.}
+  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true plot correlation of -0.707 in position 8, which was not detected in five evaluations.
+Instead, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225.
+When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data as the subjects ignore the eventual separation in the lines, similar to in figure ~\ref{cny}.}
   \label{anom2}
@@
 Plot 13 has a correlation of 0.225, smaller (in absolute value) than the -0.707 of plot 8.
-This selection is in-line with Wertheimer’s gestalt law of common fate.
-It appears to be strongly correlated in the beginning of the series but breaks apart towards the end.
-The selection implies that subjects selecting this plot focused on the joint path of the two series early in the plot and ignored the contradictory evidence of no relationship as they diverged later in the plot.
-It also confounded by the seeming preference to find positive correlation over negative correlation, and we suspect that if one of the series in plot 8 had the values flipped, then the strong positive correlation would have been noticed.
+This selection gives evidence towards Wertheimer’s gestalt law of common fate, as subjects may be selecting this plot focused on the joint path of the two series early in Plot 13 and ignored the contradictory evidence of no relationship as they diverged later in the plot similar to the series in figure ~\ref{cny}.
+It also confounded by the seeming preference to find positive correlation over negative correlation, and we suspect that if one of the series in plot 8 had the values flipped, then the strong positive correlation may have been noticed.
 
@@
 \item $x_{4,t}$ is the sample size, with $t \in \{12, 24, 48, 96\}$ ,
-\item $x_{5,l}$ denotes the plot design, with $l =0$ for the overlaid lines type,
+\item $x_{5,l}$ si a dummy variable for the plot design, with $l = 0$ for the overlaid lines type,
 \item $z_{s}$ is an observer specific random effect, with $s = 1, 2, \dots , 160$.
@@
 Table~\ref{tab2} shows a similar story for time as Table~\ref{tab1} did for power, that the overlaid lines have better performance close to zero correlation, but the scatter plot is much faster at high correlations.
-Postivie correlation reduces time for both plot designs, but only the scatter plot sees improvements in negative correlation.
-Detecing the true plot was also faster than a null.
+Postivie correlation reduces time for both plot designs, but only the scatter plot sees improvements in negative correlation, time to detect is flat accross negative correlations for the overlaid line graph.
 Neither smoothness nor sample size has an impact on the time for either plot design, despite having statistically significant impacts on power.
@@
 -In low sample sizes, both designs have similarly weak performance - 
--Relationships between t=12 is dubious at best, requires $\rho$ > 0.7 to have more than 50\% power
--In higher sample sizes, the scatterplot improves dramatically while the line plot remains unchanged
--The line plot has much weaker performance in negative correlation than positive
--If possible, smooth the lines to improve performance
+-Relationships between t=12 time series is dubious at best, requires $\rho$ > 0.7 to have more than 50\% power to see.
+-In higher sample sizes, the scatterplot improves dramatically while the line plot remains mostly unchanged
+-Positive correlation is very similar between the two designs
+-The line plot has much weaker performance in negative correlation
+-If possible, smooth the lines to improve performance - noisy data hurts visual performance
 -If negative association is expected, flip one line to make it positive
@@
 Time
--The scatterplot is faster than the lines at high correlation
+-The scatterplot is faster than the lines, particularly at high correlation
 -Smoothing and sample size have no real effect
@@
 -People could be distracted by spurious patterns in the null plots
--These were very confident selections
+-These were typically very confident selections
 -The data analyst needs to understand what kinds of patterns can be generated randomly
@@
 Overall
--Despite hiding trends etc., the simplistic scatter plot seems to be best suited to these tasks.
+-Despite hiding trends etc., the seemingly simplistic scatter plot seems to be better suited to these tasks.
 
@@
 
-Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: dplyr, ggplot2, grid, gridExtra, gridSVG, knitr, maps, mFilter, mnormt, nullabor, string and tidyr.
+Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: dplyr, ggplot2, grid, gridExtra, gridSVG, knitr, maps, mFilter, mnormt, nullabor, stringr and tidyr.
 %\citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
          scale_y_continuous(breaks=c(0,0.5,1)) +
-         labs(y="Proportion",x="Correlation between series") + theme(legend.position="none")
+         labs(y="Proportion Selected",x="Correlation between series") + theme(legend.position="none")
 #ggplot(data=subset(pin, design == "Scatterplot")) + 
@@
 
-Occasionally there was an anomaly where a single null plot was selected with a high frequency despite having a weak correlation.
-Occasionally the null distribution cangenerate visually interesting patterns that draw the analyst’s attention.
+Occasionally there was an anomaly where a single null plot was selected with a high frequency despite having a weak correlation, as the null distribution can generate visually interesting patterns that draw the analyst’s attention.
 We asked subjects for reasons behind their selections and found that choosing on the basis that “the points in the plot were close together”, or had “a unique pattern” more often led to an unsuccessful detection.
@@
 Taking these into account, the scatter plot shows superior performance relative to the overlaid lines at higher values of abs($\rho$).
-It needs to be restated that we only tested correlations at  $0.3, 0.5, 0.7$ and $0.9$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
+It needs to be restated that we only tested correlations at an absolute value $\in \{0.3, 0.5, 0.7, 0.9\}$, so the results below an absolute value of $0.3$ are unreliable and may be artifacts of regression smoothing methods.
 
 %The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
-Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (Figure~\ref{smoothing}.
+Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, demonstrated by $\beta_{4} + \beta_{10}$, but it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (See Figure~\ref{smoothing}.
 Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
 
-Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y axis) plotted against correlation (x axis) for each of the four sample sizes tested.
+Figure~\ref{powermodel} shows the fitted unrestricted model, with power predictions (y-axis) plotted against correlation (x-axis) for each of the four sample sizes tested.
 The thicker lines represent the fixed effect fit, or the average performance across all tested individuals whilst the thinner lines represent the random effects fit, the variation in power that can be attributed to an individual's visual skill.
@@
 We did not include these characteristics of the individual in the modelling due to this lack of significance.
-Figure~\ref{powermodel} especially highlights the weak performance of the overlaid lines and the benefits of smoothing.
+Figure~\ref{powermodel} especially highlights the weak performance of the overlaid lines and the benefits of smoothing and increased sample sizes.
 
@@
 The thicker lines shows the fixed effects prediction while the thinner lines show the range of random effects, the observer's visual ability.
-The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
+The weakness of the overlaid line graph in negative correlation and gain from increasing scatter plot sample size or smoothing overlaid lines are highlighted.}
   \label{powermodel}

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 
-Figures~\ref{anom1} and \ref{anom2} show lineups where a null plot with a strong pattern led to its selection over the data plot.
-Mostly, when a null plot was chosen over the data plot it was because the correlation was stronger, but in a few lineups a null plot with small correlation was selected by many subjects.
+Figures~\ref{anom1} and \ref{anom2} show examples of this in some lineups where a null plot with an unusually strong pattern led to selection of a null over the data plot.
 In Figure~\ref{anom1}, the true data plot is in position 4, but 9 out of 11 subjects chose plot 6.

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 The dots are data points, the lines represent a naive model average fit.
-Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at revealing negative correlation.}
+Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at displaying negative correlation.}
     \label{results}
@@
  \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a joint 95\% confidence interval.
-The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid lines power drops substantially.}
+The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid line graph's power drops substantially.}
  \label{powerdiff}
@@
 @
- \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data.
-The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
+ \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size with unsmoothed data.
+The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines have a mixed benefit, with improvements only in positive correlation.}
  \label{samplesize}
@@
 Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, demonstrated by $\beta_{4} + \beta_{10}$, but it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (See Figure~\ref{smoothing}.
-Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.
+Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph possibly due to the negative correlation results interfering with the modeling (See Figure~\ref{samplesize}).
 

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 
-If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability $p_{j} = s/m$, where that observer made $s$ selections out of a lineup with $m$ plots \citep{Follett}.
-In the event that a lineup has observers making different numbers of selections, we have a binomial distribution with different success probabilities, and a Poisson Binomial distribution must be used to calculate the p-value.
+If a subject selects multiple plots from the lineup, the p-value has a Bernoulli distribution with a probability $p_{j} = s/m$, where that observer made $s$ selections out of a lineup with $m$ plots.
+In the event that a lineup has observers making different numbers of selections, we have a binomial distribution with different success probabilities, and a Poisson Binomial distribution must be used to calculate the p-value demonstrated by \citet{Follett}.
 The probability mass function of a Poisson Binomial distribution can be written as:
@@
 \centering
-P(X=x) = \sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})
+P(X \geq x) = 1 - P(X < (x - 1)) = 1 - \sum^{x-1}_{i=0} \left\{\sum_{A \in T_{x}} \prod_{j \in A} p_{j} \prod_{j \in A^{c}} (1-p_{j})\right\}
 \end{equation}

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 Where $T_{x}$ is all possible sets of $x$ trials out of $K$ resulting in a success, $A \in T_{x}$ is a specific set of $x$ trials resulting in a success and $A^{c}$ is the complement to $A$, the set of $K-x$ trials resulting in a failure.
+This can easily be calculated using the poibin package for R.
 
@@
 
-Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: dplyr, ggplot2, grid, gridExtra, gridSVG, knitr, maps, mFilter, mnormt, nullabor, stringr and tidyr.
-%\citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{stringr} and \citet{tidyr}.
+Further analysis was carried out using \citet{R} version 3.1.2 and \citet{Rstudio} version 0.99.484 using the packages: dplyr, ggplot2, grid, gridExtra, gridSVG, knitr, maps, mFilter, mnormt, nullabor, poibin, stringr and tidyr.
+%\citet{dplyr}, \citet{ggplot2}, \citet{grid}, \citet{gridExtra}, \citet{gridSVG}, \citet{knitr}, \citet{maps}, \citet{mFilter}, \citet{mnormt}, \citet{nullabor}, \citet{poibin}, \citet{stringr} and \citet{tidyr}.
 

### 2016-02-09 paper/paper.Rnw
editing section 4

@@
 full <- full %>% mutate(design = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"),
-                         r = round(correlation, 1),
+                         r = round(correlation, 1), #remove slight variations in correlation caused by simulation
                          rsgn = ifelse(r >= 0, 1, 0),
                          rabs = abs(r),
-                         alpha = log(0.05/0.95),
+                         alpha = log(0.05/0.95), #an offset used to fit the model
                          attempts = sapply(strsplit(as.character(full$response_0), ","), length), 
@@
   ##split multiple selections in response_0 into different rows
-  mutate(correct = ifelse(response_0 == actual_location, 1, 0), 
+  mutate(correct = ifelse(response_0 == actual_location, 1, 0), #ensure only the true detection in mult. selections has a 1
          score = ifelse(correct==1, (20-attempts)/19, 0)) %>% 
@@
 \begin{itemize}
-  \item Factor 1: Plot design, levels = scatter plot, overlaid time series
-  \item Factor 2: Sample size (t), levels = \{12, 24, 48, 96\}
-  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$
+  \item Factor 1: Plot design, levels = scatter plot, overlaid time series.
+  \item Factor 2: Sample size (t), levels = $\{12, 24, 48, 96\}$.
+  \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$.
 \end{itemize}
@@
 \begin{itemize}
-  \item Factor 4: Smoothed, levels = yes, no
+  \item Factor 4: Smoothed, levels = yes, no.
 \end{itemize}
@@
 If it was detected, the trial data was discarded and the subject moved onto the rest of the lineups.
-Earlier work by \citet{Majumder:2014} has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, justifying our treatment of the results as independent evaluations.
+Earlier work by \citet{Majumder:2014} has found that repeated evaluations from a subject does not increase their ability to detect the true plot from a lineup, however random effects were used to model the dependence between different evaluations from the same subject.
 We recruited subjects from Amazon's Mechanical Turk service and ensured that each lineup was viewed by multiple subjects.
-In total there were 2091 lineup evaluations, however to better control each subject’s individual visual ability we removed data from subjects with less than ten individual evaluations.
-Some subjects chose to participate in the experiment more than once, so we removed all data from repeated trials by these subjects.
+In total there were 2091 lineup evaluations, however to better control correlation between subject’s individual visual ability we removed data from subjects with less than ten individual evaluations.
+Some subjects chose to participate in the experiment more than once, so we removed all data from any repeated trials leaving only the original ten evaluations.
 We were left with 1600 evaluations by 160 subjects.
@@
 
+If the true plot correlation were equal to $0$, it should be indistinguishable from the null plots and the detection rate would fall to $1/20$, the Type I error rate.
+An offset is included in the model to incorporate this effect.
+
 If a subject made multiple selections from the same lineup, each selection is treated as a separate data point with a probability weighting equal to the inverse of the number of selections.
@@
 \begin{table}[htp]
-\caption{Weighted Mixed Effects Logit Model for the predicted power of plot design.
+\caption{Weighted Mixed Effects Logit Model for the predicted power of both plot designs.
 Sample size for scatter plot evaluations: 859, overlaid lines: 903, total: 1762.
@@
 \begin{table}[htp]
-\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
+\caption{Weighted Mixed Effects Linear Model for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
 \label{tab2}
@@
 
-The models used for this paper were generated using SAS software, Version 9 of the SAS System for Unix with the MIXED and GLIMMIX procedures.
+The models used for this paper were generated using SAS software, Version 9 of the SAS System for Unix, with the MIXED and GLIMMIX procedures.
 Copyright \copyright\ 2015 SAS Institute Inc.

### 2016-02-09 paper/paper.Rnw
editing

@@
     <<nzd, error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
-    library(dplyr)
-    library(tidyr)
-    library(ggplot2)
     rates <- select(read.csv("data/rates.csv"), date, AUD, CNY, NZD)[36:69,]
@@
   \caption{Plots illustrating the experimental design.
-The same AR(1) data both unsmoothed (\subref{smootha} and \subref{smoothb}) and smoothed (\subref{smoothc} and \subref{smoothd}) for both plot designs, with t = 96 and correlation = 0.86}
+The same AR(1) data both unsmoothed (\subref{smootha} and \subref{smoothb}) and smoothed (\subref{smoothc} and \subref{smoothd}) for both plot designs, with t = 96 and correlation $\rho$ = 0.86}
   \label{smoothing}
@@
 \end{subfigure}
-\caption{Plots illustrating the experimental design: the two different plot types of the same data with $t=24$, $r= 0.946$ (\subref{designa} and \subref{designc}), $r= -0.943$ (\subref{designb} and \subref{designd}).}
+\caption{Plots illustrating the experimental design: the two different plot types of the same data with $t=24$, $\rho= 0.946$ (\subref{designa} and \subref{designc}), $\rho= -0.943$ (\subref{designb} and \subref{designd}).}
 \label{designs}

### 2016-02-09 paper/paper.Rnw
editing

@@
 #write.csv(sasdata, "sas/sasdata.csv", row.names = FALSE)
-#The code that is exported to fit the model into SAS, see sas/model fit.sas
+#The code that is exported to fit the model into SAS, see code/fitting model.sas
 

### 2016-02-09 paper/supplementary.Rnw
editing

@@
 The most selections made on any lineup evaluation is 8.
+No subject demographic had an increased rate of making multiple selections.
 
@@
 
-\begin{figure}[htbp]
+We originally hypothesised that factor combinations with a low power are most likely to have multiple selections due to a decreased ability to differentiate between true and null plots.
+This would cause most lineups with multiple selection lineups to be concentrated around $\rho = \pm 0.3$ for both designs as well as any negative value for overlaid line graphs.
+Figure ~\ref{lineupmulti} plots lineups with at least one case of a subject selecting multiple plots in an evaluation.
+The hypothesised effect only appears at sample sizes of 48 and 96.
+The overlaid line graph lineups were more likely to have multiple selections, but this is also occuring in high positive correlations, where the power was similar to that of a scatter plot.
+
+\begin{figure}[h]
 \centering
@@
 \caption{Characteristics of lineups with at least one subject making multiple selections.
-Overlaid line lineups have multiple selections more often, and true plot correlation does not appear to have an impact except at large sample sizes.}
+Overlaid line graph lineups have subjects make multiple selections more often, but true plot correlation does not appear to have an impact except at large sample sizes.}
 \label{lineupmulti}
@@
 \caption{Number of times $n$ selections from a lineup evaluation were made.
-The subject who made the most selections is additionally reported.
+The subject who made the most plot selections within their ten evaluations is additionally reported.
 There is no relationship between a subject's demographics and a likelihood to make multiple selections.}
@@
 
-Analysing the level one residuals, $\hat{\epsilon_{i}}$, and level two residuals, $\hat{\gamma_{i}}$, allows us to test the model fit as described \citet{Loy}.
-There are some negative extreme outliers but the distributional assumptions appear to hold.
+Analysing the level one residuals, $\hat{\epsilon_{i}}$, and level two residuals, $\hat{\gamma_{i}}$, allows us to test the model fit as described by \citet{Loy}.
+There are some extreme negative outliers but the distributional assumptions appear to hold.
 
@@
 
-$${E[score]}_{i} = \frac{1}{K}\sum_{j}{score}_{ij}$$
+$$E[{score}_{i}] = \frac{1}{K}\sum_{j}{score}_{ij}$$
 
@@
 
-$${E[score]}_{j} = \sum_{i}\{{E[score]}_{i} | i \in L_{j}\}$$
+$$E[{score}_{j}]  = \sum_{i}\{E[{score}_{i}] | i \in L_{j}\}$$
 
-Finally, we measure visual ability by:
+Finally, we proxy visual ability by:
 
-$${Visual Ability}_{j} = {score}_{j}-{E[score]}_{j}$$.
+$${Visual Ability}_{j} = {score}_{j}-E[{score}_{j}] $$.
 

### 2016-02-09 paper/supplementary.Rnw
supp material

@@
 @
-\caption{Age and education by gender. Younger age brackets and males are over-represented in the sample, but these factors caused no significant impact on results.}
+\caption{Age and education by gender. Younger age brackets, males and people with college education are over-represented in the sample, but these factors caused no significant impact on results.}
 \label{ageed}
@@
 Where $m = 20$, the number of plots in a linup, $s_{ij}$ is the number of selections made by subject $i$ on lineup $j$ and $\mathbbm{I}_{ij} = 1$ if the true plot is detected, $0$ otherwise.
-The most selections made on any lineup evaluation is 8.
-No subject demographic had an increased rate of making multiple selections.
 
@@
 subjplot <- mutate(subjplot, avscore= total/eval)
-ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.1)
+ggplot(data=subjplot, aes(x=avscore)) + geom_histogram(binwidth=0.1) + labs(y=NULL, x="Average Score")
 @
-\caption{The average score of our subjects, according to formula \ref{score}. There is a wide range of visual ability to be controlled with random effects in the modelling}
+\caption{The average score of our subjects, according to formula \ref{score}. There is a wide range of differing visual abilities to be controlled with random effects in the modelling}
 \label{avscore}
@@
 
-\section{Evaluation Descriptive Statistics}
+\section{Lineup Evaluation Descriptive Statistics}
 
@@
 ggplot(data=choice, aes(x=choice_reason, fill=correct)) + geom_bar() + facet_grid(design~rsgn) +
-  theme(legend.position="bottom", axis.title=element_blank()) + scale_fill_discrete(name="True Plot")
+  theme(legend.position="bottom", axis.title=element_blank(), axis.text.x = element_text(angle = 270)) + scale_fill_discrete(name="True Plot")
 @
-\caption{Reasons given for plot selection. After controlling for lineup factors, there is no explanatory power.}
+\caption{Reasons given for plot selection. After controlling for lineup factors, these have no explanatory power.}
 \label{reasons}
@@
 m.lineup <- merge(m.freq, bylineup, by.x="Var1", by.y="pic_id")
-ggplot(data=m.lineup) + geom_bar(aes(x=factor(r), fill=design)) + facet_grid(n~smoothed) + labs(y="Count", x="True plot correlation")
+ggplot(data=m.lineup) + geom_bar(aes(x=factor(r), fill=design)) + facet_grid(n~smoothed) + labs(y=NULL, x="True plot correlation")
 @

### 2016-02-09 paper/paper.Rnw
supp material

@@
 
-To allow for a greater control of distributions of the data in the lineups, we use simulated data.
-The data must contain autocorrelation for there to be meaningful information in the time dimension.
-To fulfil this requirement, each plot has a standardised pair of AR(1) models defined by:
+For there to be meaningful information in the time dimension, the data must have autocorrelation. To fulfil this requirement, each plot has a pair of standardised AR(1) models generated by:
 
@@
 
-With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ being generated with from a bivariate normal distribution:
+With $t \in \left\{12, 24, 48, 96\right\}$ and $e_{t}$ generated from a bivariate normal distribution:
 
@@
 The variation in the covariance parameter allows for control in the correlation in the pairs of data.
-The null data has no structure, so uncorrelated pairs are produced by setting $\rho = 0$.
+The null data has no underlying structure, so uncorrelated pairs are produced by setting $\rho = 0$.
 True data pairs are generated with $\rho \in \pm \{0.3, 0.5, 0.7, 0.9\}$.
@@
 Simulating data does not produce correlations exactly equal to $\rho$, so simulated data pairs were accepted as real data if pairwise correlation between the two series is within 0.015 of the desired value of $\rho$.
-The null pairs often had correlation generated spuriously through the simulation process.
-This was particularly problematic for generating small sample data, but setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
+The null pairs often had correlation generated spuriously through the simulation process, which was particularly problematic for small sample data.
+Setting $\beta = min(0.5,100t)$ reduced spurious correlation to manageable levels.
 For particularly large time series, the AR(1) model can create particularly 'jagged' lines that may be unrealistic for many actual applications.
-To counter this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component.
+To reduce this, additional sets of data were created with a Hodrick-Prescott filter to capture the trend of the model and remove the noise in the cycle component.
 This filter creates a trend by penalising both errors in smoothness and in fit.
-The trend of the series, $\tau_{t}$ is defined by the equation:
+The trend of each series $y_{t}$, $\tau_{t}$ is defined by the equation:
 
@@
 Setting $\lambda = 1$ produced a smoothing that is more consistent with real world time series applications and did not introduce excessive spurious null correlation.
-However, for $t = \{12, 24\}$, any form of smoothing was unfeasible without introducing extreme null correlation, hence smoothing was only used for $t = \{48, 96\}$.
-Each generated lineup was produced as both a scatter plot and an overlaid line graph.
+However, for $t = \{12, 24\}$, smoothing becomes unfeasible without introducing extreme null correlation, often with an absolute value greater than 0.8.
+Smoothing was only used for $t = \{48, 96\}$.
+Each real data set was placed with a collection of 19 null plots and produced as both a scatter plot and an overlaid line graph.
 
@@
   \item Factor 1: Plot design, levels = scatter plot, overlaid time series.
-  \item Factor 2: Sample size (t), levels = $\{12, 24, 48, 96\}$.
+  \item Factor 2: Sample size, t, levels = $\{12, 24, 48, 96\}$.
   \item Factor 3: True plot correlation, $\rho$, levels = $\pm \{0.3, 0.5, 0.7, 0.9\}$.

### 2016-02-15 paper/supplementary.Rnw
supp material
touch up

@@
 \centering
-\caption{Number of times $n$ selections from a lineup evaluation were made.
+\caption{Number of times $n$ selections were made in a lineup evaluation.
 The subject who made the most plot selections within their ten evaluations is additionally reported.
@@
 @
-\caption{Visual ability plotted against random effects.
-The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
+\caption{Visual ability plotted against random effects. The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model suitably measure visual ability.}
 \label{random2}
